---
title: "Causal Survival - Case Base"
output: pdf_document
date: "2026-05-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newpage
# Imports

### EDA
```{r, warning=FALSE, message=FALSE}
library("tableone")
library("xtable")
```

### Cox PH Model
```{r, warning=FALSE, message=FALSE}
library("survival")
```

### SMD Balance
```{r, warning=FALSE, message=FALSE}
library("survey")
library("cobalt")
```

### Case Base
```{r, warning=FALSE, message=FALSE}
library("casebase")
```

\newpage
# EDA

## Data Set
```{r}
  #set seed
  set.seed(2026)
  
  #re-factor the data set
  valid.indices <- seq(2, nrow(colon), by = 2)
  
  #main variables of interest
  time <- colon$time[valid.indices]
  Delta <- colon$status[valid.indices]
  A <- ifelse(
    colon$rx[valid.indices] == "Obs",
    0,
    1
  )
  
  #covriates
  Gender <- colon$sex[valid.indices]
  Age <- colon$age[valid.indices]
  Obstruction <- colon$obstruct[valid.indices]
  Perforation <- colon$perfor[valid.indices]
  Adherence <- colon$adhere[valid.indices]
  NumNodes <- colon$nodes[valid.indices]
  Differentiation <- colon$differ[valid.indices]
  Extent <- colon$extent[valid.indices]
  Surgery <- colon$surg[valid.indices]
  Nodes4 <- colon$node4[valid.indices]
  
  #data set design
  data.set <- data.frame(
    time = time,
    Delta = Delta,
    A = A,
    Gender = Gender,
    Age = Age,
    Obstruction = Obstruction,
    Perforation = Perforation,
    Adherence = Adherence,
    NumNodes = NumNodes,
    Differentiation = Differentiation,
    Extent = Extent,
    Surgery = Surgery,
    Nodes4 = Nodes4
  )
```

\newpage
## Table One
```{r}
covariates = c(
  "Gender",
  "Age",
  "Obstruction",
  "Perforation",
  "Adherence",
  "NumNodes",
  "Differentiation",
  "Extent",
  "Surgery",
  "Nodes4"
)

table.one.original <- CreateTableOne(
  vars = covariates,
  strata = "A",
  data = data.set,
  test = FALSE
)

print(table.one.original, smd = TRUE)
```
\newpage
## Propensity Scores
```{r}
#generate propensity scores
propensity.score <- fitted(
  glm(
    A ~ Gender +
      Age + 
      Obstruction + 
      Perforation + 
      Adherence + 
      NumNodes + 
      Differentiation + 
      Extent + 
      Surgery + 
      Nodes4,
    data = data.set,
    family = binomial(link = "logit")
  )
)

#separate propensity scores for exposed & unexposed groups
propensity.score.0 <- propensity.score[A == 0]
propensity.score.1 <- propensity.score[A == 1]

#visualize the propensity scores overlap
par(
  mar=c(2,2,2,0)
)

hist(
  propensity.score.0,
  col=rgb(0,0,1,0.5),
  breaks=seq(0,1,by=0.02),
  ylim=c(0,175),
  main="Propensity Score Overlap",
  xlab="Propensity Scores"
)

hist(
  propensity.score.1,
  col=rgb(1,0,0,0.5),
  breaks=seq(0,1,by=0.02),
  add=T
)

box()

legend(
  "topright",
  legend=c('A = 0', 'A = 1'),
  fill=c(rgb(0,0,1,0.5),rgb(1,0,0,0.5)),
  border="black", bty="n"
)
```

\newpage
## Stabilized IPW
```{r}
p.A1 <- mean(A)

#calculate stabilized ipw
ipw <- ifelse(
  A == 1,
  p.A1 / propensity.score,
  (1 - p.A1) / (1 - propensity.score)
)
```
\newpage
## Covariate Balance
```{r}
#calculate the covariate balance before & after IPW
balance <- bal.tab(
  A ~ Gender + 
    Age + 
    Obstruction + 
    Perforation +
    Adherence +
    NumNodes +
    Differentiation +
    Extent +
    Surgery +
    Nodes4,
  data = data.set,
  weights = ipw,
  method = "weighting",
  estimand = "Average Treatment Effect",
  un = TRUE
)

#visualize the covariate balance before & after IPW
love.plot(
  balance,
  stats = "mean.diffs",
  threshold = 0.1,
  abs = TRUE,
  line = TRUE,
  var.order = "unadjusted",
  colors = c("red", "blue"),
  shapes = c(17, 16),
  limits = c(m = c(0, 0.7)), #hard-coded 0.7??
  title = "Covariate Balance Before and After IPW",
  labels = TRUE
)
```

\newpage
## Table Two
```{r}
data.set.ipw <- svydesign(
  ids = ~ 0,
  data = data.set,
  weights = ipw
)

table.one.ipw <- svyCreateTableOne(
  vars = covariates,
  strata = "A",
  data = data.set.ipw,
  test = FALSE
)

print(table.one.ipw, smd = TRUE)
```
\newpage
# Estimation

## Assumptions
*Non-Informative Censoring*

\newpage
## Model 1: Unadjusted Cox
```{r}
#fit model
model.1 <- coxph(
  Surv(time, Delta) ~ A,
  data = data.set,
  robust = TRUE
)

#generate summary
summary(model.1)
```
\newpage
## Model 2: Covariate Adjusted Cox
```{r}
#fit model
model.2 <- coxph(
  Surv(time, Delta) ~ A +
    Gender +
    Age +
    Obstruction +
    Perforation +
    Adherence +
    NumNodes +
    Differentiation +
    Extent +
    Surgery +
    Nodes4,
  data = data.set,
  robust = TRUE
)

#generate summary
summary(model.2)
```
\newpage
## Model 3: Marginal Structural Cox
```{r}
#fit model
model.3 <- coxph(
  Surv(time, Delta) ~ A,
  data = data.set,
  weights = ipw,
  robust = TRUE
)

#generate summary
summary(model.3)
```
\newpage
## Model 4: IPW Case Base PH
```{r}
#fit model
model.4 <- fitSmoothHazard(
  Delta ~ time * A,
  data = data.set,
  ratio = 10,
  weights = ipw
)

#generate summary
summary(model.4)
```
\newpage
## Model 5: IPW Case Base (Variable Hazard over Time)
```{r}
#fit model
model.5 <- fitSmoothHazard(
  Delta ~ pspline(time, df = 2) * A,
  data = data.set,
  ratio = 10,
  weights = ipw
)

#generate summary
summary(model.5)
```