---
title: "Causal RMST under case-base sampling with IPW: Right Heart Catheterization"
author: "Bilal Shaikh and Sumeet Kalia"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = FALSE, warning = FALSE, message = FALSE)
```

\newpage
# Configuration
```{r}
SEED <- 20260811
CB_RATIO <- 20
TAU_MAX <- 180
TAU_GRID <- seq(0, TAU_MAX, by = 1)
TAU_REPORT <- c(30, 90, 180)
ALPHA <- 0.05
Z_CRIT <- qnorm(1 - ALPHA / 2)
WEIGHT_TRUNCATE <- c(0.01, 0.99)

set.seed(SEED)

dir.create("Figures", recursive = TRUE, showWarnings = FALSE)
dir.create("Tables", recursive = TRUE, showWarnings = FALSE)
```

\newpage
# Plot theme
A single minimalist theme and a single palette are used for every figure so that
the panels remain legible in greyscale. Colour and shape are mapped to the same
grouping variable throughout, so each series is identified twice.

```{r}
palette <- c("red", "blue", "green4", "black")
shapes <- c(16, 17, 15, 3)
line_types <- c("solid", "dashed", "dotdash", "dotted")

plot_theme <- function(base.size = 10) {
  ggplot2::theme_minimal(base_size = base.size) +
    ggplot2::theme(
      panel.grid.minor = ggplot2::element_blank(),
      panel.grid.major = ggplot2::element_line(
        colour = "grey90", linewidth = 0.3
      ),
      panel.border = ggplot2::element_rect(
        colour = "grey30", fill = NA, linewidth = 0.4
      ),
      strip.background = ggplot2::element_blank(),
      strip.text = ggplot2::element_text(face = "bold", size = base.size),
      axis.text = ggplot2::element_text(colour = "grey20"),
      legend.position = "bottom",
      legend.direction = "horizontal",
      legend.title = ggplot2::element_blank(),
      legend.key.width = ggplot2::unit(1.4, "lines"),
      legend.margin = ggplot2::margin(t = 0)
    )
}

legend <- ggplot2::guides(
  colour = ggplot2::guide_legend(nrow = 1, ncol = 2, byrow = TRUE),
  fill = ggplot2::guide_legend(nrow = 1, ncol = 2, byrow = TRUE),
  shape = ggplot2::guide_legend(nrow = 1, ncol = 2, byrow = TRUE),
  linetype = ggplot2::guide_legend(nrow = 1, ncol = 2, byrow = TRUE)
)
```

\newpage
# Data preparation
Time zero is admission to intensive care.
The treatment is right heart catheterization within the first twenty-four hours.
Follow-up is administratively censored at `TAU_MAX` days.

```{r}
Hmisc::getHdata(rhc)

rhc$time.death <- as.numeric(
  pmax(rhc$dthdte, rhc$lstctdte, na.rm = TRUE) - rhc$sadmdte
)
rhc$time.discharge <- as.numeric(rhc$dschdte - rhc$sadmdte)
rhc$death <- as.integer(rhc$death == "Yes")
rhc$A <- as.integer(rhc$swang1 == "RHC")

hx_vars <- grep("hx$", names(rhc), value = TRUE)
rhc[hx_vars] <- lapply(rhc[hx_vars], function(variable) {
  factor(
    ifelse(as.numeric(variable) == 1, "Yes", "No"), levels = c("No", "Yes")
  )
})

organ_vars <- c(
  "resp", "card", "neuro", "gastr", "renal", "meta", "hema",
  "seps", "trauma", "ortho"
)
rhc[organ_vars] <- lapply(rhc[organ_vars], function(variable) {
  droplevels(as.factor(variable))
})

continuous_vars <- c(
  "age", "edu", "das2d3pc", "aps1", "scoma1", "surv2md1",
  "wtkilo1", "temp1", "meanbp1", "resp1", "hrt1", "pafi1",
  "paco21", "ph1", "wblc1", "hema1", "sod1", "pot1",
  "crea1", "bili1", "alb1", "urin1"
)
categorical_vars <- c(
  "sex", "race", "income", "ninsclas", "ca", "cat1", "cat2","dnr1",
  hx_vars, organ_vars
)

impute_median <- function(variable) {
  miss <- (is.na(variable) | variable == 0)
  variable[miss] <- median(variable, na.rm = TRUE)
  list(value = variable, missing = as.integer(miss))
}

covariate_frame <- data.frame(row.names = seq_len(nrow(rhc)))
for (variable in continuous_vars) {
  imputed_vector <- impute_median(as.numeric(rhc[[variable]]))
  covariate_frame[[variable]] <- imputed_vector$value
  if (any(imputed_vector$missing == 1)) {
    covariate_frame[[paste0(variable, ".miss")]] <- imputed_vector$missing
  }
}
for (variable in categorical_vars) {
  copy <- rhc[[variable]]
  copy <- droplevels(as.factor(copy))
  if (any(is.na(copy))) {
    copy <- addNA(copy)
    levels(copy)[is.na(levels(copy))] <- "Missing"
  }
  covariate_frame[[variable]] <- copy
}

covariate_frame <- covariate_frame[
  , vapply(
    covariate_frame,
    function(covariate) length(unique(covariate)) > 1,
    logical(1)
  )
]

data_set <- data.frame(
  A = rhc$A,
  time.death = rhc$time.death,
  death = rhc$death,
  time.discharge = rhc$time.discharge,
  covariate_frame
)

data_set <- data_set[
  is.finite(data_set$time.death) & data_set$time.death >= 0, 
]
data_set$time.death[data_set$time.death == 0] <- 0.5

covariate_names <- setdiff(names(covariate_frame), character(0))
p_score_formula <- as.formula(
  paste("A ~", paste(covariate_names, collapse = " + "))
)
```

\newpage
## Follow-up under each analytic principle
Under the intention-to-treat analogue the outcome is death, with administrative
censoring at `TAU_MAX` days and censoring at last contact. Under the per-protocol
analogue each individual is additionally censored at discharge from the study
hospital, which is the point at which continued management under the assigned
strategy ends.

```{r}
build_followup <- function(data.set, principle = c("itt", "pp")) {
  principle <- match.arg(principle)
  admin.censor.time <- pmin(data.set$time.death, TAU_MAX)
  event <- as.integer(data.set$death == 1 & data.set$time.death <= TAU_MAX)
  
  if (principle == "pp") {
    discharge.time <- ifelse(
      is.finite(data.set$time.discharge) & data.set$time.discharge > 0,
      data.set$time.discharge, Inf
    )
    stop.time <- pmin(admin.censor.time, discharge.time)
    event <- as.integer(event == 1 & admin.censor.time <= discharge.time)
    data.set$time <- stop.time
  
  } else {
    data.set$time <- admin.censor.time
  }
  
  data.set$status <- event
  data.set$time[data.set$time <= 0] <- 0.5
  data.set$cens <- as.integer(data.set$status == 0 & data.set$time < TAU_MAX)
  return(data.set)
}

itt_data <- build_followup(data_set, "itt")
pp_data <- build_followup(data_set, "pp")
```

\newpage
# Propensity score and weights
```{r}
truncate_weights <- function(weights, probs = WEIGHT_TRUNCATE) {
  weight.limits <- quantile(weights, probs = probs, na.rm = TRUE)
  return(pmin(pmax(weights, weight.limits[1]), weight.limits[2]))
}

p_score_fit <- glm(p_score_formula, data = data_set, family = binomial)
p_score_treat <- pmin(pmax(fitted(p_score_fit), 1e-6), 1 - 1e-6)
p_marginal <- mean(data_set$A)

iptw <- ifelse(
  data_set$A == 1, p_marginal / p_score_treat,
  (1 - p_marginal) / (1 - p_score_treat)
)
```

\newpage
# Censoring Weights
Censoring weights are constructed from a Cox model for the censoring hazard
fitted separately within each treatment arm, evaluated at each individual's
observed follow-up time. The stabilized weight uses a numerator model that
conditions on treatment alone.

```{r}
censoring_weights <- function(data.set) {
  weights <- numeric(nrow(data.set))
  for (a.value in c(0, 1)) {
    row.idx <- which(data.set$A == a.value)
    arm.data <- data.set[row.idx, ]
    denom.form <- as.formula(
      paste("Surv(time, cens) ~", paste(covariate_names, collapse = " + "))
    )
    denom.fit <- survival::coxph(denom.form, data = arm.data)
    base.denom <- survival::basehaz(denom.fit, centered = FALSE)
    lp.denom <- predict(
      denom.fit, newdata = arm.data, type = "lp", reference = "zero"
    )
    h.denom <- approx(
      base.denom$time, base.denom$hazard, xout = arm.data$time,
      method = "constant", rule = 2
    )$y
    surv.denom <- exp(-h.denom * exp(lp.denom))

    num.fit <- survival::survfit(Surv(time, cens) ~ 1, data = arm.data)
    surv.num <- approx(
      num.fit$time, num.fit$surv, xout = arm.data$time,
      method = "constant", rule = 2
    )$y

    weights[row.idx] <- surv.num / pmax(surv.denom, 1e-6)
  }
  return(weights)
}

itt_ipcw <- truncate_weights(censoring_weights(itt_data))
pp_ipcw <- truncate_weights(censoring_weights(pp_data))

itt_data$w.iptw <- truncate_weights(iptw)
itt_data$w.both <- truncate_weights(itt_data$w.iptw * itt_ipcw)

pp_data$w.iptw <- truncate_weights(iptw)
pp_data$w.both <- truncate_weights(pp_data$w.iptw * pp_ipcw)

weight_row <- function(principle, label, weights) {
  return(data.frame(
    principle = principle, weight = label,
    mean = mean(weights),
    p99 = unname(quantile(weights, 0.99)),
    max = max(weights),
    ess.pct = 100 * (sum(weights)^2 / sum(weights^2)) / length(weights),
    row.names = NULL
  ))
}

weight_summary <- rbind(
  weight_row("ITT", "Treatment", itt_data$w.iptw),
  weight_row("ITT", "Censoring", itt_ipcw),
  weight_row("ITT", "Product", itt_data$w.both),
  weight_row("PP", "Treatment", pp_data$w.iptw),
  weight_row("PP", "Censoring", pp_ipcw),
  weight_row("PP", "Product", pp_data$w.both),
  weight_row("Untruncated", "Treatment", iptw)
)

weight_summary

print(
  xtable::xtable(weight_summary, digits = c(0, 0, 0, 3, 3, 3, 1)),
  include.rownames = FALSE, file = "Tables/weights.tex"
)
```

\newpage
# Exploratory data analysis
```{r}
table_unweighted <- tableone::CreateTableOne(
  vars = covariate_names, strata = "A", data = data_set, test = FALSE
)
print(table_unweighted, smd = TRUE)

design_weighted <- survey::svydesign(
  ids = ~ 1, data = data_set, weights = ~ iptw
)
table_weighted <- tableone::svyCreateTableOne(
  vars = covariate_names, strata = "A", data = design_weighted, test = FALSE
)
print(table_weighted, smd = TRUE)

smd_for <- function(weights) {
  if (is.null(weights)) {
    table.result <- tableone::CreateTableOne(
      vars = covariate_names, strata = "A", data = data_set, test = FALSE
    )
  } else {
    survey.design <- survey::svydesign(
      ids = ~ 1, data = cbind(data_set, .w = weights), weights = ~ .w
    )
    table.result <- tableone::svyCreateTableOne(
      vars = covariate_names, strata = "A", data = survey.design, test = FALSE
    )
  }
  smd <- tableone::ExtractSmd(table.result)
  data.frame(covariate = rownames(smd), smd = abs(as.numeric(smd)))
}

smd_raw <- smd_for(NULL)
smd_iptw <- smd_for(itt_data$w.iptw)
smd_both <- smd_for(itt_data$w.both)

balance_long <- rbind(
  cbind(smd_raw, adjustment = "Unweighted"),
  cbind(smd_iptw, adjustment = "IPTW"),
  cbind(smd_both, adjustment = "IPTW x IPCW")
)

balance_long$adjustment <- factor(
  balance_long$adjustment, levels = c("Unweighted", "IPTW", "IPTW x IPCW")
)

covariate_order <- smd_raw$covariate[order(smd_raw$smd)]
balance_long$covariate <- factor(
  balance_long$covariate, levels = covariate_order
)

ggplot2::ggplot(
  balance_long, ggplot2::aes(
    x = smd, y = covariate, colour = adjustment,
    shape = adjustment, group = adjustment
  )
) + 
  ggplot2::geom_vline(xintercept = 0.1, linetype = "dashed", colour = "grey40") + 
  ggplot2::geom_line(orientation = "y", linewidth = 0.35, alpha = 0.8) +
  ggplot2::geom_point(size = 1.7) +
  ggplot2::scale_colour_manual(values = palette[1:3]) +
  ggplot2::scale_shape_manual(values = shapes[1:3]) +
  ggplot2::scale_x_continuous(
    breaks = seq(0, 1, by = 0.1), limits = c(0, max(balance_long$smd) * 1.02)
  ) +
  ggplot2::labs(x = "Absolute standardized mean difference", y = NULL) +
  plot_theme() +
  ggplot2::guides(
    colour = ggplot2::guide_legend(nrow = 1),
    shape = ggplot2::guide_legend(nrow = 1)
  )

ggplot2::ggsave(
  "Figures/covariate-balance.pdf", width = 7, height = 9
)

p_score_cens <- numeric(nrow(itt_data))

for (a_value in c(0, 1)) {
  row_idx <- which(itt_data$A == a_value)
  cens_fit <- glm(
    as.formula(paste("cens ~", paste(covariate_names, collapse = " + "))),
    data = itt_data[row_idx, ], family = binomial
  )
  p_score_cens[row_idx] <- fitted(cens_fit)
}

overlap_frame <- rbind(
  data.frame(
    score = p_score_treat, panel = "Treatment: P(RHC | X)",
    group = factor(data_set$A, levels = c(0, 1), labels = c("No RHC", "RHC"))
  ),
  data.frame(
    score = p_score_cens, panel = "Censoring: P(censored | X, A)",
    group = factor(itt_data$A, levels = c(0, 1), labels = c("No RHC", "RHC"))
  )
)
overlap_frame$panel <- factor(
  overlap_frame$panel,
  levels = c("Treatment: P(RHC | X)", "Censoring: P(Censored | X, A)")
)

ggplot2::ggplot(
  overlap_frame, ggplot2::aes(x = score, fill = group, colour = group)
) +
  ggplot2::geom_histogram(
    bins = 50, alpha = 0.35, position = "identity", linewidth = 0.25
  ) +
  ggplot2::facet_wrap(~ panel, scales = "free_y") +
  ggplot2::scale_fill_manual(values = palette[1:2]) +
  ggplot2::scale_colour_manual(values = palette[1:2]) +
  ggplot2::scale_x_continuous(breaks = seq(0, 1, by = 0.25)) +
  ggplot2::labs(x = "Estimated probability", y = "Count") +
  plot_theme() +
  legend

ggplot2::ggsave(
  "Figures//propensity-overlap.pdf", width = 9, height = 4
)
```

\newpage
# Model fitting
```{r}
interior_knots <- function(event.times, boundary, probs = c(0.25, 0.5, 0.75)) {
  if (length(event.times) < 20) {
    return(numeric(0))
  }
  
  knots <- unique(
    as.numeric(quantile(event.times, probs = probs, names = FALSE))
  )
  knots <- knots[
    knots > boundary[1] +
      .Machine$double.eps & knots < boundary[2] -
      .Machine$double.eps
  ]
  
  return(sort(unique(knots)))
}

fit_casebase <- function(
    data.set, weights, adjust = c("none", "covariate"),
    time.varying = FALSE, ratio = CB_RATIO,
    knot.probs = c(0.25, 0.5, 0.75), seed = SEED  
  ) {
  
  adjust <- match.arg(adjust)
  data.set.augmented <- data.set
  data.set.augmented$w <- weights
  
  if (!is.null(seed)) set.seed(seed)
  cb.data <- casebase::sampleCaseBase(
    data.set.augmented, time = "time", event = "status", ratio = ratio
  )
  boundary <- range(cb.data$time)
  interior.knots <- interior_knots(
    data.set$time[data.set$status == 1], boundary, probs = knot.probs
  )
  time.term <- if (length(interior.knots) > 0) {
    sprintf(
      "splines::bs(time, knots = c(%s), Boundary.knots = c(%s), degree = 3)",
      paste(interior.knots, collapse = ", "), paste(boundary, collapse = ", ")
    )
  } else {
    sprintf(
      "splines::bs(time, df = 3, Boundary.knots = c(%s), degree = 3)",
      paste(boundary, collapse = ", ")
    )
  }
  right.hand.side <- ifelse(
    time.varying, paste0("A * ", time.term), paste0("A + ", time.term)
  )
  
  if (adjust == "covariate") {
    right.hand.side <- paste(
      right.hand.side, "+", paste(covariate_names, collapse = " + ")
    )
  }
  form <- as.formula(paste("status ~", right.hand.side, "+ offset(offset)"))
  fit <- glm(form, data = cb.data, family = binomial, weights = w)
  
  return(list(
    fit = fit,
    vcov = sandwich::vcovHC(fit, type = "HC0"),
    knots = interior.knots,
    boundary = boundary,
    time.varying = time.varying
  ))
}

casebase_design <- function(cb.object, a.value, time.grid, reference.row) {
  new.data <- reference.row[rep(1, length(time.grid)), , drop = FALSE]
  new.data$time <- pmax(time.grid, 1e-6)
  new.data$A <- a.value
  new.data$offset <- 0
  design.matrix <- model.matrix(
    delete.response(terms(cb.object$fit)), data = new.data
  )
  return(design.matrix[, names(coef(cb.object$fit)), drop = FALSE])
}

rmst_from_survival <- function(time.grid, surv.vec) {
  d.t <- diff(time.grid)
  surv.vec.mids <- (head(surv.vec, -1) + tail(surv.vec, -1)) / 2
  return(c(0, cumsum(surv.vec.mids * d.t)))
}

casebase_rmst <- function(cb.object, time.grid, reference.row) {
  d.t <- diff(time.grid)
  surv <- list()
  grad <- list()
  for (a.value in c(0, 1)) {
    design.matrix <- casebase_design(
      cb.object, a.value, time.grid, reference.row
    )
    
    hazard <- as.vector(exp(design.matrix %*% coef(cb.object$fit)))
    cum.haz <- c(0, cumsum(d.t * (head(hazard, -1) + tail(hazard, -1)) / 2))
    
    surv.vec <- exp(-cum.haz)
    
    hazard.weighted.design <- design.matrix * hazard
    
    grad.cum.hazard <- rbind(0, apply(hazard.weighted.design, 2, function(col) {
      cumsum(d.t * (head(col, -1) + tail(col, -1)) / 2)
    }))
    
    grad.surv <- -surv.vec * grad.cum.hazard
    
    grad.rmst <- rbind(0, apply(grad.surv, 2, function(col) {
      cumsum(d.t * (head(col, -1) + tail(col, -1)) / 2)
    }))
    
    surv[[as.character(a.value)]] <- surv.vec
    grad[[as.character(a.value)]] <- grad.rmst
  }
  rmst.grad.diff <- grad[["1"]] - grad[["0"]]
  
  var.matrix <- cb.object$vcov[
    colnames(rmst.grad.diff), colnames(rmst.grad.diff), drop = FALSE
  ]
  
  rmst.se <- sqrt(
    pmax(rowSums((rmst.grad.diff %*% var.matrix) * rmst.grad.diff), 0)
  )
  
  return(data.frame(
    time = time.grid,
    s0 = surv[["0"]], s1 = surv[["1"]],
    rmst =
      rmst_from_survival(time.grid, surv[["1"]]) -
      rmst_from_survival(time.grid, surv[["0"]]),
    se = rmst.se
  ))
}

cox_rmst <- function(cox.fit, time.grid, reference.row) {
  new.data <- reference.row[rep(1, 2), , drop = FALSE]
  new.data$A <- c(0, 1)
  survival.fit <- survival::survfit(cox.fit, newdata = new.data)
  s0 <- approx(
    survival.fit$time, survival.fit$surv[, 1], time.grid,
    method = "constant", rule = 2
  )$y
  s1 <- approx(
    survival.fit$time, survival.fit$surv[, 2], time.grid,
    method = "constant", rule = 2
  )$y
  se.psi <- sqrt(diag(vcov(cox.fit))["A"])
  integrand <- ifelse(s1 > 0, s1 * log(pmax(s1, 1e-300)), 0)
  gradient.vec <- rmst_from_survival(time.grid, integrand)
  
  return(data.frame(
    time = time.grid, s0 = s0, s1 = s1,
    rmst = rmst_from_survival(time.grid, s1) - rmst_from_survival(time.grid, s0),
    se = abs(gradient.vec) * as.numeric(se.psi)
  ))
}

hr_ci <- function(estimate, se) {
  return(sprintf(
    "%.3f (%.3f, %.3f)",
    exp(estimate),
    exp(estimate - Z_CRIT * se),
    exp(estimate + Z_CRIT * se)
  ))
}
```

\newpage
## The ten configurations
```{r}
run_configurations <- function(data.set, principle.label) {
  reference.row <- data.set[1, , drop = FALSE]
  cox.covariate.form <- as.formula(
    paste("Surv(time, status) ~ A +", paste(covariate_names, collapse = " + "))
  )
  results <- list()

  c1.cox.unadjusted <- survival::coxph(
    Surv(time, status) ~ A, data = data.set, robust = TRUE
  )
  c2.cox.covariate.adjusted <- survival::coxph(
    cox.covariate.form, data = data.set, robust = TRUE
  )
  c3.cox.iptw <- survival::coxph(
    Surv(time, status) ~ A, data = data.set, weights = data.set$w.iptw,
    robust = TRUE
  )
  c4.cox.iptw.ipcw <- survival::coxph(
    Surv(time, status) ~ A, data = data.set, weights = data.set$w.both,
    robust = TRUE
  )

  c5.cb.unweighted.crude <- fit_casebase(
    data.set, rep(1, nrow(data.set)), adjust = "none"
  )
  c6.cb.unweighted.covariate <- fit_casebase(
    data.set, rep(1, nrow(data.set)), adjust = "covariate"
  )
  c7.cb.iptw <- fit_casebase(data.set, data.set$w.iptw, adjust = "none")
  c8.cb.iptw.ipcw <- fit_casebase(data.set, data.set$w.both, adjust = "none")

  c9.cox.time.varying <- survival::coxph(
    Surv(time, status) ~ A + tt(A), data = data.set, weights = data.set$w.both,
    robust = TRUE, tt = function(x, t, ...) x * log(pmax(t, 1e-6))
  )
  c10.cb.time.varying <- fit_casebase(
    data.set, data.set$w.both, adjust = "none", time.varying = TRUE
  )

  cox.list <- list(
    C1 = c1.cox.unadjusted, C2 = c2.cox.covariate.adjusted,
    C3 = c3.cox.iptw, C4 = c4.cox.iptw.ipcw
  )
  cb.list <- list(
    C5 = c5.cb.unweighted.crude, C6 = c6.cb.unweighted.covariate,
    C7 = c7.cb.iptw, C8 = c8.cb.iptw.ipcw
  )

  results$hr <- rbind(
    do.call(rbind, lapply(names(cox.list), function(config.key) {
      model.fit <- cox.list[[config.key]]
      return(data.frame(
        principle = principle.label, configuration = config.key,
        model = "MS-Cox", estimate = unname(coef(model.fit)["A"]),
        se = unname(sqrt(diag(vcov(model.fit))["A"]))
      ))
    })),
    do.call(rbind, lapply(names(cb.list), function(config.key) {
      model.fit <- cb.list[[config.key]]
      return(data.frame(
        principle = principle.label, configuration = config.key,
        model = "Case-base", estimate = unname(coef(model.fit$fit)["A"]),
        se = unname(sqrt(diag(model.fit$vcov)["A"]))
      ))
    }))
  )
  results$hr$hr.ci <- hr_ci(results$hr$estimate, results$hr$se)

  results$rmst.cox <- cbind(
    principle = principle.label, configuration = "C4", method = "MS-Cox",
    cox_rmst(c4.cox.iptw.ipcw, TAU_GRID, reference.row)
  )
  results$rmst.cb <- cbind(
    principle = principle.label, configuration = "C8", method = "Case-base",
    casebase_rmst(c8.cb.iptw.ipcw, TAU_GRID, reference.row)
  )

  c9.time.varying.coef <- coef(c9.cox.time.varying)
  c9.time.varying.vcov <- vcov(c9.cox.time.varying)
  
  log.t <- log(pmax(TAU_GRID, 1e-6))
  results$hr.time <- rbind(
    data.frame(
      principle = principle.label, method = "MS-Cox (C9)",
      time = TAU_GRID,
      loghr = c9.time.varying.coef[1] + c9.time.varying.coef[2] * log.t,
      se = sqrt(
        c9.time.varying.vcov[1, 1] +
          log.t^2 * c9.time.varying.vcov[2, 2] +
          2 * log.t * c9.time.varying.vcov[1, 2]
      )
    ),
    local({
      x1 <- casebase_design(c10.cb.time.varying, 1, TAU_GRID, reference.row)
      x0 <- casebase_design(c10.cb.time.varying, 0, TAU_GRID, reference.row)
      design.diff <- x1 - x0
      data.frame(
        principle = principle.label, method = "Case-base (C10)",
        time = TAU_GRID,
        loghr = as.vector(design.diff %*% coef(c10.cb.time.varying$fit)),
        se = sqrt(
          pmax(
            rowSums((design.diff %*% c10.cb.time.varying$vcov) * design.diff), 0
          )
        )
      )
    })
  )
  
  return(results)
}

itt_results <- run_configurations(itt_data, "Intention-to-treat")
pp_results <- run_configurations(pp_data, "Per-protocol")
```

\newpage
# Results (HR)
```{r}
hr_table <- rbind(itt_results$hr, pp_results$hr)
hr_wide <- reshape(
  hr_table[, c("configuration", "model", "principle", "hr.ci")],
  direction = "wide", idvar = c("configuration", "model"),
  timevar = "principle"
)

hr_wide

print(
  xtable::xtable(hr_wide), include.rownames = FALSE, file = "Tables/hr.tex"
)
```

\newpage
# Results (RMST)
```{r}
rmst_all <- rbind(
  itt_results$rmst.cox, itt_results$rmst.cb,
  pp_results$rmst.cox, pp_results$rmst.cb
)
rmst_all$lower <- rmst_all$rmst - Z_CRIT * rmst_all$se
rmst_all$upper <- rmst_all$rmst + Z_CRIT * rmst_all$se

rmst_report <- subset(rmst_all, time %in% TAU_REPORT)
rmst_report$cell <- sprintf(
  "%.2f (%.2f, %.2f)",
  rmst_report$rmst,
  rmst_report$lower,
  rmst_report$upper
)
rmst_wide <- reshape(
  rmst_report[, c("principle", "method", "time", "cell")],
  direction = "wide", idvar = c("principle", "method"),
  timevar = "time"
)
rmst_wide

print(
  xtable::xtable(rmst_wide), include.rownames = FALSE, file = "Tables/rmst.tex"
)
```

```{r}
ggplot2::ggplot(
  rmst_all, ggplot2::aes(
    x = time, y = rmst, colour = method, fill = method, linetype = method
  )
) +
  ggplot2::geom_hline(yintercept = 0, linetype = "dashed", colour = "grey40") +
  ggplot2::geom_ribbon(
    ggplot2::aes(ymin = lower, ymax = upper), alpha = 0.15, colour = NA
  ) +
  ggplot2::geom_line(linewidth = 0.6) +
  ggplot2::facet_wrap(~ principle) +
  ggplot2::scale_colour_manual(values = palette[1:2]) +
  ggplot2::scale_fill_manual(values = palette[1:2]) +
  ggplot2::scale_linetype_manual(values = line_types[1:2]) +
  ggplot2::labs(x = "Restriction horizon (days)", y = "RMST contrast (days)") +
  plot_theme() +
  legend

ggplot2::ggsave(
  "Figures//rmst-contrast.pdf", width = 9, height = 4.4
)

survival_long <- do.call(rbind, lapply(
  list(
    itt_results$rmst.cox, itt_results$rmst.cb,
    pp_results$rmst.cox, pp_results$rmst.cb
  ),
  function(rmst) {
    return(rbind(
        data.frame(
          principle = rmst$principle, method = rmst$method, time = rmst$time,
          arm = "No RHC", surv = rmst$s0
        ),
        data.frame(
          principle = rmst$principle, method = rmst$method, time = rmst$time,
          arm = "RHC", surv = rmst$s1
        )
    ))
  }
))

survival_long$series <- paste(
  survival_long$arm, survival_long$method, sep = ", "
)

ggplot2::ggplot(
  survival_long, ggplot2::aes(
    x = time, y = surv, colour = arm, linetype = method
  )
) +
  ggplot2::geom_line(linewidth = 0.6) +
  ggplot2::facet_wrap(~ principle) +
  ggplot2::scale_colour_manual(values = palette[1:2]) +
  ggplot2::scale_linetype_manual(values = line_types[1:2]) +
  ggplot2::scale_y_continuous(limits = c(0, 1)) +
  ggplot2::labs(
    x = "Days since admission",
    y = "Counterfactual survival probability"
  ) +
  plot_theme() +
  ggplot2::guides(
    colour = ggplot2::guide_legend(nrow = 1, order = 1),
    linetype = ggplot2::guide_legend(nrow = 1, order = 2)
  )

ggplot2::ggsave(
  "Figures//counterfactual-survival.pdf", width = 9, height = 4.4
)

hr_time_all <- rbind(itt_results$hr.time, pp_results$hr.time)
hr_time_all$hr <- exp(hr_time_all$loghr)
hr_time_all$lower <- exp(hr_time_all$loghr - Z_CRIT * hr_time_all$se)
hr_time_all$upper <- exp(hr_time_all$loghr + Z_CRIT * hr_time_all$se)

ggplot2::ggplot(
  subset(hr_time_all, time >= 1), ggplot2::aes(
    x = time, y = hr, colour = method, fill = method, linetype = method
  )
) +
  ggplot2::geom_hline(yintercept = 1, linetype = "dashed", colour = "grey40") +
  ggplot2::geom_ribbon(
    ggplot2::aes(ymin = lower, ymax = upper), alpha = 0.15, colour = NA
  ) +
  ggplot2::geom_line(linewidth = 0.6) +
  ggplot2::facet_wrap(~ principle) +
  ggplot2::scale_colour_manual(values = palette[1:2]) +
  ggplot2::scale_fill_manual(values = palette[1:2]) +
  ggplot2::scale_linetype_manual(values = line_types[1:2]) +
  ggplot2::coord_cartesian(ylim = c(0.5, 2.5)) +
  ggplot2::labs(x = "Days since admission", y = "Marginal hazard ratio") +
  plot_theme() +
  legend

ggplot2::ggsave(
  "Figures//hr-time-varying.pdf", width = 9, height = 4.4
)
```

\newpage
# Sensitivity analyses
The primary analysis truncates the stabilized weights at the first and
ninety-ninth percentiles. The two sensitivity analyses below refit the weighted
configurations without truncation and with the propensity score model restricted
to the demographic and comorbidity blocks, so that the influence of the
physiological measurements on the estimated contrast can be assessed.

```{r}
rmst_ci <- function(est, se, row.idx) {
  sprintf(
    "%.2f (%.2f, %.2f)",
    est[row.idx],
    est[row.idx] - Z_CRIT * se[row.idx],
    est[row.idx] + Z_CRIT * se[row.idx]
  )
}

TAU_IDX <- match(180, TAU_GRID)

sensitivity_fit <- function(
    data.set, weights, knot.probs = c(0.25, 0.5, 0.75),
    ratio = CB_RATIO, include.cox = TRUE, seed = SEED  
  ) {
  
  reference.row <- data.set[1, , drop = FALSE]
  results <- list()
  if (include.cox) {
    cox.fit <- survival::coxph(
      Surv(time, status) ~ A, data = data.set, weights = weights,
                     robust = TRUE
    )
    
    cox.rmst.result <- cox_rmst(cox.fit, TAU_GRID, reference.row)
    results$cox.hr <- hr_ci(coef(cox.fit)["A"], sqrt(diag(vcov(cox.fit))["A"]))
    results$cox.rmst <- rmst_ci(
      cox.rmst.result$rmst, cox.rmst.result$se, TAU_IDX
    )
  
  } else {
    results$cox.hr <- "n/a"
    results$cox.rmst <- "n/a"
  }
  
  cb.object <- fit_casebase(
    data.set, weights, adjust = "none", ratio = ratio,
    knot.probs = knot.probs, seed = seed
  )
  cb.rmst.result <- casebase_rmst(cb.object, TAU_GRID, reference.row)
  results$cb.hr <- hr_ci(
    coef(cb.object$fit)["A"], sqrt(diag(cb.object$vcov)["A"])
  )
  results$cb.rmst <- rmst_ci(cb.rmst.result$rmst, cb.rmst.result$se, TAU_IDX)
  as.data.frame(results, stringsAsFactors = FALSE)
}

untruncated <- iptw * itt_ipcw

reduced_names <- intersect(
  covariate_names, c("age", "sex", "race", "edu", "income", "ninsclas", hx_vars)
)
p_score_reduced <- glm(
  as.formula(paste("A ~", paste(reduced_names, collapse = " + "))),
  data = data_set, family = binomial
)
p_reduced <- pmin(pmax(fitted(p_score_reduced), 1e-6), 1 - 1e-6)
w_reduced <- truncate_weights(ifelse(
  data_set$A == 1,
  p_marginal / p_reduced,
  (1 - p_marginal) / (1 - p_reduced)
))

sensitivity_table <- rbind(
  cbind(
    analysis = "Primary analysis",
    sensitivity_fit(itt_data, itt_data$w.both)
  ),
  cbind(
    analysis = "No weight truncation",
    sensitivity_fit(itt_data, untruncated)
  ),
  cbind(
    analysis = "Reduced propensity score model",
    sensitivity_fit(itt_data, w_reduced)
  ),
  cbind(
    analysis = "Two interior spline knots", sensitivity_fit(
      itt_data, itt_data$w.both,
      knot.probs = c(1/3, 2/3), include.cox = FALSE
    )
  ),
  cbind(
    analysis = "Five interior spline knots", sensitivity_fit(
      itt_data, itt_data$w.both,
      knot.probs = seq(1, 5) / 6, include.cox = FALSE
    )
  ),
  cbind(
    analysis = "Sampling ratio 10", sensitivity_fit(
      itt_data, itt_data$w.both, ratio = 10,
      include.cox = FALSE
    )
  ),
  cbind(
    analysis = "Sampling ratio 100", sensitivity_fit(
      itt_data, itt_data$w.both, ratio = 100,
      include.cox = FALSE
    )
  )
)

sensitivity_table

print(
  xtable::xtable(sensitivity_table), include.rownames = FALSE,
  file = "Tables/sensitivity.tex"
)
```

\newpage
# Case-base resampling variability
The base series is drawn at random, so two fits with identical specifications
differ by the realization of that draw. The sandwich variance conditions on the
realized sample and does not include this component. The analysis below refits
configuration C8 under independent draws and compares the resulting spread with
the reported standard error, at two sampling ratios.

```{r}
N_DRAWS <- 200

resample_spread <- function(data.set, weights, ratio, n.draws = N_DRAWS) {
  reference.row <- data.set[1, , drop = FALSE]
  results <- t(
    vapply(
      seq_len(n.draws),
      function(curr.draw) {
        model.fit <- fit_casebase(
          data.set, weights, adjust = "none", ratio = ratio,
          seed = SEED + curr.draw
        )
        rmst.result <- casebase_rmst(model.fit, TAU_GRID, reference.row)
        
        return(c(
          loghr = unname(coef(model.fit$fit)["A"]),
          se.loghr = unname(sqrt(diag(model.fit$vcov)["A"])),
          rmst = rmst.result$rmst[TAU_IDX],
          se.rmst = rmst.result$se[TAU_IDX]
        ))
      }, numeric(4)
    )
  )
  
  return(data.frame(
    ratio = ratio,
    mean.loghr = mean(results[, "loghr"]),
    sd.loghr = sd(results[, "loghr"]),
    ase.loghr = mean(results[, "se.loghr"]),
    share.loghr = 100 * (sd(results[, "loghr"]) / mean(results[, "se.loghr"]))^2,
    mean.rmst = mean(results[, "rmst"]),
    sd.rmst = sd(results[, "rmst"]),
    ase.rmst = mean(results[, "se.rmst"]),
    share.rmst = 100 * (sd(results[, "rmst"]) / mean(results[, "se.rmst"]))^2
  ))
}

resampling_table <- do.call(rbind, lapply(c(10, 20, 100), function(curr.ratio) {
  cat(sprintf("resampling spread at ratio %d\n", curr.ratio))
  flush.console()
  return(resample_spread(itt_data, itt_data$w.both, ratio = curr.ratio))
}))

resampling_table

print(
  xtable::xtable(resampling_table, digits = c(0, 0, 4, 4, 4, 1, 3, 3, 3, 1)),
  include.rownames = FALSE, file = "Tables/resampling.tex"
)
```

\newpage
# Session information
```{r}
sessionInfo()
```
