---
title: "Causal Survival - RHC Application"
output: pdf_document
date: "2026-05-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newpage
# Package Imports

### EDA
```{r, warning=FALSE, message=FALSE}
library("Hmisc")
library("tableone")
library("xtable")
library("ggplot2")
library("survey")
library("cobalt")
```

### Estimation
```{r, warning=FALSE, message=FALSE}
library("survival")
library("casebase")
library("sandwich")
library("foreach")
library("doParallel")
```

\newpage
# Data Preparation
```{r}
## Set random seed for reproducibility
s = 123

## Load the Right Heart Catheterization (RHC) dataset
getHdata(rhc)
```

# Primary Analysis Variables
```{r, cache=TRUE}
## Define follow-up time as the time from hospital admission to the last
## available contact or death date
time <- as.numeric(
  pmax(rhc$dthdte, rhc$lstctdte, na.rm = TRUE) - rhc$sadmdte
)

## Define the binary event indicator:
## 1 = death, 0 = censored
Delta <- ifelse(
  rhc$death == "Yes",
  1,
  0
)

## Define the binary treatment indicator:
## 1 = right heart catheterization (RHC) performed, 0 = not performed
A <- ifelse(
  rhc$swang1 == "RHC",
  1,
  0
)
```

\newpage
# Baseline Covariates
```{r}
### Demographic and Socioeconomic Characteristics

## Patient sex
Gender <- rhc$sex

## Patient age at baseline
Age <- as.numeric(rhc$age)

## Patient weight in kilograms
## Note: The original dataset encodes some missing values as 0. These values
## are retained because replacing them with NA substantially alters the
## estimated propensity scores and covariate balance following weighting.
Weight <- as.numeric(rhc$wtkilo1)

## Patient race
Race <- droplevels(rhc$race)

## Educational attainment
Education <- as.numeric(rhc$edu)

## Income category
Income <- rhc$income

## Insurance classification
InsuranceClass <- rhc$ninsclas


### Pre-existing Comorbidities

## Names of baseline comorbidity variables to be recoded
hx.vars <- c(
  "chfhx",
  "cardiohx",
  "dementhx",
  "psychhx",
  "chrpulhx",
  "renalhx",
  "liverhx",
  "gibledhx",
  "malighx",
  "immunhx",
  "transhx",
  "amihx"
)

## Convert labelled comorbidity variables to binary factors:
## "No" = absence of comorbidity, "Yes" = presence of comorbidity
rhc[hx.vars] <- lapply(rhc[hx.vars], function(x) {
  x <- as.numeric(x)
  
  factor(
    ifelse(x == 1, "Yes", "No"),
    levels = c("No", "Yes")
  )
})

## Extract recoded comorbidity variables for the analysis dataset
CHFhx <- rhc$chfhx
CARDIOhx <- rhc$cardiohx
DEMENThx <- rhc$dementhx
PSYCHhx <- rhc$psychhx
CHRPULhx <- rhc$chrpulhx
RENALhx <- rhc$renalhx
LIVERhx <- rhc$liverhx
GIBLEDhx <- rhc$gibledhx
MALIGhx <- rhc$malighx
IMMUNhx <- rhc$immunhx
TRANShx <- rhc$transhx
AMIhx <- rhc$amihx


### Disease Severity and Clinical Characteristics

## Indicator of cancer diagnosis
Cancer <- rhc$ca

## Primary disease category
Category <- droplevels(rhc$cat1)

## Baseline disease severity score
Severity <- as.numeric(rhc$das2d3pc)
```

\newpage
# Construct Analysis Dataset
```{r}
## Combine the outcome, treatment, and baseline covariates into a single
## analysis-ready dataset
data.set <- data.frame(
  time = time,
  Delta = Delta,
  A = A,
  
  Gender = Gender,
  Age = Age,
  Weight = Weight,
  Race = Race,
  Education = Education,
  Income = Income,
  InsuranceClass = InsuranceClass,
  
  CHFhx = CHFhx,
  CARDIOhx = CARDIOhx,
  DEMENThx = DEMENThx,
  PSYCHhx = PSYCHhx,
  CHRPULhx = CHRPULhx,
  RENALhx = RENALhx,
  LIVERhx = LIVERhx,
  GIBLEDhx = GIBLEDhx,
  MALIGhx = MALIGhx,
  IMMUNhx = IMMUNhx,
  TRANShx = TRANShx,
  AMIhx = AMIhx,
  
  Cancer = Cancer,
  Category = Category,
  Severity = Severity
)
```

# Plotting Parameters
```{r}
## Define colours used to distinguish the marginal structural Cox model
## and the case-base model in survival and hazard ratio plots
col.cox <- "red"
col.cb <- "blue"

## Define semi-transparent fill colours for confidence interval bands
fill.cox <- rgb(1,0,0,0.20)
fill.cb  <- rgb(0,0,1,0.20)

## Define semi-transparent colours for the exposed and control treatment groups
col.exposed <- rgb(0,0,1,0.5)
col.control <- rgb(1,0,0,0.5)
```

\newpage
# Propensity Score Estimation
```{r}
## Set seed to ensure reproducibility of any downstream procedures involving
## randomization or resampling
set.seed(s)

## Estimate the propensity score, defined as the probability of receiving
## treatment conditional on the observed baseline covariates
propensity.score <- fitted(
  glm(
    A ~ Gender +
      Age +
      Weight +
      Race +
      Education +
      Income +
      InsuranceClass +
      CHFhx +
      CARDIOhx +
      DEMENThx +
      PSYCHhx +
      CHRPULhx +
      RENALhx +
      LIVERhx +
      GIBLEDhx +
      MALIGhx +
      IMMUNhx +
      TRANShx +
      AMIhx + 
      Cancer + 
      Category + 
      Severity,
    data = data.set,
    family = binomial(link = "logit")
  )
)

## Separate estimated propensity scores by treatment group to assess the
## degree of overlap in treatment assignment probabilities
propensity.score.0 <- propensity.score[A == 0]
propensity.score.1 <- propensity.score[A == 1]
```

# Propensity Score Overlap
```{r}
## Plot the propensity score distributions for treated and untreated patients
## to visually assess the positivity/overlap condition
par(
  mar=c(2,2,2,0)
)

hist(
  propensity.score.0,
  col=col.control,
  breaks=seq(0,1,by=0.02),
  ylim=c(0,200),
  main="Propensity Score Overlap",
  xlab="Propensity Scores"
)

hist(
  propensity.score.1,
  col=col.exposed,
  breaks=seq(0,1,by=0.02),
  add=T
)

box()

legend(
  "topright",
  legend=c('A = 0', 'A = 1'),
  fill=c(col.control,col.exposed),
  border="black",
  bty="n"
)
```

\newpage
# Stabilized Inverse Probability Weights
```{r}
## Estimate the marginal probability of treatment, used to construct the
## numerator of the stabilized treatment weights
p.A1 <- mean(A)

## Calculate unstabilized inverse probability of treatment weights (IPWs)
## by taking the inverse of the estimated probability of the treatment
## actually received
ipw.unstabilized <- ifelse(
  A == 1,
  1 / propensity.score,
  1 / (1 - propensity.score)
)

## Calculate stabilized IPWs by incorporating the marginal probability of
## treatment in the numerator. Stabilization reduces the variability of the
## weights while preserving their role in creating a weighted pseudo-population
ipw.stabilized <- ifelse(
  A == 1,
  p.A1 / propensity.score,
  (1 - p.A1) / (1 - propensity.score)
)
```

# IPW Distribution
```{r}
## Compare the distributions of the unstabilized and stabilized IPWs to
## assess the magnitude and variability of the resulting weights
par(
  mar=c(2,2,2,0)
)

hist(
  ipw.unstabilized,
  col=col.control,
  breaks=seq(0, max(ipw.unstabilized) + 0.5,by=0.1),
  ylim=c(0,1000),
  xlim=c(0, 10),
  main="Unstablized vs. Stabilized IPWs",
  ylab="Frequency",
  xlab="IPWs"
)

hist(
  ipw.stabilized,
  col=col.exposed,
  breaks=seq(0, max(ipw.stabilized) + 0.5,by=0.1),
  add=T
)

box()

legend(
  "topright",
  legend=c('Unstabilized', 'Stabilized'),
  fill=c(col.control,col.exposed),
  border="black", bty="n"
)
```

\newpage
# Descriptive Statistics: Table One
```{r}
## Specify the baseline covariates to be included in the descriptive table
covariates = c(
  "Gender",
  "Age",
  "Weight",
  "Race",
  "Education",
  "Income",
  "InsuranceClass",
  
  "CHFhx",
  "CARDIOhx",
  "DEMENThx",
  "PSYCHhx",
  "CHRPULhx",
  "RENALhx",
  "LIVERhx",
  "GIBLEDhx",
  "MALIGhx",
  "IMMUNhx",
  "TRANShx",
  "AMIhx",
  
  "Cancer",
  "Category",
  "Severity"
)

## Generate the unweighted baseline characteristics table, stratified by
## treatment group. Statistical significance tests are omitted because the
## primary objective is descriptive comparison of baseline characteristics.
table.one.original <- CreateTableOne(
  vars = covariates,
  strata = "A",
  data = data.set,
  test = FALSE
)

## Display standardized mean differences (SMDs) to quantify baseline
## differences between the treatment groups
print(table.one.original, smd = TRUE)
```

\newpage
# Weighted Descriptive Statistics: Table Two
```{r, warning=FALSE, message=FALSE}
## Construct a survey design object using the stabilized IPWs. Setting ids = ~0
## specifies that observations are treated as independent sampling units.
data.set.ipw <- svydesign(
  ids = ~ 0,
  data = data.set,
  weights = ipw.stabilized
)

## Generate the weighted baseline characteristics table using the IPW-defined
## pseudo-population, stratified by treatment group
table.one.ipw <- svyCreateTableOne(
  vars = covariates,
  strata = "A",
  data = data.set.ipw,
  test = FALSE
)

## Display standardized mean differences for the weighted covariate
## distributions to assess residual imbalance after IPW
print(table.one.ipw, smd = TRUE)
```

\newpage
# Covariate Balance Before and After IPW
```{r, warning=FALSE, message=FALSE}
## Assess covariate balance before and after application of the stabilized
## inverse probability weights. The unweighted and weighted SMDs are retained
## to facilitate direct comparison of balance across treatment groups.
balance <- bal.tab(
  A ~ Gender +
    Age +
    Weight +
    Race +
    Education +
    Income +
    InsuranceClass +
    CHFhx +
    CARDIOhx +
    DEMENThx +
    PSYCHhx +
    CHRPULhx +
    RENALhx +
    LIVERhx +
    GIBLEDhx +
    MALIGhx +
    IMMUNhx +
    TRANShx +
    AMIhx + 
    Cancer + 
    Category + 
    Severity,
  data = data.set,
  weights = ipw.stabilized,
  method = "weighting",
  estimand = "Average Treatment Effect",
  un = TRUE
)

## Visualize the standardized mean differences before and after IPW.
## An absolute SMD below 0.1 is used as the criterion for acceptable
## covariate balance.
love.plot(
  balance,
  stats = "mean.diffs",
  threshold = 0.1,
  abs = TRUE,
  line = TRUE,
  var.order = "unadjusted",
  colors = c(col.control, col.exposed),
  shapes = c(17, 16),
  limits = c(m = c(0, 0.3)),
  title = "Covariate Balance Before and After IPW",
  labels = TRUE
)

## Covariate balance improved substantially following application of the
## stabilized inverse probability weights, indicating improved comparability
## between treatment groups in the weighted pseudo-population.
```

\newpage
# Estimation

## Assumptions
*Non-Informative Censoring*

# Modified Case-Base Smooth Hazard Model
```{r}
## Modified implementation of fitSmoothHazard() from the casebase package.
## The function extends the standard case-base fitting procedure to allow
## observation-level weights, including inverse probability weights (IPWs).
FSH <- function(
    formula,
    data,
    time,
    w = rep(1, nrow(data)),
    ratio = 100, ...  
) {
  
  ## Preserve the original function call for reproducibility and method
  ## dispatch in downstream model summaries
  cl <- match.call()
  
  ## Explicitly expand "." in the model formula so that all covariates in the
  ## supplied dataset are represented in the model specification
  formula <- expand.dot.formula(
    formula,
    data = data
  )
  
  ## Extract the event variable from the left-hand side of the model formula
  eventVar <- all.vars(formula[[2]])
  
  ## If the follow-up time variable is not explicitly supplied, infer it from
  ## the data using the event variable and the standard case-base structure
  if (missing(time)) {
    varNames <- checkArgsTimeEvent(
      data = data,
      event = eventVar
    )
    
    timeVar <- varNames$time
    
  } else {
    ## Otherwise, use the time variable explicitly supplied by the user
    timeVar <- time
  }
  
  ## Identify the distinct event types present in the data to determine whether
  ## the analysis involves a single event or competing risks
  typeEvents <- sort(unique(data[[eventVar]]))
  
  
  # Case-Base Sampling
  
  ## Generate the case-base sample when the input data have not already been
  ## sampled. The original dataset is retained so that the supplied observation
  ## weights can be incorporated into the case-base sampling procedure.
  if (!inherits(data, "cbData")) {
    
    originalData <- as.data.frame(data)
    originalData$w <- w
    
    ## Sample cases and controls from the person-time data. The sampling ratio
    ## controls the number of person-moments sampled relative to observed cases.
    sampleData <- sampleCaseBase(
      originalData,
      timeVar,
      eventVar,
      comprisk = (length(typeEvents) > 2),
      ratio
    )
    
  } else {
    ## If the input is already a case-base dataset, retain it directly.
    ## The original unsampled data are unavailable at this stage.
    originalData <- data.frame()
    sampleData <- data
  }
  
  
  # Weighted Case-Base Model
  
  ## Include the case-base sampling offset in the model formula. This offset
  ## accounts for the sampling mechanism when estimating the hazard function.
  formula <- update(formula, ~ . + offset(offset))
  
  ## Fit a weighted binomial regression model to the case-base sample.
  ## Observation-level weights allow the model to incorporate IPWs in addition
  ## to the case-base sampling structure.
  fittingFunction <- function(formula) glm(
    formula,
    data = sampleData,
    family = binomial,
    weights = w
  )
  
  ## Estimate the smooth hazard model using the weighted case-base sample
  out <- fittingFunction(formula)
  
  
  # Model Object and Metadata
  
  ## Preserve the lower-level glm call while replacing the displayed call with
  ## the original FSH call for consistency with the custom model interface
  out$lower_call <- out$call 
  out$call <- cl
  
  ## Store the original data and key variable names for downstream methods
  ## such as prediction and plotting
  out$originalData <- originalData
  out$typeEvents <- typeEvents
  out$timeVar <- timeVar
  out$eventVar <- eventVar
  
  ## Record the number of case and background person-moments in the sampled
  ## case-base dataset
  num_pm <- table(sampleData[[eventVar]])
  out$num_cm <- num_pm[2]
  out$num_bm <- num_pm[1]
  out$formula <- formula
  
  ## Store the case-base sampling offset separately, then reset the model's
  ## offset to zero so that absolute risk estimation can be performed without
  ## retaining the sampling correction in the prediction stage
  out$offset <- out$data$offset
  out$data$offset <- 0
  
  ## Assign a custom class so that methods defined for single-event case-base
  ## models can be applied to the fitted object
  class(out) <- c("singleEventCB", class(out))
  
  return(out)
}
```

# Formula Expansion Helper
```{r}
## Expand "." in a model formula into the complete set of covariates available
## in the supplied dataset. This ensures that the modified FSH function can
## explicitly evaluate formulas containing "." before fitting the model.
expand.dot.formula <- function(formula, data = NULL) {
  
  ## Check whether the formula contains the "." shorthand for all available
  ## predictors
  if (isTRUE("." %in% all.vars(formula))) {
    
    ## Preserve the original formula attributes while expanding the terms
    att <- attributes(formula)
    
    try.terms <- try(
      stats::terms(formula, data = data),
      silent = TRUE
    )
    
    ## Replace the shorthand formula with its explicitly expanded version when
    ## the terms object can be constructed successfully
    if (!is(try.terms, "try-error")) {
      formula <- formula(try.terms)
    }
    
    ## Restore the original formula attributes after expansion
    attributes(formula) <- att
  }
  
  ## Return the expanded formula
  return(formula)
}
```

\newpage
# Case-Base Model Setup
```{r}
## Set the case-base sampling ratio used across all case-base model fits.
## A ratio of 100 samples 100 background person-moments per observed case.
fsh.ratio <- 100
```

# Counterfactual Survival Curve Setup
```{r}
## Define the two treatment levels used to generate counterfactual survival
## curves: A = 0 represents no RHC and A = 1 represents RHC
trt.types <- c(0, 1)

## Create a single-treatment profile for prediction under the treated condition
individual.profile.data <- data.frame(A = 1)

## Create a prediction dataset containing both treatment levels for estimation
## of counterfactual survival under each treatment condition
survival.curves.data <- data.frame(A = trt.types)

# Time Grid Definition

## Determine the maximum observed follow-up time in the analysis dataset
tau.max <- max(data.set$time, na.rm = TRUE)

## Restrict the primary survival-curve display to the first 150 days
## (approximately four months) to focus on the relevant follow-up period
t.max <- 150

## Define the time grid used for RMST integration, spanning the full observed
## follow-up period at one-day intervals
tau.grid <- seq(0, tau.max, by = 1)

## Define the time grid used for plotting and estimating survival curves over
## a restricted four-month follow-up period
t.grid <- seq(0, t.max, by = 1)
```

\newpage
# Model 1: Unadjusted Marginal Structural Cox Model
```{r, cache=TRUE}
## Set seed to ensure reproducibility of any stochastic procedures used
## during model fitting
set.seed(s)

## Fit an unadjusted Cox proportional hazards model with treatment as the
## sole predictor. The resulting coefficient estimates the marginal
## log-hazard ratio comparing RHC with no RHC.
model.1 <- coxph(
  Surv(time, Delta) ~ A,
  data = data.set,
  robust = TRUE
)

## Extract the treatment coefficient and its standard error
beta.1 <- coef(model.1)["A"]
se.1 <- sqrt(diag(vcov(model.1))["A"])

## Transform the treatment coefficient from the log-hazard ratio scale to the
## hazard ratio scale and calculate the corresponding 95% Wald confidence
## interval
cat("HR-hat:", exp(beta.1), "\n")
cat("SE of HR-hat:", se.1, "\n")
cat("95% CI: (", exp(beta.1 + c(-1,1)*1.96*se.1), ")\n")
```

# Model 2: Unadjusted Case-Base Model
```{r, cache=TRUE, warning=FALSE, message=FALSE}
## Set seed to ensure reproducibility of the case-base sampling procedure
set.seed(s)

## Fit an unadjusted smooth hazard model using case-base sampling, with
## treatment as the sole predictor. The sampling ratio determines the number
## of background person-moments sampled for each observed case.
model.2 <- fitSmoothHazard(
  Delta ~ A,
  data = data.set,
  ratio = fsh.ratio
)

## Extract the treatment coefficient and calculate its robust standard error
## using the heteroskedasticity-consistent covariance estimator
beta.2 <- coef(model.2)["A"]
se.2 <- sqrt(diag(vcovHC(model.2))["A"])

## Transform the treatment coefficient to the hazard ratio scale and calculate
## the corresponding 95% Wald confidence interval
cat("HR-hat:", exp(beta.2), "\n")
cat("SE of HR-hat:", se.2, "\n")
cat("95% CI: (", exp(beta.2 + c(-1,1)*1.96*se.2), ")\n")
```

\newpage
# Model 3: Covariate-Adjusted Cox Model
```{r, cache=TRUE}
## Set seed to ensure reproducibility of any stochastic procedures used
## during model fitting
set.seed(s)

## Fit a covariate-adjusted Cox proportional hazards model. Treatment is
## adjusted for the full set of observed baseline demographic, socioeconomic,
## comorbidity, and disease-severity covariates.
model.3 <- coxph(
  Surv(time, Delta) ~ A +
    Gender +
    Age +
    Weight +
    Race +
    Education +
    Income +
    InsuranceClass +
    CHFhx +
    CARDIOhx +
    DEMENThx +
    PSYCHhx +
    CHRPULhx +
    RENALhx +
    LIVERhx +
    GIBLEDhx +
    MALIGhx +
    IMMUNhx +
    TRANShx +
    AMIhx + 
    Cancer + 
    Category + 
    Severity,
  data = data.set,
  robust = TRUE
)

## Extract the treatment coefficient and its robust standard error
beta.3 <- coef(model.3)["A"]
se.3 <- sqrt(diag(vcov(model.3))["A"])

## Transform the treatment coefficient to the hazard ratio scale and calculate
## the corresponding 95% Wald confidence interval
cat("HR-hat:", exp(beta.3), "\n")
cat("SE of HR-hat:", se.3, "\n")
cat("95% CI: (", exp(beta.3 + c(-1,1)*1.96*se.3), ")\n")
```

\newpage
# Model 4: Covariate-Adjusted Case-Base Model
```{r, cache=TRUE, warning=FALSE, message=FALSE}
## Set seed to ensure reproducibility of the case-base sampling procedure
set.seed(s)

## Fit a covariate-adjusted smooth hazard model using case-base sampling.
## The model uses the same treatment and baseline covariates as the adjusted
## Cox model to provide a directly comparable alternative estimator.
model.4 <- fitSmoothHazard(
  Delta ~ A +
    Gender +
    Age +
    Weight +
    Race +
    Education +
    Income +
    InsuranceClass +
    CHFhx +
    CARDIOhx +
    DEMENThx +
    PSYCHhx +
    CHRPULhx +
    RENALhx +
    LIVERhx +
    GIBLEDhx +
    MALIGhx +
    IMMUNhx +
    TRANShx +
    AMIhx + 
    Cancer + 
    Category + 
    Severity,
  data = data.set,
  ratio = fsh.ratio
)

## Extract the treatment coefficient and calculate its heteroskedasticity-
## consistent standard error
beta.4 <- coef(model.4)["A"]
se.4 <- sqrt(diag(vcovHC(model.4))["A"])

## Transform the treatment coefficient to the hazard ratio scale and calculate
## the corresponding 95% Wald confidence interval
cat("HR-hat:", exp(beta.4), "\n")
cat("SE of HR-hat:", se.4, "\n")
cat("95% CI: (", exp(beta.4 + c(-1,1)*1.96*se.4), ")\n")
```

\newpage
# Model 5: Marginal Structural Cox Model
```{r, cache=TRUE}
## Set seed to ensure reproducibility of any stochastic procedures used
## during model fitting
set.seed(s)

## Fit the marginal structural Cox model using stabilized inverse probability
## weights (IPWs). The treatment-only model estimates the marginal treatment
## effect in the IPW-defined pseudo-population.
model.5 <- coxph(
  Surv(time, Delta) ~ A,
  data = data.set,
  weights = ipw.stabilized,
  robust = TRUE
)

## Extract the treatment coefficient and its robust standard error
beta.5 <- coef(model.5)["A"]
se.5 <- sqrt(diag(vcov(model.5))["A"])

## Transform the treatment coefficient to the hazard ratio scale and calculate
## the corresponding 95% Wald confidence interval
cat("HR-hat:", exp(beta.5), "\n")
cat("SE of HR-hat:", se.5, "\n")
cat("95% CI: (", exp(beta.5 + c(-1,1)*1.96*se.5), ")\n")
```

\newpage
# Model 6: IPW Case-Base Proportional Hazards Model
```{r, cache=TRUE, warning=FALSE, message=FALSE}
## Set seed to ensure reproducibility of the case-base sampling procedure
set.seed(s)

## Fit the modified smooth hazard model using case-base sampling and stabilized
## IPWs. The custom FSH function incorporates the IPWs as observation-level
## weights while accounting for the case-base sampling mechanism.
model.6 <- FSH(
  Delta ~ A,
  data = data.set,
  ratio = fsh.ratio,
  w = ipw.stabilized
)

## Extract the treatment coefficient and calculate its heteroskedasticity-
## consistent standard error
beta.6 <- coef(model.6)["A"]
se.6 <- sqrt(diag(vcovHC(model.6))["A"])

## Transform the treatment coefficient to the hazard ratio scale and calculate
## the corresponding 95% Wald confidence interval
cat("HR-hat:", exp(beta.6), "\n")
cat("SE of HR-hat:", se.6, "\n")
cat("95% CI: (", exp(beta.6 + c(-1,1)*1.96*se.6), ")\n")
```

\newpage
# Survival Curve Prediction Setup
```{r, cache=TRUE}
## Define the two treatment levels used to generate counterfactual survival
## curves under no RHC (A = 0) and RHC (A = 1)
trt.types <- c(0, 1)

## Define the treatment profile used for individual-level prediction
individual.profile.data <- data.frame(A = 1)

## Create a prediction dataset containing both treatment levels for estimation
## of the corresponding counterfactual survival curves
survival.curves.data <- data.frame(A = trt.types)

## Determine the maximum observed follow-up time and define the restricted
## four-month follow-up period used for focused visualization
tau.max <- max(data.set$time, na.rm = TRUE)
t.max <- 150

## Define the full follow-up grid and the restricted four-month grid
tau.grid <- seq(0, tau.max, by = 1)
t.grid <- seq(0, t.max, by = 1)
```

\newpage
# Model 7: Unadjusted Case-Base Model with Time-Varying Hazard Ratio
```{r, cache=TRUE, warning=FALSE, message=FALSE}
## Set seed to ensure reproducibility of the case-base sampling procedure
set.seed(s)

## Fit an unadjusted case-base smooth hazard model allowing the treatment
## effect to vary over time. The spline term models the baseline time effect,
## while its interaction with treatment permits a non-constant hazard ratio.
model.7 <- FSH(
  Delta ~ pspline(time, df = 2) * A,
  data = data.set,
  ratio = fsh.ratio
)
```

# Time-Varying Hazard Ratio: Full Follow-Up
```{r, message=FALSE}
## Display the estimated treatment hazard ratio as a function of follow-up
## time, including its 95% confidence interval
plot(
  model.7,
  type = "hr",
  newdata = expand.grid(
    time = tau.grid,
    A = trt.types
  ),
  var = "A",
  xvar = "time",
  ci = TRUE,
  ylim = c(0,2),
  xlab="Time",
  ylab="Hazard Ratio",
  main="Model 7: Variable Hazard Ratio"
)

## Add a reference line corresponding to a hazard ratio of 1, representing
## no instantaneous treatment effect
abline(h = 1, col = "red", lty = 2)
```

# Time-Varying Hazard Ratio: First Four Months
```{r}
## Restrict the hazard-ratio plot to the first 150 days (approximately four
## months) to provide a more detailed view of the early follow-up period
plot(
  model.7,
  type = "hr",
  newdata = expand.grid(
    time = t.grid,
    A = trt.types
  ),
  var = "A",
  xvar = "time",
  ci = TRUE,
  ylim = c(1,1.5),
  xlab="Time",
  ylab="Hazard Ratio",
  main="Model 7: Variable Hazard Ratio - First 4 Months"
)

## Add the no-effect reference line
abline(h = 1, col = "red", lty = 2)
```

\newpage
# Model 8: IPW Case-Base Model with Time-Varying Hazard Ratio
```{r, cache=TRUE, warning=FALSE, message=FALSE}
## Set seed to ensure reproducibility of the case-base sampling procedure
set.seed(s)

## Fit the IPW case-base smooth hazard model, allowing the treatment effect
## to vary over time. Stabilized IPWs are incorporated through the modified
## FSH function to estimate the treatment effect in the weighted
## pseudo-population.
model.8 <- FSH(
  Delta ~ pspline(time, df = 2) * A,
  data = data.set,
  ratio = fsh.ratio,
  w = ipw.stabilized
)
```

# Time-Varying Hazard Ratio: Full Follow-Up
```{r, message=FALSE}
## Display the estimated time-varying treatment hazard ratio and its 95%
## confidence interval over the full follow-up period
plot(
  model.8,
  type = "hr",
  newdata = expand.grid(
    time = tau.grid,
    A = trt.types
  ),
  var = "A",
  xvar = "time",
  ci = TRUE,
  ylim = c(0,2),
  xlab="Time",
  ylab="Hazard Ratio",
  main="Model 8: Variable Hazard Ratio"
)

## Add the no-effect reference line
abline(h = 1, col = "red", lty = 2)
```

# Time-Varying Hazard Ratio: First Four Months
```{r}
## Restrict the hazard-ratio plot to the first 150 days to provide a more
## detailed visualization of the early treatment effect
plot(
  model.8,
  type = "hr",
  newdata = expand.grid(
    time = t.grid,
    A = trt.types
  ),
  var = "A",
  xvar = "time",
  ci = TRUE,
  ylim = c(1,1.5),
  xlab="Time",
  ylab="Hazard Ratio",
  main="Model 8: Variable Hazard Ratio - First 4 Months"
)

## Add the no-effect reference line
abline(h = 1, col = "red", lty = 2)
```

# Head-to-Head Comparison: Models 7 and 8
```{r}
## Extract the estimated time-varying treatment hazard ratios and their 95%
## confidence intervals from the unadjusted and IPW-adjusted case-base models.
## The resulting estimates are used to construct a common comparison plot.
get.hr <- function(model) {
  invisible(pdf(NULL))
  
  out <- plot(
    model,
    type = "hr",
    newdata = data.frame(
      time = tau.grid,
      A = 0
    ),
    var = "A",
    xvar = "time",
    ci = TRUE
  )
  
  invisible(dev.off())
  
  out
}

## Extract the time-varying hazard-ratio estimates from Models 7 and 8
hr.7 <- get.hr(model.7)
hr.8 <- get.hr(model.8)

# Combined Hazard-Ratio Plot

## Initialize an empty plotting region using the time and hazard-ratio ranges
## from the extracted estimates. The curves and confidence intervals for both
## models are added to this common plotting region below.
plot(
  hr.7$time,
  hr.7$hazard_ratio,
  type = "n",
  ylim = c(0, 2),
  xlab = "Time",
  ylab = "Hazard Ratio",
  main = "Head-to-head Comparison of Variable HR Models"
)

# 95% Confidence Intervals

## Add the 95% confidence interval band for the unadjusted case-base model
polygon(
  c(hr.7$time, rev(hr.7$time)),
  c(hr.7$lowerbound, rev(hr.7$upperbound)),
  col = adjustcolor("red", alpha.f = 0.20),
  border = NA
)

## Add the 95% confidence interval band for the IPW-adjusted case-base model
polygon(
  c(hr.8$time, rev(hr.8$time)),
  c(hr.8$lowerbound, rev(hr.8$upperbound)),
  col = adjustcolor("blue", alpha.f = 0.20),
  border = NA
)

# Estimated Hazard-Ratio Curves

## Overlay the estimated time-varying hazard-ratio curve from the unadjusted
## case-base model
lines(
  hr.7$time,
  hr.7$hazard_ratio,
  col = "red",
  lwd = 2
)

## Overlay the estimated time-varying hazard-ratio curve from the IPW-adjusted
## case-base model
lines(
  hr.8$time,
  hr.8$hazard_ratio,
  col = "blue",
  lwd = 2
)

# Reference Line and Legend

## Add a horizontal reference line at HR = 1, corresponding to no difference
## in the instantaneous hazard between treatment groups
abline(
  h = 1,
  col = "black",
  lty = 2
)

## Identify the two estimation approaches shown in the combined plot
legend(
  "bottomleft",
  legend = c("Unadjusted", "IPW-Adjusted"),
  col = c("red", "blue"),
  lwd = 2,
  bty = "n"
)
```

\newpage
# Restricted Mean Survival Time (RMST)
```{r}
## Numerically integrate a survival curve over the specified follow-up period
## using the trapezoidal rule. The cumulative integral at each time point
## represents the RMST up to that point.
rmst.fn <- function(time, surv) {
  dt <- diff(time)
  mids <- (head(surv, -1) + tail(surv, -1)) / 2
  c(0, cumsum(mids * dt))
}
```

\newpage
# MS-Cox RMST and Standard Error
```{r}
## Calculate the counterfactual RMST contrast and its standard error from the
## marginal structural Cox model. The RMST contrast is defined as the
## difference in restricted mean survival between RHC and no RHC.
cox.rmst.list <- function(
    ms.cox, t.grid, cox.psi.se, survival.curves.data 
) {
  
  # Counterfactual Survival Curves
  
  ## Estimate the counterfactual survival curves under both treatment levels
  ## from the fitted marginal structural Cox model
  sf <- survfit(
    ms.cox,
    newdata = survival.curves.data
  )
  
  ## Interpolate the counterfactual survival curve under no RHC onto the
  ## specified time grid
  S0 <- approx(
    sf$time,
    sf$surv[, 1],
    t.grid,
    method = "constant",
    rule = 2
  )$y
  
  ## Interpolate the counterfactual survival curve under RHC onto the
  ## specified time grid
  S1 <- approx(
    sf$time,
    sf$surv[, 2],
    t.grid,
    method = "constant",
    rule = 2
  )$y
  
  # RMST Contrast
  
  ## Calculate the RMST contrast as the difference between the restricted mean
  ## survival under RHC and no RHC
  cox.rmst <- rmst.fn(t.grid, S1) - rmst.fn(t.grid, S0)
  
  # Delta-Method Standard Error
  
  ## Construct the gradient of the RMST functional with respect to the
  ## treatment log-hazard coefficient. For the Cox model, the derivative
  ## involves the integral of S1(t) log{S1(t)} over the restriction period.
  integrand <- ifelse(S1 > 0, S1 * log(S1), 0)
  g.cox <- rmst.fn(t.grid, integrand)
  
  ## Propagate uncertainty from the estimated Cox treatment coefficient to the
  ## RMST contrast using the delta method
  cox.rmst.se <- abs(g.cox) * cox.psi.se
  
  return(list(
    cox.rmst = cox.rmst,
    cox.rmst.se = cox.rmst.se
  ))
}
```

\newpage
# IPW Case-Base RMST and Standard Error
```{r}
## Calculate the RMST contrast and its standard error from the IPW case-base
## model. The survival curves are reconstructed from the fitted model and the
## variance of the RMST contrast is obtained using a numerical delta method.
cb.rmst.list <- function(cb.ipw, cb.var, t.grid, constant.HR) {
  
  ## Extract the estimated model coefficients and the corresponding model
  ## terms used to construct the design matrix
  beta.hat <- coef(cb.ipw)
  term.form <- delete.response(terms(cb.ipw))
  
  # Counterfactual Survival Function
  
  ## Construct the counterfactual survival function under treatment level a
  ## directly from the case-base model's design matrix
  cb.surv <- function(beta, a, grid) {
    
    ## Create prediction data over the specified time grid under treatment
    ## level a
    nd <- data.frame(
      time = grid,
      A = a,
      offset = 0
    )
    
    ## Construct the model matrix using the same terms as the fitted model and
    ## align its columns with the estimated coefficient vector
    X <- model.matrix(term.form, data = nd)
    X <- X[, names(beta), drop = FALSE]
    
    ## Convert the linear predictor to the estimated hazard function
    lambda <- exp(as.vector(X %*% beta))
    
    ## Numerically integrate the estimated hazard using the trapezoidal rule
    ## to obtain the cumulative hazard
    Lambda <- c(
      0,
      cumsum(
        diff(grid) * (head(lambda, -1) + tail(lambda, -1)) / 2
      )
    )
    
    ## Convert cumulative hazard to survival probability
    exp(-Lambda)
  }
  
  # RMST Contrast
  
  ## Define the RMST contrast as a function of the model coefficient vector.
  ## This function is subsequently used to obtain the numerical gradient
  ## required for the delta-method standard error.
  rmst.diff <- function(beta) {
    rmst.fn(t.grid, cb.surv(beta, 1, t.grid)) -
      rmst.fn(t.grid, cb.surv(beta, 0, t.grid))
  }
  
  ## Evaluate the RMST contrast at the estimated model coefficients
  cb.rmst <- rmst.diff(beta.hat)
  
  # Numerical Gradient
  
  ## Calculate the numerical gradient of the RMST contrast with respect to
  ## each model coefficient using a central finite-difference approximation
  p <- length(beta.hat)
  m <- length(t.grid)
  
  grad <- matrix(NA_real_, nrow = m, ncol = p)
  colnames(grad) <- names(beta.hat)
  
  ## Step size used for the finite-difference approximation
  eps <- 1e-6
  
  ## Perturb each coefficient in turn and approximate the corresponding
  ## derivative of the RMST contrast
  for (j in seq_len(p)) {
    bp <- bm <- beta.hat
    bp[j] <- bp[j] + eps
    bm[j] <- bm[j] - eps
    
    grad[, j] <- (rmst.diff(bp) - rmst.diff(bm)) / (2 * eps)
  }
  
  # Delta-Method Standard Error
  
  ## Align the model covariance matrix with the coefficient ordering used by
  ## the numerical gradient before applying the delta-method variance formula
  cb.var <- cb.var[colnames(grad), colnames(grad), drop = FALSE]
  
  ## Calculate the RMST standard error at each restriction time using
  ## Var{g(beta)} = grad' Var(beta) grad
  cb.rmst.se <- vapply(
    seq_len(m),
    function(i) {
      g <- grad[i, ]
      sqrt(drop(t(g) %*% cb.var %*% g))
    },
    numeric(1)
  )
  
  return(list(
    cb.rmst = cb.rmst,
    cb.rmst.se = cb.rmst.se
  ))
}
``` 

\newpage
# Counterfactual Survival Curves: MS-Cox and IPW Case-Base Models

# MS-Cox Counterfactual Survival Curves
```{r, cache=TRUE}
## Estimate counterfactual survival under both treatment levels using the
## marginal structural Cox model. A = 0 corresponds to no RHC, while A = 1
## corresponds to RHC.
sf.cox <- survfit(
  model.5,
  newdata = survival.curves.data
)

## Interpolate the estimated survival curve under no RHC onto the common
## time grid used for subsequent comparison and RMST calculation
S0.cox <- approx(
  sf.cox$time,
  sf.cox$surv[, 1],
  xout = tau.grid,
  method = "constant",
  rule = 2
)$y

## Interpolate the corresponding lower 95% confidence limit under no RHC
S0.lower.cox <- approx(
  sf.cox$time,
  sf.cox$lower[, 1],
  xout = tau.grid,
  method = "constant",
  rule = 2
)$y

## Interpolate the corresponding upper 95% confidence limit under no RHC
S0.upper.cox <- approx(
  sf.cox$time,
  sf.cox$upper[, 1],
  xout = tau.grid,
  method = "constant",
  rule = 2
)$y

## Interpolate the estimated survival curve under RHC onto the common time grid
S1.cox <- approx(
  sf.cox$time,
  sf.cox$surv[, 2],
  xout = tau.grid,
  method = "constant",
  rule = 2
)$y

## Interpolate the corresponding lower 95% confidence limit under RHC
S1.lower.cox <- approx(
  sf.cox$time,
  sf.cox$lower[, 2],
  xout = tau.grid,
  method = "constant",
  rule = 2
)$y

## Interpolate the corresponding upper 95% confidence limit under RHC
S1.upper.cox <- approx(
  sf.cox$time,
  sf.cox$upper[, 2],
  xout = tau.grid,
  method = "constant",
  rule = 2
)$y
```

\newpage
# IPW Case-Base Counterfactual Survival Curves
```{r, cache=TRUE, message=FALSE, warning=FALSE}
## Estimate counterfactual survival under no RHC and RHC using the IPW
## case-base model over the common time grid
Sf.cb <- absoluteRisk(
  object = model.8,
  time = tau.grid,
  newdata = survival.curves.data,
  type = "survival"
)

## Extract the estimated survival curves for no RHC and RHC, respectively
S0.cb <- Sf.cb[, 2]
S1.cb <- Sf.cb[, 3]

# Bootstrap Confidence Intervals for the IPW Case-Base Curves

## Use nonparametric bootstrap resampling to obtain pointwise confidence
## intervals for the counterfactual survival curves estimated by the IPW
## case-base model
set.seed(s)
n.boot <- 200

## For each bootstrap sample, refit the IPW case-base model and obtain
## counterfactual survival estimates under both treatment levels
boot.surv <- replicate(n.boot, {
  idx <- sample(nrow(data.set), replace = TRUE)
  boot.data <- data.set[idx, ]
  
  boot.fit <- FSH(
    Delta ~ pspline(time, df = 2) * A,
    data = boot.data,
    ratio = 10,
    w = ipw.stabilized[idx]
  )
  
  ar <- absoluteRisk(
    boot.fit,
    time = tau.grid,
    newdata = survival.curves.data,
    type = "survival"
  )
  
  list(
    S0 = ar[, 2],
    S1 = ar[, 3]
  )
})

## Organize the bootstrap estimates into separate matrices for the no-RHC
## and RHC counterfactual survival curves
S0.boot <- sapply(1:n.boot, function(i) boot.surv[["S0", i]])
S1.boot <- sapply(1:n.boot, function(i) boot.surv[["S1", i]])

## Calculate the pointwise lower 95% bootstrap confidence limit for the
## no-RHC counterfactual survival curve
S0.lower.cb <- apply(
  S0.boot,
  1,
  quantile,
  probs = 0.025,
  na.rm = TRUE
)

## Calculate the pointwise upper 95% bootstrap confidence limit for the
## no-RHC counterfactual survival curve
S0.upper.cb <- apply(
  S0.boot,
  1,
  quantile,
  probs = 0.975,
  na.rm = TRUE
)

## Calculate the pointwise lower 95% bootstrap confidence limit for the
## RHC counterfactual survival curve
S1.lower.cb <- apply(
  S1.boot,
  1,
  quantile,
  probs = 0.025,
  na.rm = TRUE
)

## Calculate the pointwise upper 95% bootstrap confidence limit for the
## RHC counterfactual survival curve
S1.upper.cb <- apply(
  S1.boot,
  1,
  quantile,
  probs = 0.975,
  na.rm = TRUE
)
```

# Head-to-Head Comparison of Counterfactual Survival Curves
```{r}
## Compare the MS-Cox and IPW case-base estimates separately under each
## treatment condition. The confidence bands are plotted first so that the
## survival curves remain visible on top of the shaded regions.

# Counterfactual Survival Under No RHC

## Initialize the plotting region for the counterfactual survival curves
## under no RHC (A = 0)
plot(
  tau.grid, S0.cox,
  type = "n",
  lwd = 2,
  col = col.cox,
  lty = 1,
  ylim = c(0, 1),
  xlab = "Time",
  ylab = "Survival Probability",
  main = "Counterfactual Survival Curves (No RHC)"
)

## Add the 95% confidence band for the MS-Cox survival estimate
polygon(
  c(tau.grid, rev(tau.grid)),
  c(S0.lower.cox, rev(S0.upper.cox)),
  col = fill.cox,
  border = NA
)

## Add the 95% bootstrap confidence band for the IPW case-base survival
## estimate
polygon(
  c(tau.grid, rev(tau.grid)),
  c(S0.lower.cb, rev(S0.upper.cb)),
  col = fill.cb,
  border = NA
)

## Overlay the estimated survival curves. The MS-Cox curve is a step function,
## whereas the case-base model produces a smooth survival curve.
lines(S0.cox, type = "s", lwd = 2, col = col.cox, lty = 1)
lines(S0.cb, type = "l", lwd = 2, col = col.cb, lty = 1)

## Identify the two estimation approaches in the comparison plot
legend(
  "topright",
  legend = c(
    "Semi-Parametric: MS-Cox",
    "Parametric: Case Base"
  ),
  col = c(
    col.cox, col.cb
  ),
  lty = c(1, 1),
  lwd = 2,
  bty = "n"
)

# Counterfactual Survival Under RHC

## Initialize the plotting region for the counterfactual survival curves
## under RHC (A = 1)
plot(
  tau.grid, S1.cox,
  type = "n",
  lwd = 2,
  col = col.cox,
  lty = 1,
  ylim = c(0, 1),
  xlab = "Time",
  ylab = "Survival Probability",
  main = "Counterfactual Survival Curves (RHC)"
)

## Add the 95% confidence band for the MS-Cox survival estimate
polygon(
  c(tau.grid, rev(tau.grid)),
  c(S1.lower.cox, rev(S1.upper.cox)),
  col = fill.cox,
  border = NA
)

## Add the 95% bootstrap confidence band for the IPW case-base survival
## estimate
polygon(
  c(tau.grid, rev(tau.grid)),
  c(S1.lower.cb, rev(S1.upper.cb)),
  col = fill.cb,
  border = NA
)

## Overlay the estimated survival curves under RHC
lines(S1.cox, type = "s", lwd = 2, col = col.cox, lty = 1)
lines(S1.cb, type = "l", lwd = 2, col = col.cb, lty = 1)

## Identify the two estimation approaches in the comparison plot
legend(
  "topright",
  legend = c(
    "Semi-Parametric: MS-Cox",
    "Parametric: Case Base"
  ),
  col = c(
    col.cox, col.cb
  ),
  lty = c(1, 1),
  lwd = 2,
  bty = "n"
)
```

\newpage
# RMST Comparison: MS-Cox vs. IPW Case-Base

# RMST Estimates and Standard Errors
```{r, cache=TRUE}
## Extract the robust standard error of the treatment log-hazard coefficient
## from the marginal structural Cox model
model.5.psi.se <- sqrt(diag(vcov(model.5))["A"])

## Obtain the heteroskedasticity-consistent covariance matrix for the IPW
## case-base model. This covariance matrix is used in the numerical
## delta-method calculation of the RMST standard error.
model.8.var <- cb.var <- vcovHC(model.8, type = "HC0")


## Calculate the RMST contrast and its delta-method standard error for the
## marginal structural Cox model
cox.rmst.list <- cox.rmst.list(
  model.5, tau.grid, model.5.psi.se, survival.curves.data
)

## Calculate the RMST contrast and its delta-method standard error for the
## IPW case-base model
cb.rmst.list <- cb.rmst.list(
  model.8, model.8.var, tau.grid, constant.HR = TRUE
)

# Extract RMST Estimates and Standard Errors

## Extract the estimated RMST contrasts and their corresponding standard errors
## from both estimation approaches
cox.rmst <- cox.rmst.list$cox.rmst
cox.rmst.se <- cox.rmst.list$cox.rmst.se

cb.rmst <- cb.rmst.list$cb.rmst
cb.rmst.se <- cb.rmst.list$cb.rmst.se

# Pointwise 95% Confidence Intervals

## Construct pointwise 95% Wald confidence intervals for the MS-Cox RMST
## contrast over the range of restriction times
cox.lower <- cox.rmst - 1.96 * cox.rmst.se
cox.upper <- cox.rmst + 1.96 * cox.rmst.se

## Construct pointwise 95% Wald confidence intervals for the IPW case-base
## RMST contrast over the range of restriction times
cb.lower <- cb.rmst - 1.96 * cb.rmst.se
cb.upper <- cb.rmst + 1.96 * cb.rmst.se
```

# RMST Comparison Plot
```{r}
## Determine a common y-axis range that accommodates the point estimates and
## confidence intervals from both estimation approaches
ylim <- range(
  c(
    cox.lower, cox.upper,
    cb.lower, cb.upper
  ),
  na.rm = TRUE
)

## Initialize the plotting region for comparing the estimated causal RMST
## contrasts over the full range of restriction times
plot(
  tau.grid, cox.rmst,
  type = "n",
  ylim = ylim,
  xlab = expression(tau),
  ylab = "RMST Difference",
  main = "Causal RMST up to Tau"
)

# 95% Confidence Bands

## Add the pointwise 95% confidence band for the MS-Cox RMST contrast
polygon(
  c(tau.grid, rev(tau.grid)),
  c(cox.lower, rev(cox.upper)),
  col = adjustcolor(col.cox, alpha.f = 0.2),
  border = NA
)

## Add the pointwise 95% confidence band for the IPW case-base RMST contrast
polygon(
  c(tau.grid, rev(tau.grid)),
  c(cb.lower, rev(cb.upper)),
  col = adjustcolor(col.cb, alpha.f = 0.2),
  border = NA
)

# RMST Contrast Estimates

## Overlay the estimated causal RMST contrast from the MS-Cox model
lines(tau.grid, cox.rmst, col = col.cox, lwd = 2)

## Overlay the estimated causal RMST contrast from the IPW case-base model
lines(tau.grid, cb.rmst, col = col.cb, lwd = 2)

## Add a reference line at an RMST difference of zero, corresponding to equal
## restricted mean survival under the two treatment conditions
abline(h = 0, lty = 3)

# Model Legend

## Identify the two estimation approaches in the comparison plot
legend(
  "bottomleft",
  legend = c("MS-Cox", "IPW Case-Base"),
  col = c(col.cox, col.cb),
  lwd = 2,
  bty = "n"
)
```