---
title: "Causal Survival - RHC Application - IPCW"
output: pdf_document
date: "2026-05-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newpage
# Design by Contract
```{r}
stop.running <- function(condition, message) {
  
  this.function.name <- rlang::call_name(rlang::current_call())
  
  if (is.null(condition)) {
    stop(paste0(this.function.name, ": condition must not be NULL"))
  }
  if (is.null(message)) {
    stop(paste0(this.function.name, ": message must not be NULL"))
  }
  if (!condition) {
    stop(message)
  }
}

contract <- function(condition, message) {
  
  this.function.name <- rlang::call_name(rlang::current_call())
  
  stop.running(
    condition = !missing(condition),
    message = paste0(this.function.name, ": condition must not be missing")
  )
  stop.running(
    condition = !is.null(message),
    message = paste0(this.function.name, ": message must not be NULL")
  )
  stop.running(
    condition = is.character(message),
    message = paste0(this.function.name, ": message must be a character string")
  )
  stop.running(
    condition = length(message) == 1,
    message = paste0(this.function.name, ": message must be a single string")
  )

  return(
    structure(
      list(
        expr = substitute(condition),
        message = message
      ),
      class = "contract"
    )
  )
}

DbC <- function(contracts) {
  
  this.function.name <- rlang::call_name(rlang::current_call())
  function.name <- rlang::call_name(rlang::caller_call())
  
  stop.running(
    condition = !is.null(contracts),
    message = paste0(this.function.name, ": contracts must not be NULL")
  )
  stop.running(
    condition = is.list(contracts),
    message = paste0(this.function.name, ": contracts must be a list")
  )
  stop.running(
    condition = length(contracts) > 0,
    message = paste0(
      this.function.name,
      ": there must be at least one condition to check"
    )
  )
  stop.running(
    condition = all(vapply(contracts, inherits, logical(1), what = "contract")),
    message = paste0(
      this.function.name,
      ": every element of contracts must be built with contract()"
    )
  )

  for (i in seq_along(contracts)) {
    stop.running(
      condition = eval(contracts[[i]]$expr, envir = parent.frame()),
      message = paste0(function.name, ": ", contracts[[i]]$message)
    )
  }
}
```

\newpage
# Construct Dataset Function (RHC)
```{r}
construct_data_set <- function() {
  Hmisc::getHdata(rhc)
  # primary analysis variables
  time <- pmax(rhc$dthdte, rhc$lstctdte, na.rm = TRUE) - rhc$sadmdte
  Delta <- ifelse(rhc$death == "Yes",1,0)
  C <- 1 - Delta
  A <- ifelse(rhc$swang1 == "RHC",1,0)
  
  # demographic and socioeconomic characteristics
  Gender <- rhc$sex
  Age <- as.numeric(rhc$age)
  Race <- droplevels(rhc$race)
  Education <- as.numeric(rhc$edu)
  Income <- rhc$income
  InsuranceClass <- rhc$ninsclas
  ## Patient weight in kilograms
  ## Note: The original dataset encodes some missing values as 0. These values
  ## are retained because replacing them with NA substantially alters the
  ## estimated propensity scores and covariate balance following weighting.
  Weight <- as.numeric(rhc$wtkilo1)
  
  # pre-existing comorbodities
  hx.vars <- grep("hx$", names(rhc), value = TRUE)
  rhc[hx.vars] <- lapply(rhc[hx.vars], function(x) {
    x <- as.numeric(x)
  
    factor(
      ifelse(x == 1, "Yes", "No"),
      levels = c("No", "Yes")
    )
  })
  
  HX.Variables <- rhc[hx.vars]
  names(HX.Variables) <- paste0(
    toupper(sub("hx$", "", names(HX.Variables))),
    "hx"
  )
  
  # clinical characteristics
  Cancer <- rhc$ca
  Category <- droplevels(rhc$cat1)
  Severity <- as.numeric(rhc$das2d3pc)
  
  # construct data set
  data.set <- data.frame(
    time = time,
    Delta = Delta,
    C = C,
    A = A,
    
    Gender = Gender,
    Age = Age,
    Weight = Weight,
    Race = Race,
    Education = Education,
    Income = Income,
    InsuranceClass = InsuranceClass,
    
    HX.Variables,
    
    Cancer = Cancer,
    Category = Category,
    Severity = Severity
  )
  
  #-----------------------------------  POST   --------------------------------#
  DbC(
    contracts = list(
      contract(
        is.data.frame(data.set),
        "construct_data_set: data.set must be a data.frame"
      ),
      contract(
        is.data.frame(data.set),
        "construct_data_set: data.set must be a data.frame"
      ),
      contract(
        nrow(data.set) > 0,
        "construct_data_set: data.set has zero rows"
      ),
      contract(
        all(c("time", "Delta", "C", "A") %in% names(data.set)),
        "construct_data_set: missing required columns time/Delta/C/A"
      ),
      contract(
        !any(is.na(data.set$time)) && all(data.set$time >= 0),
        "construct_data_set: time must be non-negative and non-missing"
      ),
      contract(
        all(data.set$Delta %in% c(0, 1)),
        "construct_data_set: Delta must be binary 0/1"
      ),
      contract(
        all(data.set$C %in% c(0, 1)),
        "construct_data_set: C must be binary 0/1"
      ),
      contract(
        all(data.set$A %in% c(0, 1)),
        "construct_data_set: A must be binary 0/1"
      )
    )
  )
  #-----------------------------------  POST   --------------------------------#
  
  return(data.set)
}
```

\newpage
# Propensity Score Formula Function
```{r}
propensity_score_form <- function(
    data.set, covariates, outcome, additional = ""  
  ) {

  #-----------------------------------   PRE   --------------------------------#  
  DbC(
    contracts = list(
      contract(
        is.data.frame(data.set),
        "data.set must be a data.frame"
      ),
      contract(
        is.character(covariates),
        "covariates must be a character vector"
      ),
      contract(
        length(covariates) > 0,
        "covariates must be non-empty"
      ),
      contract(
        all(covariates %in% names(data.set)),
        "all covariates must be columns of data.set"
      ),
      contract(
        is.character(outcome),
        "outcome must be a character string"
      ),
      contract(
        length(outcome) == 1,
        "outcome must be a single string"
      ),
      contract(
        outcome %in% names(data.set),
        "outcome must be a column of data.set"
      ),
      contract(
        is.character(additional),
        "additional must be a character vector"
      ),
      contract(
        !any(additional != "" & !(additional %in% names(data.set))),
        "additional must reference columns of data.set"
      )
    )
  )
  #-----------------------------------   PRE   --------------------------------#
  
  all.covariates <- c(
    covariates,
    additional[additional != ""]
  )

  continuous <- vapply(
    all.covariates,
    function(covariate) {

      x <- data.set[[covariate]]

      if (!is.numeric(x) || is.integer(x)) {
        return(FALSE)
      }

      x <- x[!is.na(x)]

      non.integer.values <- unique(
        x[x != trunc(x)]
      )

      length(non.integer.values) >= nrow(data.set) * 0.1
    },
    logical(1)
  )

  cont.covariates <- all.covariates[continuous]
  non.cont.covariates <- all.covariates[!continuous]

  spline.terms <- paste0(
    "splines::ns(",
    cont.covariates,
    ", df = 2)"
  )

  linear.terms <- non.cont.covariates

  rhs.terms <- c(
    spline.terms,
    linear.terms
  )
  
  result <- as.formula(
    paste(
      outcome,
      "~ 0 +",
      paste(rhs.terms, collapse = " + ")
    )
  )
  
  #-----------------------------------  POST   --------------------------------#
  DbC(
    contracts = list(
      contract(
        inherits(result, "formula"),
        "failed to construct a valid formula"
      ),
      contract(
        outcome %in% all.vars(result),
        "formula must contain the outcome variable"
      ),
      contract(
        all(all.covariates %in% all.vars(result)),
        "formula must contain all covariates and additional variables"
      )   
    )
  )
  #-----------------------------------  POST   --------------------------------#

  return(result)
}
```

\newpage
# Propensity Score Function
```{r}
propensity_score <- function(
    data.set,
    outcome, covariates,
    propensity.score.form,
    additional = ""  
  ) {
  
  #-----------------------------------   PRE   --------------------------------#
  DbC(
    contracts = list(
      contract(
        is.data.frame(data.set),
        "data.set must be a data.frame"
      ),
      contract(
        is.character(outcome) && length(outcome) == 1,
        "outcome must be a single string"
      ),
      contract(
        outcome %in% names(data.set),
        "outcome must be a column of the data set"
      ),
      contract(
        is.character(covariates) && length(covariates) > 0,
        "covariates must be a non-empty string vector"
      ),
      contract(
        all(covariates %in% names(data.set)),
        "covariates must be columns of the data set"
      ),
      contract(
        is.function(propensity.score.form),
        "propensity.score.form must be a function"
      ),
      contract(
        is.character(additional),
        "additional must be a string vector"
      ),
      contract(
        !any(additional != "" & !(additional %in% names(data.set))),
        "additional must reference columns of the data set"
      )
    )
  )
  #-----------------------------------   PRE   --------------------------------#
  
  formula <- propensity.score.form(
    data.set = data.set,
    covariates = covariates,
    outcome = outcome,
    additional = additional
  )
  
  result <- fitted(
    glm(
      formula = formula,
      data = data.set,
      family = binomial(link = "logit")
    )
  )
  
  #-----------------------------------  POST   --------------------------------#
  DbC(
    contracts = list(
      contract(
        is.numeric(result) && all(result >= 0 & result <= 1, na.rm = TRUE),
        "propensity scores must be numeric values between 0 and 1"
      )   
    )
  )
  #-----------------------------------  POST   --------------------------------#
  
  # return the fitted propensity score model
  return(result)
}
```

\newpage
# Plot Propensity Score Function
```{r}
plot_propensity_scores <- function(
    propensity.score.0, propensity.score.1,
    cols, fills,
    title = "Propensity Score Overlap"
  ) {
  
  #-----------------------------------   PRE   --------------------------------#
  DbC(
    contracts = list(
      contract(
        is.numeric(propensity.score.0),
        "propensity.score.0 must be numeric"
      ),
      contract(
        all(
          propensity.score.0 >= 0 &
          propensity.score.0 <= 1,
          na.rm = TRUE
        ),
        "propensity.score.0 must contain values between 0 and 1"
      ),
      contract(
        is.numeric(propensity.score.1),
        "propensity.score.1 must be numeric"
      ),
      contract(
        all(
          propensity.score.1 >= 0 &
          propensity.score.1 <= 1,
          na.rm = TRUE
        ),
        "propensity.score.1 must contain values between 0 and 1"
      ),
      contract(
        is.character(cols),
        "cols must be a character vector"
      ),
      contract(
        length(cols) >= 2,
        "cols must contain at least two elements"
      ),
      contract(
        is.character(fills),
        "fills must be a character vector"
      ),
      contract(
        length(fills) >= 2,
        "fills must contain at least two elements"
      ),
      contract(
        is.character(title) && length(title) == 1,
        "title must be a single string"
      )
    )
  )
  #-----------------------------------   PRE   --------------------------------#
  
  par(
    mar = c(2,2,2,0)
  )
  
  # plot propensity scores for individuals with A == 0
  hist(
    propensity.score.0,
    col = fills[1],
    breaks = seq(0,1,by=0.02),
    ylim = c(0,200),
    main = title,
    xlab = "Propensity Scores"
  )
  
  # plot propensity scores for individuals with A == 1
  hist(
    propensity.score.1,
    col = fills[2],
    breaks = seq(0,1,by=0.02),
    add = T
  )
  
  box()
  
  legend(
    "topright",
    legend = c('A = 0', 'A = 1'),
    fill = cols,
    border = "black",
    bty = "n"
  )
}
```

\newpage
# Stabilized IPW Function
```{r}
stabilized_IPW <- function(propensity.score, outcome) {
  
  #-----------------------------------   PRE   --------------------------------#
  DbC(
    contracts = list(
      contract(
        is.numeric(propensity.score),
        "propensity.score must be numeric"
      ),
      contract(
        is.numeric(outcome),
        "outcome must be numeric"
      ),
      contract(
        length(propensity.score) == length(outcome),
        "propensity.score and outcome must have the same length"
      ),
      contract(
        all(propensity.score >= 0 & propensity.score <= 1, na.rm = TRUE),
        "propensity.score must contain values between 0 and 1"
      ),
      contract(
        all(outcome %in% c(0, 1)),
        "outcome must contain only 0 and 1"
      )
    )
  )
  #-----------------------------------   PRE   --------------------------------#
  
  p.mean <- mean(outcome)
  
  # calculate stabilized ipws
  ipw.stabilized <- ifelse(
    outcome == 1,
    p.mean / propensity.score,
    (1 - p.mean) / (1 - propensity.score)
  )
  
  #-----------------------------------  POST   --------------------------------#
  DbC(
    contracts = list(
      contract(
        is.numeric(ipw.stabilized),
        "stabilized IPWs must be numeric"
      ),
      contract(
        all(ipw.stabilized >= 0, na.rm = TRUE),
        "stabilized IPWs must be non-negative"
      )
    )
  )
  #-----------------------------------  POST   --------------------------------#
  
  return(ipw.stabilized)
}
```

\newpage
# Plot Multiple IPWs Function
```{r}
plot_ipw <- function(ipws, titles, cols) {
  
  #-----------------------------------   PRE   --------------------------------#
  DbC(
    contracts = list(
      contract(
        is.list(ipws),
        "ipws must be a list"
      ),
      contract(
        length(ipws) > 0,
        "ipws must be non-empty"
      ),
      contract(
        all(vapply(ipws, is.numeric, logical(1))),
        "all IPWs must be numeric"
      ),
      contract(
        all(vapply(ipws, function(x) all(x >= 0, na.rm = TRUE), logical(1))),
        "all IPWs must be non-negative"
      ),
      contract(
        is.character(titles),
        "titles must be a string vector"
      ),
      contract(
        length(titles) >= length(ipws),
        "titles must contain at least one title for each IPW"
      ),
      contract(
        is.character(cols),
        "cols must be a string vector"
      ),
      contract(
        length(cols) >= length(ipws),
        "cols must contain at least one color for each IPW"
      )
    )
  )
  #-----------------------------------   PRE   --------------------------------#
  
  cols.alpha <- adjustcolor(cols, alpha.f = 0.5)
  par(mar = c(2, 2, 2, 0))
  
  # common breaks for all IPWs
  breaks <- seq(
    0,
    max(unlist(ipws)) + 0.5,
    by = 0.1
  )
  
  # plot first ipw distribution
  hist(
    ipws[[1]],
    col = cols.alpha[1],
    breaks = breaks,
    xlim = c(0, 5),
    main = "IPW Comparison",
    ylab = "Frequency",
    xlab = "IPWs"
  )
  
  # add remaining ipw distributions
  if (length(ipws) > 1) {
    for (i in 2:length(ipws)) {
      hist(
        ipws[[i]],
        col = cols.alpha[i],
        breaks = breaks,
        add = TRUE
      )
    }
  }
  
  box()
  
  legend(
    "topright",
    legend = titles,
    fill = cols,
    border = "black",
    bty = "n"
  )
}
```

\newpage
# Weighted Descriptive Statistics: Table Two Function
```{r}
weighted_table_one <- function(data.set, weights, covariates, outcome) {
  
  #-----------------------------------   PRE   --------------------------------#
  DbC(
    contracts = list(
      contract(
        is.data.frame(data.set),
        "data.set must be a data.frame"
      ),
      contract(
        is.numeric(weights),
        "weights must be numeric"
      ),
      contract(
        length(weights) == nrow(data.set),
        "weights must have the same length as the number of rows in data.set"
      ),
      contract(
        all(weights >= 0, na.rm = TRUE),
        "weights must be non-negative"
      ),
      contract(
        is.character(covariates) && length(covariates) > 0,
        "covariates must be a non-empty string vector"
      ),
      contract(
        all(covariates %in% names(data.set)),
        "covariates must be columns of the data set"
      ),
      contract(
        is.character(outcome) && length(outcome) == 1,
        "outcome must be a single string"
      ),
      contract(
        outcome %in% names(data.set),
        "outcome must be a column of the data set"
      )
    )
  )
  #-----------------------------------   PRE   --------------------------------#
  
  # create weighted data set
  data.set.ipw <- survey::svydesign(
    ids = ~ 0,
    data = data.set,
    weights = weights
  )
  
  # create table one for the weighted data set
  table.one.ipw <- tableone::svyCreateTableOne(
    vars = covariates,
    strata = outcome,
    data = data.set.ipw,
    test = FALSE
  )
  
  print(table.one.ipw, smd = TRUE)
}
```

\newpage
# Plot Covariate Balance
```{r}
plot_cov_balance <- function(
    data.set, covariates.added, outcome, weights.list, weights.titles, cols  
  ) {
  
  #-----------------------------------   PRE   --------------------------------#
  DbC(
    contracts = list(
      contract(
        is.data.frame(data.set),
        "data.set must be a data.frame"
      ),
      contract(
        is.character(covariates.added) && length(covariates.added) == 1,
        "covariates.added must be a single string"
      ),
      contract(
        nzchar(covariates.added),
        "covariates.added must be non-empty"
      ),
      contract(
        is.list(weights.list) && length(weights.list) > 0,
        "weights.list must be a non-empty list"
      ),
      contract(
        all(vapply(weights.list, is.numeric, logical(1))),
        "all elements of weights.list must be numeric"
      ),
      contract(all(vapply(
        weights.list, function(x) length(x) == nrow(data.set), logical(1)
      )),
        "all weight vectors must have the same length as the data set"
      ),
      contract(all(vapply(
        weights.list, function(x) all(x >= 0, na.rm = TRUE), logical(1))
      ),
        "all weights must be non-negative"
      ),
      contract(
        is.character(weights.titles),
        "weights.titles must be a string vector"
      ),
      contract(
        length(weights.titles) >= length(weights.list),
        "weights.titles must contain at least one title for each weight set"
      ),
      contract(
        is.character(cols),
        "cols must be a character vector"
      ),
      contract(
        length(cols) >= length(weights.list) + 1,
        "cols must contain at least one color for each weight set"
      ),
      contract(
        is.character(outcome) && length(outcome) == 1,
        "outcome must be a single string"
      ),
      contract(
        outcome %in% names(data.set),
        "outcome must be a column of the data set"
      )
    )
  )
  #-----------------------------------   PRE   --------------------------------#
  
  balance.types <- length(weights.list)
  
  # access covariate balance before and after weighting
  balance <- cobalt::bal.tab(
    as.formula(paste0(outcome, " ~ ", covariates.added)),
    data = data.set,
    weights = weights.list,
    method = "weighting",
    estimand = "ATE",
    un = TRUE
  )
  
  # plot the covariate balance
  cobalt::love.plot(
    balance,
    stats = "mean.diffs",
    threshold = 0.1,
    abs = TRUE,
    line = TRUE,
    var.order = "unadjusted",
    sample.names = c("Unadjusted", weights.titles),
    colors = cols[seq(1, 1+balance.types, 1)],
    shapes = seq(17, 17-balance.types, -1),
    limits = c(m = c(0, 0.3)),
    title = "Covariate Balance Before and After Weighting",
    labels = TRUE
  )
}
```

\newpage
# Estimation

# Modified Case-Base Smooth Hazard Model
```{r}
## Modified implementation of fitSmoothHazard() from the casebase package.
## The function extends the standard case-base fitting procedure to allow
## observation-level weights, including inverse probability weights (IPWs).
FSH <- function(
    formula,
    data,
    time,
    weights = rep(1, nrow(data)),
    ratio = 100,
    ...
) {
  
  #-----------------------------------   PRE   --------------------------------#
  DbC(
    contracts = list(
      contract(
        inherits(try(formula(formula), silent = TRUE), "formula"),
        "formula must be a valid formula"
      ),
      contract(
        length(all.vars(formula[[2]])) == 1,
        "formula must contain exactly one event variable"
      ),
      contract(
        is.data.frame(data) || inherits(data, "singleEventCB"),
        "data must be a data.frame or a singleEventCB object"
      ),
      contract(
        is.numeric(weights),
        "weights must be numeric"
      ),
      contract(
        length(weights) == nrow(data),
        "weights must have the same length as the number of rows in data"
      ),
      contract(
        all(!is.na(weights)),
        "weights must not contain missing values"
      )
    )
  )
  #-----------------------------------   PRE   --------------------------------#
  
  # preserve the original function call for reproducibility
  orig.call <- match.call()
  
  ## Expand "." in a model formula into the complete set of covariates available
  ## in the supplied dataset. This ensures that the modified FSH function can
  ## explicitly evaluate formulas containing "." before fitting the model.
  expand.dot.formula <- function(formula, data = NULL) {
    
    ## Check whether the formula contains the "." shorthand for all available
    ## predictors
    if (isTRUE("." %in% all.vars(formula))) {
      
      ## Preserve the original formula attributes while expanding the terms
      formula.attrs <- attributes(formula)
      
      try.terms <- try(
        terms(formula, data = data),
        silent = TRUE
      )
      
      ## Replace the shorthand formula with its explicitly expanded version when
      ## the terms object can be constructed successfully
      if (!is(try.terms, "try-error")) {
        formula <- formula(try.terms)
      }
      
      ## Restore the original formula attributes after expansion
      attributes(formula) <- formula.attrs
    }
    
    ## Return the expanded formula
    return(formula)
  }
  
  formula <- expand.dot.formula(formula, data = data)
  
  # extract the event variable from the left-hand side of the model formula
  event.var <- all.vars(formula[[2]])
  
  #extract time variable from the data set
  if (missing(time)) {
    var.names <- casebase::checkArgsTimeEvent(
      data = data,
      event = event.var
    )
    time.var <- var.names$time
    
  } else {
    time.var <- time
  }
  
  ## Identify the distinct event types present in the data to determine whether
  ## the analysis involves a single event or competing risks
  type.events <- sort(unique(data[[event.var]]))
  
  
  # Case-Base Sampling
  
  ## Generate the case-base sample when the input data have not already been
  ## sampled. The original dataset is retained so that the supplied observation
  ## weights can be incorporated into the case-base sampling procedure.
  if (!inherits(data, "cbData")) {
    
    original.data <- as.data.frame(data)
    original.data$weights <- weights
    
    ## Sample cases and controls from the person-time data. The sampling ratio
    ## controls the number of person-moments sampled relative to observed cases.
    sample.data <- casebase::sampleCaseBase(
      original.data,
      time.var,
      event.var,
      comprisk = (length(type.events) > 2),
      ratio
    )
    
  } else {
    ## If the input is already a case-base dataset, retain it directly.
    ## The original unsampled data are unavailable at this stage.
    original.data <- data.frame()
    sample.data <- data
  }
  
  
  # Weighted Case-Base Model
  
  ## Include the case-base sampling offset in the model formula. This offset
  ## accounts for the sampling mechanism when estimating the hazard function.
  formula <- update(formula, ~ . + offset(offset))
  
  ## Fit a weighted binomial regression model to the case-base sample.
  ## Observation-level weights allow the model to incorporate IPWs in addition
  ## to the case-base sampling structure.
  fitting.function <- function(formula) glm(
    formula,
    data = sample.data,
    family = binomial,
    weights = weights
  )
  
  ## Estimate the smooth hazard model using the weighted case-base sample
  out <- fitting.function(formula)
  
  
  # Model Object and Metadata
  
  ## Preserve the lower-level glm call while replacing the displayed call with
  ## the original FSH call for consistency with the custom model interface
  out$lower_call <- out$call 
  out$call <- orig.call
  
  ## Store the original data and key variable names for downstream methods
  ## such as prediction and plotting
  out$originalData <- original.data
  out$typeEvents <- type.events
  out$timeVar <- time.var
  out$eventVar <- event.var
  
  ## Record the number of case and background person-moments in the sampled
  ## case-base dataset
  num.person.moments <- table(sample.data[[event.var]])
  out$num_cm <- num.person.moments[2]
  out$num_bm <- num.person.moments[1]
  out$formula <- formula
  
  ## Store the case-base sampling offset separately, then reset the model's
  ## offset to zero so that absolute risk estimation can be performed without
  ## retaining the sampling correction in the prediction stage
  out$offset <- out$data$offset
  out$data$offset <- 0
  
  ## Assign a custom class so that methods defined for single-event case-base
  ## models can be applied to the fitted object
  class(out) <- c("singleEventCB", class(out))
  
  #-----------------------------------  POST   --------------------------------#
  DbC(
    contracts = list(
      contract(
        inherits(out, "singleEventCB"),
        "result must inherit from singleEventCB"
      ),
      contract(
        is.data.frame(out$originalData),
        "originalData must be a data.frame"
      ),
      contract(
        is.double(out$typeEvents),
        "typeEvents must be a double vector"
      ),
      contract(
        is.character(out$timeVar) && length(out$timeVar) == 1,
        "timeVar must be a single character string"
      ),
      contract(
        out$timeVar %in% names(out$originalData),
        "timeVar must be a column of out$originalData"
      ),
      contract(
        is.character(out$eventVar) && length(out$eventVar) == 1,
        "eventVar must be a single character string"
      ),
      contract(
        out$eventVar %in% names(out$originalData),
        "eventVar must be a column of out$originalData"
      ),
      contract(
        out$eventVar == event.var,
        "eventVar must equal the event variable from the input formula"
      ),
      contract(
        is.integer(out$num_cm) && length(out$num_cm) == 1,
        "num_cm must be one integer"
      ),
      contract(
        is.integer(out$num_bm) && length(out$num_bm) == 1,
        "num_bm must be one integer"
      ),
      contract(
        out$num_bm > out$num_cm,
        "num_bm must be greater than out$num_cm"
      ),
      contract(
        inherits(out$formula, "formula"),
        "formula must be a valid formula"
      ),
      contract(
        is.double(out$offset),
        "offset must be a double vector"
      ),
      contract(
        length(out$offset) == nrow(out$data),
        "offset must have the same length as the number of rows in out$data"
      ),
      contract(
        is.data.frame(out$data),
        "data must be a data.frame"
      ),
      contract(
        nrow(out$data) > 0,
        "data must have at least one row"
      ),
      contract(
        all(c(out$eventVar, out$timeVar) %in% names(out$data)),
        "data must contain the event and time variables"
      ),
      contract(
        is.language(out$lower_call),
        "lower_call must be a language object"
      ),
      contract(
        is.language(out$call),
        "call must be a language object"
      )
    )
  )
  #-----------------------------------  POST   --------------------------------#
  
  return(out)
}
```

\newpage
# Extract and Report Hazard Ratio
```{r}
## Extract the treatment coefficient, calculate its standard error, and report
## the hazard ratio with its corresponding 95% Wald confidence interval.
extract_print_hr <- function(model, se.func) {
  
  #-----------------------------------   PRE   --------------------------------#
  DbC(
    contracts = list(
      contract(
        is.function(se.func),
        "se.func must be a function"
      ),
      contract(
        "A" %in% names(coef(model)),
        "model must contain a coefficient named A"
      ),
      contract(
        "A" %in% rownames(se.func(model)),
        "se.func(model) must contain a standard error for A"
      )
    )
  )
  #-----------------------------------   PRE   --------------------------------#
  
  # Extract the treatment coefficient and its standard error
  beta <- coef(model)["A"]
  se <- sqrt(diag(se.func(model))["A"])
  
  # Transform the treatment coefficient from the log-hazard ratio scale to the
  # hazard ratio scale and calculate the corresponding 95% Wald confidence
  # interval
  cat("HR-hat:", exp(beta), "\n")
  cat("SE of HR-hat:", se, "\n")
  cat(
    "95% CI: (",
    exp(beta + c(-1, 1) * 1.96 * se),
    ")\n"
  )
}
```

\newpage
# Fit Model
```{r}
fit_model <- function(
    fitting.function,
    formula,
    data.set,
    seed,
    extract.print.hr,
    constant.HR = TRUE,
    se.func = NULL,
    ...  
  ) {
  
  #-----------------------------------   PRE   --------------------------------#
  DbC(
    contracts = list(
      contract(
        is.function(fitting.function),
        "fitting.function must be a function"
      ),
      contract(
        inherits(try(formula(formula), silent = TRUE), "formula"),
        "formula must be a valid formula"
      ),
      contract(
        is.data.frame(data.set),
        "data.set must be a data.frame"
      ),
      contract(
        is.numeric(seed) && length(seed) == 1,
        "seed must be a single numeric value"
      ),
      contract(
        is.function(extract.print.hr),
        "extract.print.hr must be a function"
      ),
      contract(
        is.logical(constant.HR) && length(constant.HR) == 1,
        "constant.HR must be a single logical value"
      ),
      contract(
        is.function(se.func) || is.null(se.func),
        "se.func must be a function or NULL"
      ),
      contract(
        !constant.HR || is.function(se.func),
        "se.func must be a function when constant.HR is TRUE"
      )
    )
  )
  #-----------------------------------   PRE   --------------------------------#
  
  ## Set seed to ensure reproducibility of any stochastic procedures used
  ## during model fitting
  set.seed(seed)
  
  ## Evaluate arguments supplied through ...
  dots <- list(...)
  
  ## Fit Model
  model <- do.call(
    fitting.function,
    c(
      list(
        formula = formula,
        data = data.set
      ),
      dots
    )
  )
  
  if (constant.HR) {
    ## Extract and report the hazard ratio
    extract.print.hr(model, se.func) 
  }
  
  result <- model
  
  #-----------------------------------  POST   --------------------------------#
  DbC(
    contracts = list(
      contract(
        !is.null(result),
        "fit_model must return a non-NULL fitted model"
      )
    )
  )
  #-----------------------------------  POST   --------------------------------#
  
  return(result)
}
```

\newpage
# Case Base Variable HR Extraction
```{r}
casebase_var_hr <- function(model, time.grid) {
  
  #-----------------------------------   PRE   --------------------------------#
  DbC(
    contracts = list(
      contract(
        !is.null(model),
        "model must be a valid fitted model object"
      ),
      contract(
        is.numeric(time.grid),
        "time.grid must be numeric"
      ),
      contract(
        length(time.grid) > 0,
        "time.grid must contain at least one point"
      ),
      contract(
        all(diff(time.grid) > 0),
        "time.grid must be strictly increasing"
      )
    )
  )
  #-----------------------------------   PRE   --------------------------------#
  
  invisible(pdf(NULL))
  
  out <- plot(
    model,
    type = "hr",
    newdata = data.frame(
      time = time.grid,
      A = 0
    ),
    var = "A",
    xvar = "time",
    ci = TRUE
  )
  
  invisible(dev.off())
  
  #-----------------------------------  POST   --------------------------------#
  DbC(
    contracts = list(
      contract(
        !is.null(out),
        "casebase_var_hr must return a non-NULL result"
      )
    )
  )
  #-----------------------------------  POST   --------------------------------#
  
  return(out)
}
```

\newpage
# Plot time-dependent Hazard Ratios
```{r}
plot_time_dep_hr <- function(
    models.list, main.title, time.grid, 
    casebase.var.hr,
    titles = "", cols = "red" 
  ) {
  
  #-----------------------------------   PRE   --------------------------------#
  DbC(
    contracts = list(
      contract(
        is.list(models.list),
        "models.list must be a list"
      ),
      contract(
        length(models.list) > 0,
        "models.list must be non-empty"
      ),
      contract(
        is.character(main.title) && length(main.title) == 1,
        "main.title must be a single string"
      ),
      contract(
        is.numeric(time.grid),
        "time.grid must be numeric"
      ),
      contract(
        length(time.grid) > 0,
        "time.grid must contain at least one point"
      ),
      contract(
        all(diff(time.grid) > 0),
        "time.grid must be strictly increasing"
      ),
      contract(
        is.function(casebase.var.hr),
        "casebase.var.hr must be a function"
      ),
      contract(
        is.character(titles),
        "titles must be a character vector"
      ),
      contract(
        length(titles) >= length(models.list),
        "titles must contain at least one title for each model"
      ),
      contract(
        is.character(cols),
        "cols must be a character vector"
      ),
      contract(
        length(cols) >= length(models.list),
        "cols must contain at least one color for each model"
      )
    )
  )
  #-----------------------------------   PRE   --------------------------------#
  
  # global function consts
  hr.list <- list()
  cols.alpha <- adjustcolor(cols, alpha.f = 0.2)
  
  # extract all hr curves
  for (i in 1:length(models.list)) {
    hr.list[[i]] <- casebase.var.hr(
      model = models.list[[i]],
      time.grid = time.grid
    )
  }
  
  # common time grid
  common.time <- time.grid
  
  # common HR grid
  common.hr <- range(
    unlist(lapply(hr.list, function(x) x$hazard_ratio)),
    na.rm = TRUE
  )
  
  # define common time and hr grids here then use in plot below
  
  # define base plot
  plot(
    NA,
    xlim = range(common.time),
    ylim = range(common.hr),
    type = "n",
    xlab = "Time",
    ylab = "Hazard Ratio",
    main = main.title
  )
  
  # plot all ci bands
  for (i in 1:length(hr.list)) {
    polygon(
      c(hr.list[[i]]$time, rev(hr.list[[i]]$time)),
      c(hr.list[[i]]$lowerbound, rev(hr.list[[i]]$upperbound)),
      col = cols.alpha[i],
      border = NA
    )
  }
  
  # plot all hr curves on top
  for (i in 1:length(hr.list)) {
    lines(
      hr.list[[i]]$time,
      hr.list[[i]]$hazard_ratio,
      col = cols[i],
      lwd = 2
    )
  }
  
  # plot null hr reference line
  abline(
    h = 1,
    col = "black",
    lty = 2
  )
  
  if (length(models.list) > 1) {
    # create legend
    legend(
      "bottomleft",
      legend = titles,
      col = cols,
      lwd = 2,
      bty = "n"
    ) 
  }
}
```

\newpage
# Restricted Mean Survival Time (RMST)
```{r}
## Numerically integrate a survival curve over the specified follow-up period
## using the trapezoidal rule. The cumulative integral at each time point
## represents the RMST up to that point.
rmst_calculation <- function(time, surv) {
  
  #-----------------------------------   PRE   --------------------------------#
  DbC(
    contracts = list(
      contract(
        is.numeric(time),
        "time must be numeric"
      ),
      contract(
        length(time) > 0,
        "time must contain at least one point"
      ),
      contract(
        all(diff(time) > 0),
        "time must be strictly increasing"
      ),
      contract(
        is.numeric(surv),
        "surv must be numeric"
      ),
      contract(
        length(surv) == length(time),
        "surv must have the same length as time"
      ),
      contract(
        all(surv >= 0 & surv <= 1, na.rm = TRUE),
        "surv must contain values between 0 and 1"
      )
    )
  )
  #-----------------------------------   PRE   --------------------------------#
  
  time.diff <- diff(time)
  midpoints <- (head(surv, -1) + tail(surv, -1)) / 2
  result <- c(0, cumsum(midpoints * time.diff))
  
  #-----------------------------------  POST   --------------------------------#
  DbC(
    contracts = list(
      contract(
        is.numeric(result),
        "RMST calculation must return a numeric vector"
      ),
      contract(
        length(result) == length(time),
        "RMST calculation result must have the same length as time"
      ),
      contract(
        result[1] == 0,
        "RMST calculation result must start at 0"
      )
    )
  )
  #-----------------------------------  POST   --------------------------------#
  
  return(result)
}
```

\newpage
# MS-Cox RMST and Standard Error
```{r}
## Calculate the Counter-factual RMST contrast and its standard error from the
## marginal structural Cox model. The RMST contrast is defined as the
## difference in restricted mean survival between RHC and no RHC.
cox_rmst_list <- function(
    ms.cox, time.grid, cox.psi.se, survival.curves.data, rmst.calculation
  ) {
  
  #-----------------------------------   PRE   --------------------------------#
  DbC(
    contracts = list(
      contract(
        !is.null(ms.cox),
        "ms.cox must be a valid fitted Cox model"
      ),
      contract(
        is.numeric(time.grid),
        "time.grid must be numeric"
      ),
      contract(
        length(time.grid) > 0,
        "time.grid must contain at least one point"
      ),
      contract(
        all(diff(time.grid) > 0),
        "time.grid must be strictly increasing"
      ),
      contract(
        is.numeric(cox.psi.se) && length(cox.psi.se) == 1,
        "cox.psi.se must be a single numeric value"
      ),
      contract(
        is.data.frame(survival.curves.data),
        "survival.curves.data must be a data.frame"
      ),
      contract(
        is.function(rmst.calculation),
        "rmst.calculation must be a function"
      )
    )
  )
  #-----------------------------------   PRE   --------------------------------#
  
  # Counter-factual Survival Curves
  surv <- survival::survfit(
    ms.cox,
    newdata = survival.curves.data
  )
  
  # Interpolate the Counter-factual survival curve under no RHC onto
  # the specified time grid
  S0 <- approx(
    surv$time,
    surv$surv[, 1],
    time.grid,
    method = "constant",
    rule = 2
  )$y
  
  # Interpolate the Counter-factual survival curve under RHC onto
  # the specified time grid
  S1 <- approx(
    surv$time,
    surv$surv[, 2],
    time.grid,
    method = "constant",
    rule = 2
  )$y
  
  # RMST Contrast
  rmst <- rmst.calculation(time.grid, S1) -
    rmst.calculation(time.grid, S0)
  
  # Delta-Method Standard Error
  integrand <- ifelse(S1 > 0, S1 * log(S1), 0)
  cox.gradient <- rmst.calculation(time.grid, integrand)
  
  rmst.se <- abs(cox.gradient) * cox.psi.se
  
  result <- list(
    rmst = rmst,
    rmst.se = rmst.se
  )
  
  #-----------------------------------  POST   --------------------------------#
  DbC(
    contracts = list(
      contract(
        is.list(result),
        "result must be a list"
      ),
      contract(
        is.numeric(result$rmst),
        "rmst must be numeric"
      ),
      contract(
        length(result$rmst) == length(time.grid),
        "rmst must have the same length as time.grid"
      ),
      contract(
        is.numeric(result$rmst.se),
        "rmst.se must be numeric"
      ),
      contract(
        length(result$rmst.se) == length(time.grid),
        "rmst.se must have the same length as time.grid"
      )
    )
  )
  #-----------------------------------  POST   --------------------------------#
  
  return(result)
}
```

\newpage
# casebase Counter-factual Survival Function
```{r}
## Construct the Counter-factual survival function under treatment level a
## directly from the case-base model's design matrix
casebase_surv <- function(beta, a, time.grid, term.formula) {
  
  #-----------------------------------   PRE   --------------------------------#
  DbC(
    contracts = list(
      contract(
        is.numeric(beta),
        "beta must be numeric"
      ),
      contract(
        is.numeric(a) && length(a) == 1,
        "a must be a single numeric value"
      ),
      contract(
        is.numeric(time.grid),
        "time.grid must be numeric"
      ),
      contract(
        length(time.grid) > 0,
        "time.grid must contain at least one point"
      ),
      contract(
        all(diff(time.grid) > 0),
        "time.grid must be strictly increasing"
      ),
      contract(
        inherits(try(formula(term.formula), silent = TRUE), "formula"),
        "term.formula must be a valid formula"
      )
    )
  )
  #-----------------------------------   PRE   --------------------------------#
  
  ## Create prediction data over the specified time grid under treatment
  ## level a
  new.data <- data.frame(
    time = time.grid,
    A = a,
    offset = 0
  )
  
  ## Construct the model matrix using the same terms as the fitted model and
  ## align its columns with the estimated coefficient vector
  covariates <- model.matrix(term.formula, data = new.data)
  covariates <- covariates[, names(beta), drop = FALSE]
  
  ## Convert the linear predictor to the estimated hazard function
  lambda <- exp(as.vector(covariates %*% beta))
  
  ## Numerically integrate the estimated hazard using the trapezoidal rule
  ## to obtain the cumulative hazard
  Lambda <- c(
    0,
    cumsum(
      diff(time.grid) * (head(lambda, -1) + tail(lambda, -1)) / 2
    )
  )
  
  ## Convert cumulative hazard to survival probability
  result <- exp(-Lambda)
  
  #-----------------------------------  POST   --------------------------------#
  DbC(
    contracts = list(
      contract(
        is.numeric(result),
        "casebase_surv must return a numeric vector"
      ),
      contract(
        length(result) == length(time.grid),
        "casebase_surv result must have the same length as time.grid"
      ),
      contract(
        all(result >= 0 & result <= 1, na.rm = TRUE),
        "casebase_surv result must contain values between 0 and 1"
      ),
      contract(
        result[1] == 1,
        "casebase_surv result must start at 1"
      )
    )
  )
  #-----------------------------------  POST   --------------------------------#
  
  return(result)
}
```

\newpage
# RMST Contrast
```{r}
## Define the RMST contrast as a function of the model coefficient vector.
## This function is subsequently used to obtain the numerical gradient
## required for the delta-method standard error.
rmst_diff <- function(
    beta, time.grid, casebase.surv, rmst.calculation, term.formula
  ) {
  
  #-----------------------------------   PRE   --------------------------------#
  DbC(
    contracts = list(
      contract(
        is.numeric(beta),
        "beta must be numeric"
      ),
      contract(
        is.numeric(time.grid),
        "time.grid must be numeric"
      ),
      contract(
        length(time.grid) > 0,
        "time.grid must contain at least one point"
      ),
      contract(
        all(diff(time.grid) > 0),
        "time.grid must be strictly increasing"
      ),
      contract(
        is.function(casebase.surv),
        "casebase.surv must be a function"
      ),
      contract(
        is.function(rmst.calculation),
        "rmst.calculation must be a function"
      ),
      contract(
        inherits(try(formula(term.formula), silent = TRUE), "formula"),
        "term.formula must be a valid formula"
      )
    )
  )
  #-----------------------------------   PRE   --------------------------------#
  
  rmst.A1 <- rmst.calculation(
    time = time.grid,
    surv = casebase.surv(
      beta = beta,
      a = 1,
      time.grid = time.grid,
      term.formula = term.formula
    )
  )
  
  rmst.A0 <- rmst.calculation(
    time = time.grid,
    surv = casebase.surv(
      beta = beta,
      a = 0,
      time.grid = time.grid,
      term.formula = term.formula
    )
  )
  
  result <- rmst.A1 - rmst.A0
  
  #-----------------------------------  POST   --------------------------------#
  DbC(
    contracts = list(
      contract(
        is.numeric(result),
        "RMST difference must be numeric"
      ),
      contract(
        length(result) == length(time.grid),
        "RMST difference must have the same length as time.grid"
      )
    )
  )
  #-----------------------------------  POST   --------------------------------#
  
  return(result)
}
```

\newpage
# IPW Case-Base RMST and Standard Error
```{r}
## Calculate the RMST contrast and its standard error from the IPW case-base
## model. The survival curves are reconstructed from the fitted model and the
## variance of the RMST contrast is obtained using a numerical delta method.
casebase_rmst_list <- function(
    casebase.ipw, casebase.var, time.grid, constant.HR,
    casebase.surv, rmst.diff, rmst.calculation  
  ) {
  
  #-----------------------------------   PRE   --------------------------------#
  DbC(
    contracts = list(
      contract(
        !is.null(casebase.ipw),
        "casebase.ipw must be a fitted model"
      ),
      contract(
        is.numeric(casebase.var),
        "casebase.var must be numeric"
      ),
      contract(
        is.numeric(time.grid),
        "time.grid must be numeric"
      ),
      contract(
        length(time.grid) > 0,
        "time.grid must contain at least one point"
      ),
      contract(
        all(diff(time.grid) > 0),
        "time.grid must be strictly increasing"
      ),
      contract(
        is.logical(constant.HR) && length(constant.HR) == 1,
        "constant.HR must be a single logical value"
      ),
      contract(
        is.function(casebase.surv),
        "casebase.surv must be a function"
      ),
      contract(
        is.function(rmst.diff),
        "rmst.diff must be a function"
      ),
      contract(
        is.function(rmst.calculation),
        "rmst.calculation must be a function"
      )
    )
  )
  #-----------------------------------   PRE   --------------------------------#
  
  ## Extract the estimated model coefficients and the corresponding model
  ## terms used to construct the design matrix
  beta.hat <- coef(casebase.ipw)
  term.formula <- delete.response(terms(casebase.ipw))
  
  ## Evaluate the RMST contrast at the estimated model coefficients
  rmst <- rmst.diff(
    beta = beta.hat,
    time.grid = time.grid,
    casebase.surv = casebase.surv,
    rmst.calculation = rmst.calculation,
    term.formula = term.formula
  )
  
  # Numerical Gradient
  
  ## Calculate the numerical gradient of the RMST contrast with respect to
  ## each model coefficient using a central finite-difference approximation
  num.coefs <- length(beta.hat)
  num.time.points <- length(time.grid)
  
  gradient <- matrix(NA_real_, nrow = num.time.points, ncol = num.coefs)
  colnames(gradient) <- names(beta.hat)
  
  ## Step size used for the finite-difference approximation
  step.size <- 1e-6
  
  ## Perturb each coefficient in turn and approximate the corresponding
  ## derivative of the RMST contrast
  for (j in seq_len(num.coefs)) {
    beta.plus <- beta.minus <- beta.hat
    beta.plus[j] <- beta.plus[j] + step.size
    beta.minus[j] <- beta.minus[j] - step.size
    
    rmst.beta.plus <- rmst.diff(
      beta = beta.plus,
      time.grid = time.grid,
      casebase.surv = casebase.surv,
      rmst.calculation = rmst.calculation,
      term.formula = term.formula
    )
    
    rmst.beta.minus <- rmst.diff(
      beta = beta.minus,
      time.grid = time.grid,
      casebase.surv = casebase.surv,
      rmst.calculation = rmst.calculation,
      term.formula = term.formula
    )
    
    gradient[, j] <- (rmst.beta.plus - rmst.beta.minus) / (2 * step.size)
  }
  
  # Delta-Method Standard Error
  
  ## Align the model covariance matrix with the coefficient ordering used by
  ## the numerical gradient before applying the delta-method variance formula
  casebase.var <- casebase.var[colnames(gradient), colnames(gradient), drop = FALSE]
  
  ## Calculate the RMST standard error at each restriction time using
  ## Var{g(beta)} = gradient' Var(beta) gradient
  rmst.se <- vapply(
    seq_len(num.time.points),
    function(i) {
      gradient.row <- gradient[i, ]
      sqrt(drop(t(gradient.row) %*% casebase.var %*% gradient.row))
    },
    numeric(1)
  )
  
  result <- list(
    rmst = rmst,
    rmst.se = rmst.se
  )
  
  #-----------------------------------  POST   --------------------------------#
  DbC(
    contracts = list(
      contract(
        is.list(result),
        "result must be a list"
      ),
      contract(
        is.numeric(result$rmst),
        "rmst must be numeric"
      ),
      contract(
        length(result$rmst) == length(time.grid),
        "rmst must have the same length as time.grid"
      ),
      contract(
        is.numeric(result$rmst.se),
        "rmst.se must be numeric"
      ),
      contract(
        length(result$rmst.se) == length(time.grid),
        "rmst.se must have the same length as time.grid"
      )
    )
  )
  #-----------------------------------  POST   --------------------------------#
  
  return(result)
}
``` 

\newpage
# Counter-factual Survival Curves: MS-Cox and IPW Case-Base Models

# Calculate Counter-factual Survival Curves Function (MS-Cox)
```{r}
cox_csc <- function(model, new.data, time.grid) {
  
  #-----------------------------------   PRE   --------------------------------#
  DbC(
    contracts = list(
      contract(
        !is.null(model),
        "model must be a valid fitted model"
      ),
      contract(
        is.data.frame(new.data),
        "new.data must be a data.frame"
      ),
      contract(
        is.numeric(time.grid),
        "time.grid must be numeric"
      ),
      contract(
        length(time.grid) > 0,
        "time.grid must contain at least one point"
      ),
      contract(
        all(diff(time.grid) > 0),
        "time.grid must be strictly increasing"
      )
    )
  )
  #-----------------------------------   PRE   --------------------------------#
  
  surv <- survival::survfit(
    model,
    newdata = new.data
  )
  
  ## Interpolate the estimated survival curve under no RHC onto the common
  ## time grid used for subsequent comparison and RMST calculation
  S0 <- approx(
    surv$time,
    surv$surv[, 1],
    xout = time.grid,
    method = "constant",
    rule = 2
  )$y
  
  ## Interpolate the corresponding lower 95% confidence limit under no RHC
  S0.lower <- approx(
    surv$time,
    surv$lower[, 1],
    xout = time.grid,
    method = "constant",
    rule = 2
  )$y
  
  ## Interpolate the corresponding upper 95% confidence limit under no RHC
  S0.upper <- approx(
    surv$time,
    surv$upper[, 1],
    xout = time.grid,
    method = "constant",
    rule = 2
  )$y
  
  ## Interpolate the estimated survival curve under RHC onto the common time grid
  S1 <- approx(
    surv$time,
    surv$surv[, 2],
    xout = time.grid,
    method = "constant",
    rule = 2
  )$y
  
  ## Interpolate the corresponding lower 95% confidence limit under RHC
  S1.lower <- approx(
    surv$time,
    surv$lower[, 2],
    xout = time.grid,
    method = "constant",
    rule = 2
  )$y
  
  ## Interpolate the corresponding upper 95% confidence limit under RHC
  S1.upper <- approx(
    surv$time,
    surv$upper[, 2],
    xout = time.grid,
    method = "constant",
    rule = 2
  )$y
  
  result <- list(
    S0 = S0,
    S0.lower = S0.lower, S0.upper = S0.upper,
    S1 = S1,
    S1.lower = S1.lower, S1.upper = S1.upper
  )
  
  #-----------------------------------  POST   --------------------------------#
  DbC(
    contracts = list(
      contract(
        is.numeric(result$S0),
        "S0 must be numeric"
      ),
      contract(
        length(result$S0) == length(time.grid),
        "S0 must have the same length as time.grid"
      ),
      contract(
        is.numeric(result$S0.lower),
        "S0.lower must be numeric"
      ),
      contract(
        length(result$S0.lower) == length(time.grid),
        "S0.lower must have the same length as time.grid"
      ),
      contract(
        is.numeric(result$S0.upper),
        "S0.upper must be numeric"
      ),
      contract(
        length(result$S0.upper) == length(time.grid),
        "S0.upper must have the same length as time.grid"
      ),
      contract(
        is.numeric(result$S1),
        "S1 must be numeric"
      ),
      contract(
        length(result$S1) == length(time.grid),
        "S1 must have the same length as time.grid"
      ),
      contract(
        is.numeric(result$S1.lower),
        "S1.lower must be numeric"
      ),
      contract(
        length(result$S1.lower) == length(time.grid),
        "S1.lower must have the same length as time.grid"
      ),
      contract(
        is.numeric(result$S1.upper),
        "S1.upper must be numeric"
      ),
      contract(
        length(result$S1.upper) == length(time.grid),
        "S1.upper must have the same length as time.grid"
      ),
      contract(
        all(
          result$S0.lower <= result$S0 &
          result$S0 <= result$S0.upper,
          na.rm = TRUE
        ),
        "S0 must satisfy lower <= estimate <= upper"
      ),
      contract(
        all(
          result$S1.lower <= result$S1 &
          result$S1 <= result$S1.upper,
          na.rm = TRUE
        ),
        "S1 must satisfy lower <= estimate <= upper"
      )
    )
  )
  #-----------------------------------  POST   --------------------------------#
  
  return(result)
}
```

\newpage
# Calculate Counter-factual Survival Curves Function (Case Base)
```{r}
casebase_csc <- function(
    seed, data.set, model, weights, new.data, time.grid, FSH  
  ) {
  
  #-----------------------------------   PRE   --------------------------------#
  DbC(
    contracts = list(
      contract(
        is.numeric(seed) && length(seed) == 1,
        "seed must be a single numeric value"
      ),
      contract(
        is.data.frame(data.set),
        "data.set must be a data.frame"
      ),
      contract(
        !is.null(model),
        "model must be a valid fitted model"
      ),
      contract(
        is.numeric(weights),
        "weights must be numeric"
      ),
      contract(
        length(weights) == nrow(data.set),
        "weights must have the same length as the number of rows in data.set"
      ),
      contract(
        all(!is.na(weights)),
        "weights must not contain missing values"
      ),
      contract(
        is.data.frame(new.data),
        "new.data must be a data.frame"
      ),
      contract(
        is.numeric(time.grid),
        "time.grid must be numeric"
      ),
      contract(
        length(time.grid) > 0,
        "time.grid must contain at least one point"
      ),
      contract(
        all(diff(time.grid) > 0),
        "time.grid must be strictly increasing"
      ),
      contract(
        is.function(FSH),
        "FSH must be a function"
      )
    )
  )
  #-----------------------------------   PRE   --------------------------------#
  
  surv <- casebase::absoluteRisk(
    object = model,
    time = time.grid,
    newdata = new.data,
    type = "survival"
  )
  
  ## Extract the estimated survival curves for no RHC and RHC, respectively
  S0 <- surv[, 2]
  S1 <- surv[, 3]
  
  # Bootstrap Confidence Intervals for the IPW Case-Base Curves
  
  ## Use nonparametric bootstrap resampling to obtain pointwise confidence
  ## intervals for the Counter-factual survival curves estimated by the IPW
  ## case-base model
  set.seed(seed)
  num.samples <- 200
  
  ## For each bootstrap sample, refit the IPW case-base model and obtain
  ## Counter-factual survival estimates under both treatment levels
  bootstrap <- replicate(num.samples, {
    boot.idx <- sample(nrow(data.set), replace = TRUE)
    boot.data <- data.set[boot.idx, ]
    
    boot.fit <- FSH(
      Delta ~ survival::pspline(time, df = 2) * A,
      data = boot.data,
      ratio = 10,
      weights = weights[boot.idx]
    )
    
    boot.surv <- casebase::absoluteRisk(
      boot.fit,
      time = time.grid,
      newdata = new.data,
      type = "survival"
    )
    
    list(
      S0 = boot.surv[, 2],
      S1 = boot.surv[, 3]
    )
  })
  
  ## Organize the bootstrap estimates into separate matrices for the no-RHC
  ## and RHC Counter-factual survival curves
  S0.boot <- sapply(1:num.samples, function(i) bootstrap[["S0", i]])
  S1.boot <- sapply(1:num.samples, function(i) bootstrap[["S1", i]])
  
  ## Calculate the pointwise lower 95% bootstrap confidence limit for the
  ## no-RHC Counter-factual survival curve
  S0.lower <- apply(
    S0.boot,
    1,
    quantile,
    probs = 0.025,
    na.rm = TRUE
  )
  
  ## Calculate the pointwise upper 95% bootstrap confidence limit for the
  ## no-RHC Counter-factual survival curve
  S0.upper <- apply(
    S0.boot,
    1,
    quantile,
    probs = 0.975,
    na.rm = TRUE
  )
  
  ## Calculate the pointwise lower 95% bootstrap confidence limit for the
  ## RHC Counter-factual survival curve
  S1.lower <- apply(
    S1.boot,
    1,
    quantile,
    probs = 0.025,
    na.rm = TRUE
  )
  
  ## Calculate the pointwise upper 95% bootstrap confidence limit for the
  ## RHC Counter-factual survival curve
  S1.upper <- apply(
    S1.boot,
    1,
    quantile,
    probs = 0.975,
    na.rm = TRUE
  )
  
  result <- list(
    S0 = S0,
    S0.lower = S0.lower, S0.upper = S0.upper,
    S1 = S1,
    S1.lower = S1.lower, S1.upper = S1.upper
  )
  
  #-----------------------------------  POST   --------------------------------#
  DbC(
    contracts = list(
      contract(
        is.numeric(result$S0),
        "S0 must be numeric"
      ),
      contract(
        length(result$S0) == length(time.grid),
        "S0 must have the same length as time.grid"
      ),
      contract(
        is.numeric(result$S0.lower),
        "S0.lower must be numeric"
      ),
      contract(
        length(result$S0.lower) == length(time.grid),
        "S0.lower must have the same length as time.grid"
      ),
      contract(
        is.numeric(result$S0.upper),
        "S0.upper must be numeric"
      ),
      contract(
        length(result$S0.upper) == length(time.grid),
        "S0.upper must have the same length as time.grid"
      ),
      contract(
        is.numeric(result$S1),
        "S1 must be numeric"
      ),
      contract(
        length(result$S1) == length(time.grid),
        "S1 must have the same length as time.grid"
      ),
      contract(
        is.numeric(result$S1.lower),
        "S1.lower must be numeric"
      ),
      contract(
        length(result$S1.lower) == length(time.grid),
        "S1.lower must have the same length as time.grid"
      ),
      contract(
        is.numeric(result$S1.upper),
        "S1.upper must be numeric"
      ),
      contract(
        length(result$S1.upper) == length(time.grid),
        "S1.upper must have the same length as time.grid"
      ),
      contract(
        all(
          result$S0.lower <= result$S0 &
          result$S0 <= result$S0.upper,
          na.rm = TRUE
        ),
        "S0 must satisfy lower <= estimate <= upper"
      ),
      contract(
        all(
          result$S1.lower <= result$S1 &
          result$S1 <= result$S1.upper,
          na.rm = TRUE
        ),
        "S1 must satisfy lower <= estimate <= upper"
      )
    )
  )
  #-----------------------------------  POST   --------------------------------#
  
  return(result)
}
```

\newpage
# Plot Counter-factual Survival Curves
```{r}
plot_csc <- function(
    time.grid,
    S0, S0.lower, S0.upper,
    S1, S1.lower, S1.upper,
    main.title,
    fills, cols
  ) {
  
  #-----------------------------------   PRE   --------------------------------#
  DbC(
    contracts = list(
      contract(
        is.numeric(time.grid),
        "time.grid must be numeric"
      ),
      contract(
        length(time.grid) > 0,
        "time.grid must contain at least one point"
      ),
      contract(
        all(diff(time.grid) > 0),
        "time.grid must be strictly increasing"
      ),
      contract(
        is.numeric(S0),
        "S0 must be numeric"
      ),
      contract(
        length(S0) == length(time.grid),
        "S0 must have the same length as time.grid"
      ),
      contract(
        is.numeric(S0.lower),
        "S0.lower must be numeric"
      ),
      contract(
        length(S0.lower) == length(time.grid),
        "S0.lower must have the same length as time.grid"
      ),
      contract(
        is.numeric(S0.upper),
        "S0.upper must be numeric"
      ),
      contract(
        length(S0.upper) == length(time.grid),
        "S0.upper must have the same length as time.grid"
      ),
      contract(
        is.numeric(S1),
        "S1 must be numeric"
      ),
      contract(
        length(S1) == length(time.grid),
        "S1 must have the same length as time.grid"
      ),
      contract(
        is.numeric(S1.lower),
        "S1.lower must be numeric"
      ),
      contract(
        length(S1.lower) == length(time.grid),
        "S1.lower must have the same length as time.grid"
      ),
      contract(
        is.numeric(S1.upper),
        "S1.upper must be numeric"
      ),
      contract(
        length(S1.upper) == length(time.grid),
        "S1.upper must have the same length as time.grid"
      ),
      contract(
        all(
          S0.lower <= S0 &
          S0 <= S0.upper,
          na.rm = TRUE
        ),
        "S0 must satisfy lower <= estimate <= upper"
      ),
      contract(
        all(
          S1.lower <= S1 &
          S1 <= S1.upper,
          na.rm = TRUE
        ),
        "S1 must satisfy lower <= estimate <= upper"
      ),
      contract(
        is.character(main.title) && length(main.title) == 1,
        "main.title must be a single string"
      ),
      contract(
        is.character(fills),
        "fills must be a character vector"
      ),
      contract(
        length(fills) >= 2,
        "fills must contain at least two elements"
      ),
      contract(
        is.character(cols),
        "cols must be a character vector"
      ),
      contract(
        length(cols) >= 2,
        "cols must contain at least two elements"
      )
    )
  )
  #-----------------------------------   PRE   --------------------------------#
  
  plot(
    NA,
    xlim = range(time.grid),
    ylim = c(0, 1),
    type = "n",
    xlab = "Time",
    ylab = "Survival Probability",
    main = main.title
  )
  
  ## Add the 95% confidence bands
  polygon(
    c(time.grid, rev(time.grid)),
    c(S0.lower, rev(S0.upper)),
    col = fills[1],
    border = NA
  )
  
  polygon(
    c(time.grid, rev(time.grid)),
    c(S1.lower, rev(S1.upper)),
    col = fills[2],
    border = NA
  )
  
  ## Overlay the estimated survival curves. The MS-Cox curve is a step function,
  ## whereas the case-base model produces a smooth survival curve.
  lines(time.grid, S0, type = "s", lwd = 2, col = cols[1], lty = 1)
  lines(time.grid, S1, type = "s", lwd = 2, col = cols[2], lty = 1)
  
  ## Identify the two estimation approaches in the comparison plot
  legend(
    "topright",
    legend = c("No RHC", "RHC"),
    col = cols[1:2],
    lty = c(1, 1),
    lwd = 2,
    bty = "n"
  )
}
```

\newpage
# RMST Comparison: MS-Cox vs. IPW Case-Base
```{r}
rmst_stats <- function(
    models.list, time.grid, new.data,
    cox.rmst.list, casebase.rmst.list,
    rmst.calculation, rmst.diff, casebase.surv
  ) {
  
  #-----------------------------------   PRE   --------------------------------#
  DbC(
    contracts = list(
      contract(
        is.list(models.list),
        "models.list must be a list"
      ),
      contract(
        length(models.list) > 0,
        "models.list must be non-empty"
      ),
      contract(
        all(c("cox", "casebase") %in% names(models.list)),
        "models.list must contain cox and casebase models"
      ),
      contract(
        is.numeric(time.grid),
        "time.grid must be numeric"
      ),
      contract(
        length(time.grid) > 0,
        "time.grid must contain at least one point"
      ),
      contract(
        all(diff(time.grid) > 0),
        "time.grid must be strictly increasing"
      ),
      contract(
        is.data.frame(new.data),
        "new.data must be a data.frame"
      ),
      contract(
        is.function(cox.rmst.list),
        "cox.rmst.list must be a function"
      ),
      contract(
        is.function(casebase.rmst.list),
        "casebase.rmst.list must be a function"
      ),
      contract(
        is.function(rmst.calculation),
        "rmst.calculation must be a function"
      ),
      contract(
        is.function(rmst.diff),
        "rmst.diff must be a function"
      ),
      contract(
        is.function(casebase.surv),
        "casebase.surv must be a function"
      )
    )
  )
  #-----------------------------------   PRE   --------------------------------#
  
  cox.psi.se <- sqrt(diag(vcov(models.list$cox))["A"])
  casebase.var <- sandwich::vcovHC(models.list$casebase, type = "HC0")
  
  cox.rmst.list <- cox.rmst.list(
    ms.cox = models.list$cox,
    time.grid = time.grid,
    cox.psi.se = cox.psi.se,
    survival.curves.data = new.data,
    rmst.calculation = rmst.calculation
  )
  casebase.rmst.list <- casebase.rmst.list(
    casebase.ipw = models.list$casebase,
    casebase.var = casebase.var,
    time.grid = time.grid,
    casebase.surv = casebase.surv,
    rmst.diff = rmst.diff,
    rmst.calculation = rmst.calculation
  )
  
  cox.lower <- cox.rmst.list$rmst - 1.96 * cox.rmst.list$rmst.se
  cox.upper <- cox.rmst.list$rmst + 1.96 * cox.rmst.list$rmst.se
  
  casebase.lower <- casebase.rmst.list$rmst - 1.96 * casebase.rmst.list$rmst.se
  casebase.upper <- casebase.rmst.list$rmst + 1.96 * casebase.rmst.list$rmst.se
  
  result <- list(
    time.grid = time.grid,
    cox.rmst = cox.rmst.list$rmst,
    cox.rmst.se = cox.rmst.list$rmst.se,
    cox.lower = cox.lower,
    cox.upper = cox.upper,
    casebase.rmst = casebase.rmst.list$rmst,
    casebase.rmst.se = casebase.rmst.list$rmst.se,
    casebase.lower = casebase.lower,
    casebase.upper = casebase.upper
  )
  
  #-----------------------------------  POST   --------------------------------#
  DbC(
    contracts = list(
      contract(
        is.numeric(result$time.grid),
        "time.grid must be numeric"
      ),
      contract(
        length(result$time.grid) == length(time.grid),
        "time.grid must have the same length as the input time grid"
      ),
      contract(
        is.numeric(result$cox.rmst),
        "cox.rmst must be numeric"
      ),
      contract(
        length(result$cox.rmst) == length(time.grid),
        "cox.rmst must have the same length as time.grid"
      ),
      contract(
        is.numeric(result$cox.rmst.se),
        "cox.rmst.se must be numeric"
      ),
      contract(
        length(result$cox.rmst.se) == length(time.grid),
        "cox.rmst.se must have the same length as time.grid"
      ),
      contract(
        is.numeric(result$cox.lower),
        "cox.lower must be numeric"
      ),
      contract(
        length(result$cox.lower) == length(time.grid),
        "cox.lower must have the same length as time.grid"
      ),
      contract(
        is.numeric(result$cox.upper),
        "cox.upper must be numeric"
      ),
      contract(
        length(result$cox.upper) == length(time.grid),
        "cox.upper must have the same length as time.grid"
      ),
      contract(
        is.numeric(result$casebase.rmst),
        "casebase.rmst must be numeric"
      ),
      contract(
        length(result$casebase.rmst) == length(time.grid),
        "casebase.rmst must have the same length as time.grid"
      ),
      contract(
        is.numeric(result$casebase.rmst.se),
        "casebase.rmst.se must be numeric"
      ),
      contract(
        length(result$casebase.rmst.se) == length(time.grid),
        "casebase.rmst.se must have the same length as time.grid"
      ),
      contract(
        is.numeric(result$casebase.lower),
        "casebase.lower must be numeric"
      ),
      contract(
        length(result$casebase.lower) == length(time.grid),
        "casebase.lower must have the same length as time.grid"
      ),
      contract(
        is.numeric(result$casebase.upper),
        "casebase.upper must be numeric"
      ),
      contract(
        length(result$casebase.upper) == length(time.grid),
        "casebase.upper must have the same length as time.grid"
      ),
      contract(
        all(
          result$cox.lower <= result$cox.rmst &
          result$cox.rmst <= result$cox.upper,
          na.rm = TRUE
        ),
        "cox RMST must satisfy lower <= estimate <= upper"
      ),
      contract(
        all(
          result$casebase.lower <= result$casebase.rmst &
          result$casebase.rmst <= result$casebase.upper,
          na.rm = TRUE
        ),
        "case-base RMST must satisfy lower <= estimate <= upper"
      )
    )
  )
  #-----------------------------------  POST   --------------------------------#
  
  return(result)
}
```

\newpage
# Plot RMST Comparison
```{r}
plot_rmst_comparison <- function(
    rmst.stats,
    fills, cols,
    additional.title = ""
  ) {
  
  #-----------------------------------   PRE   --------------------------------#
  DbC(
    contracts = list(
      contract(
        is.list(rmst.stats),
        "rmst.stats must be a list"
      ),
      contract(
        is.numeric(rmst.stats$time.grid),
        "time.grid must be numeric"
      ),
      contract(
        length(rmst.stats$time.grid) > 0,
        "time.grid must contain at least one point"
      ),
      contract(
        all(diff(rmst.stats$time.grid) > 0),
        "time.grid must be strictly increasing"
      ),
      contract(
        is.numeric(rmst.stats$cox.rmst),
        "cox.rmst must be numeric"
      ),
      contract(
        is.numeric(rmst.stats$cox.lower),
        "cox.lower must be numeric"
      ),
      contract(
        is.numeric(rmst.stats$cox.upper),
        "cox.upper must be numeric"
      ),
      contract(
        is.numeric(rmst.stats$casebase.rmst),
        "casebase.rmst must be numeric"
      ),
      contract(
        is.numeric(rmst.stats$casebase.lower),
        "casebase.lower must be numeric"
      ),
      contract(
        is.numeric(rmst.stats$casebase.upper),
        "casebase.upper must be numeric"
      ),
      contract(
        length(rmst.stats$cox.rmst) == length(rmst.stats$time.grid),
        "cox.rmst must have the same length as time.grid"
      ),
      contract(
        length(rmst.stats$cox.lower) == length(rmst.stats$time.grid),
        "cox.lower must have the same length as time.grid"
      ),
      contract(
        length(rmst.stats$cox.upper) == length(rmst.stats$time.grid),
        "cox.upper must have the same length as time.grid"
      ),
      contract(
        length(rmst.stats$casebase.rmst) == length(rmst.stats$time.grid),
        "casebase.rmst must have the same length as time.grid"
      ),
      contract(
        length(rmst.stats$casebase.lower) == length(rmst.stats$time.grid),
        "casebase.lower must have the same length as time.grid"
      ),
      contract(
        length(rmst.stats$casebase.upper) == length(rmst.stats$time.grid),
        "casebase.upper must have the same length as time.grid"
      ),
      contract(
        all(
          rmst.stats$cox.lower <= rmst.stats$cox.rmst &
          rmst.stats$cox.rmst <= rmst.stats$cox.upper,
          na.rm = TRUE
        ),
        "ox RMST must satisfy lower <= estimate <= upper"
      ),
      contract(
        all(
          rmst.stats$casebase.lower <= rmst.stats$casebase.rmst &
          rmst.stats$casebase.rmst <= rmst.stats$casebase.upper,
          na.rm = TRUE
        ),
        "case-base RMST must satisfy lower <= estimate <= upper"
      ),
      contract(
        is.character(fills),
        "fills must be a character vector"
      ),
      contract(
        length(fills) >= 2,
        "fills must contain at least two elements"
      ),
      contract(
        is.character(cols),
        "cols must be a character vector"
      ),
      contract(
        length(cols) >= 2,
        "cols must contain at least two elements"
      ),
      contract(
        is.character(additional.title) && length(additional.title) == 1,
        "additional.title must be a single string"
      )
    )
  )
  #-----------------------------------   PRE   --------------------------------#
  
  ylim <- range(
    c(
      rmst.stats$cox.lower, rmst.stats$cox.upper,
      rmst.stats$casebase.lower, rmst.stats$casebase.upper
    ),
    na.rm = TRUE
  )
  
  plot(
    NA,
    xlim = range(rmst.stats$time.grid),
    ylim = ylim,
    type = "n",
    xlab = expression(tau),
    ylab = "RMST Difference",
    main = paste("Causal RMST up to Tau", additional.title)
  )
  
  # 95% Confidence Bands

  ## Add the pointwise 95% confidence band for the MS-Cox RMST contrast
  polygon(
    c(rmst.stats$time.grid, rev(rmst.stats$time.grid)),
    c(rmst.stats$cox.lower, rev(rmst.stats$cox.upper)),
    col = fills[1],
    border = NA
  )
  
  ## Add the pointwise 95% confidence band for the IPW case-base RMST contrast
  polygon(
    c(rmst.stats$time.grid, rev(rmst.stats$time.grid)),
    c(rmst.stats$casebase.lower, rev(rmst.stats$casebase.upper)),
    col = fills[2],
    border = NA
  )
  
  # RMST Contrast Estimates

  ## Overlay the estimated causal RMST contrast from the MS-Cox model
  lines(rmst.stats$time.grid, rmst.stats$cox.rmst, col = cols[1], lwd = 2)
  
  ## Overlay the estimated causal RMST contrast from the IPW case-base model
  lines(rmst.stats$time.grid, rmst.stats$casebase.rmst, col = cols[2], lwd = 2)
  
  ## Add a reference line at an RMST difference of zero, corresponding to equal
  ## restricted mean survival under the two treatment conditions
  abline(h = 0, lty = 3)
  
  # Model Legend
  
  ## Identify the two estimation approaches in the comparison plot
  legend(
    "bottomleft",
    legend = c("MS-Cox", "IPW Case-Base"),
    col = cols[1:2],
    lwd = 2,
    bty = "n"
  )
}
```

\newpage
# Data Preparation
```{r}
## Set random seed for output reproducibility
seed = 123
```

# Construct Analysis Dataset
```{r}
data_set <- construct_data_set()

covariates_vector <- names(data_set)[-(1:4)]
covariates_added <- paste(covariates_vector, collapse = " + ")
```

# Plotting Parameters
```{r}
# two main colors used in different plots
colors <- c("black", "red", "blue", "green")

# lighter shades of colors for confidence bands and overlapped plots
fills <- adjustcolor(colors, alpha.f = 0.2)
dark_fills <- adjustcolor(colors, alpha.f = 0.5)
```

\newpage
# Calculate IPTW, IPCW and Combined Weights
```{r}
# calculate treatment propensity scores
treatment_ps <- propensity_score(
  data.set = data_set,
  outcome = "A",
  covariates = covariates_vector,
  propensity.score.form = propensity_score_form
)
treatment_ps_0 <- treatment_ps[data_set$A == 0]
treatment_ps_1 <- treatment_ps[data_set$A == 1]

# calculate censoring propensity scores
censoring_ps <- propensity_score(
  data.set = data_set,
  outcome = "C",
  covariates = covariates_vector,
  propensity.score.form = propensity_score_form,
  additional = "A"
)
censoring_ps_0 <- censoring_ps[data_set$A == 0]
censoring_ps_1 <- censoring_ps[data_set$A == 1]

# calculate iptw and ipcw
iptw <- stabilized_IPW(
  propensity.score = treatment_ps,
  outcome = data_set$A
)

ipcw <- stabilized_IPW(
  propensity.score = censoring_ps,
  outcome = data_set$C
)

# calculate combined weights
combined_weights <- iptw * ipcw
```

\newpage
# Plot Propensity Scores
```{r}
plot_propensity_scores(
  propensity.score.0 = treatment_ps_0, propensity.score.1 = treatment_ps_1,
  cols = colors[2:3], fills = dark_fills[2:3],
  title = "Treatment PS Overlap"
)
plot_propensity_scores(
  propensity.score.0 = censoring_ps_0, propensity.score.1 = censoring_ps_1,
  cols = colors[2:3], fills = dark_fills[2:3],
  title = "Censoring PS Overlap"
)
```

\newpage
# Plot IPW Distributions
```{r}
plot_ipw(
  ipws = list(iptw, ipcw, combined_weights),
  titles = c("IPTW", "IPCW", "Combined"),
  cols = colors[2:4]
)
```

\newpage
# Descriptive Statistics: Table One
```{r}
# create table one for the unweighted original data set
table_one_original <- tableone::CreateTableOne(
  vars = covariates_vector,
  strata = "A",
  data = data_set,
  test = FALSE
)

print(table_one_original, smd = TRUE)
```

\newpage
# Weighted Table One (Intention-To-Treat)
```{r, warning=FALSE, message=FALSE}
weighted_table_one(
  data.set = data_set,
  weights = iptw,
  covariates = covariates_vector,
  outcome = "A"
)
```

\newpage
# Weighted Table One (Per-Protocol)
```{r, warning=FALSE, message=FALSE}
weighted_table_one(
  data.set = data_set,
  weights = combined_weights,
  covariates = covariates_vector,
  outcome = "A"
)
```

\newpage
# Covariate Balance Before and After Weighting
```{r, warning=FALSE, message=FALSE}
plot_cov_balance(
  data.set = data_set,
  covariates.added = covariates_added,
  outcome = "A",
  weights.list = list(iptw = iptw, combined = combined_weights),
  weights.titles = c("IPTW", "Combined Weights"),
  cols = colors[2:4]
)
```

\newpage
# Case-Base Model Setup
```{r}
## Set the case-base sampling ratio used across all case-base model fits.
## A ratio of 100 samples 100 background person-moments per observed case.
fsh_ratio <- 100
```

# Counter-factual Survival Curve Setup
```{r}
## Define the two treatment levels used to generate Counter-factual survival
## curves: A = 0 represents no RHC and A = 1 represents RHC
treatment_types <- c(0, 1)

## Create a prediction dataset containing both treatment levels for estimation
## of Counter-factual survival under each treatment condition
survival_curves_data <- data.frame(A = treatment_types)

# Time Grid Definition

## Determine the maximum observed follow-up time in the analysis dataset
tau_max <- max(data_set$time, na.rm = TRUE)

## Restrict the primary survival-curve display to the first 150 days
## (approximately four months) to focus on the relevant follow-up period
focus_time <- 150

## Define the time grid used for RMST integration, spanning the full observed
## follow-up period at one-day intervals
tau_grid <- seq(0, tau_max, by = 1)

## Define the time grid used for plotting and estimating survival curves over
## a restricted four-month follow-up period
focused_grid <- seq(0, focus_time, by = 1)
```

\newpage
# Model 1: Unadjusted Marginal Structural Cox Model
```{r, cache=TRUE}
unadjusted_cox <- fit_model(
  fitting.function = survival::coxph,
  formula = Surv(time, Delta) ~ A,
  data.set = data_set,
  seed = seed,
  extract.print.hr = extract_print_hr,
  se.func = vcov,
  robust = TRUE
)
```

\newpage
# Model 2: Unadjusted Case-Base Model
```{r, cache=TRUE, warning=FALSE, message=FALSE}
unadjusted_casebase <- fit_model(
  fitting.function = FSH,
  formula = Delta ~ A,
  data.set = data_set,
  seed = seed,
  extract.print.hr = extract_print_hr,
  se.func = sandwich::vcovHC,
  ratio = fsh_ratio
)
```

\newpage
# Model 3: Covariate-Adjusted Cox Model
```{r, cache=TRUE}
covariate_adjusted_cox <- fit_model(
  fitting.function = survival::coxph,
  formula = as.formula(
    paste("Surv(time, Delta) ~ A +", covariates_added)
  ),
  data.set = data_set,
  seed = seed,
  extract.print.hr = extract_print_hr,
  se.func = vcov,
  robust = TRUE
)
```

\newpage
# Model 4: Covariate-Adjusted Case-Base Model
```{r, cache=TRUE, warning=FALSE, message=FALSE}
covariate_adjusted_casebase <- fit_model(
  fitting.function = FSH,
  formula = as.formula(paste("Delta ~ A +", covariates_added)),
  data.set = data_set,
  seed = seed,
  extract.print.hr = extract_print_hr,
  se.func = sandwich::vcovHC,
  ratio = fsh_ratio
)
```

\newpage
# Model 5: Marginal Structural Cox Model (Intention-To-Treat)
```{r, cache=TRUE}
ms_cox_itt <- fit_model(
  fitting.function = survival::coxph,
  formula = Surv(time, Delta) ~ A,
  data.set = data_set,
  seed = seed,
  extract.print.hr = extract_print_hr,
  se.func = vcov,
  weights = iptw,
  robust = TRUE
)
```

# Model 5: Marginal Structural Cox Model (Per-Protocol)
```{r, cache=TRUE}
ms_cox_pp <- fit_model(
  fitting.function = survival::coxph,
  formula = Surv(time, Delta) ~ A,
  data.set = data_set,
  seed = seed,
  extract.print.hr = extract_print_hr,
  se.func = vcov,
  weights = combined_weights,
  robust = TRUE
)
```

\newpage
# Model 6: IPW Case-Base Proportional Hazards Model (Intention-To-Treat)
```{r, cache=TRUE, warning=FALSE, message=FALSE}
ipw_casebase_itt <- fit_model(
  fitting.function = FSH,
  formula = Delta ~ A,
  data.set = data_set,
  seed = seed,
  extract.print.hr = extract_print_hr,
  se.func = sandwich::vcovHC,
  weights = iptw,
  ratio = fsh_ratio
)
```

# Model 6: IPW Case-Base Proportional Hazards Model (Per-Protocol)
```{r, cache=TRUE, warning=FALSE, message=FALSE}
ipw_casebase_pp <- fit_model(
  fitting.function = FSH,
  formula = Delta ~ A,
  data.set = data_set,
  seed = seed,
  extract.print.hr = extract_print_hr,
  se.func = sandwich::vcovHC,
  weights = combined_weights,
  ratio = fsh_ratio
)
```

\newpage
# Model 7: Unadjusted Case-Base Model with Time-Varying Hazard Ratio
```{r, cache=TRUE, warning=FALSE, message=FALSE}
unadjusted_var_hr_casebase <- fit_model(
  fitting.function = FSH,
  formula = Delta ~ survival::pspline(time, df = 2) * A,
  data.set = data_set,
  seed = seed,
  extract.print.hr = extract_print_hr,
  constant.HR = FALSE,
  ratio = fsh_ratio
)
```

\newpage
# Time-Varying Hazard Ratio
```{r, message=FALSE}
plot_time_dep_hr(
  list(unadjusted_var_hr_casebase),
  main.title = "Unadjusted Case Base: Variable Hazard Ratio",
  time.grid = tau_grid,
  casebase.var.hr = casebase_var_hr
)
```

\newpage
# Model 8: IPW Case-Base Model with Time-Varying Hazard Ratio (Intention-To-Treat)
```{r, cache=TRUE, warning=FALSE, message=FALSE}
ipw_var_hr_casebase_itt <- fit_model(
  fitting.function = FSH,
  formula = Delta ~ survival::pspline(time, df = 2) * A,
  data.set = data_set,
  seed = seed,
  extract.print.hr = extract_print_hr,
  constant.HR = FALSE,
  weights = iptw,
  ratio = fsh_ratio
)
```

# Model 8: IPW Case-Base Model with Time-Varying Hazard Ratio (Per-Protocol)
```{r, cache=TRUE, warning=FALSE, message=FALSE}
ipw_var_hr_casebase_pp <- fit_model(
  fitting.function = FSH,
  formula = Delta ~ survival::pspline(time, df = 2) * A,
  data.set = data_set,
  seed = seed,
  extract.print.hr = extract_print_hr,
  constant.HR = FALSE,
  weights = combined_weights,
  ratio = fsh_ratio
)
```

\newpage
# Time-Varying Hazard Ratio
```{r, message=FALSE}
plot_time_dep_hr(
  list(ipw_var_hr_casebase_itt, ipw_var_hr_casebase_pp),
  main.title = "
    IPW Case Base: Variable Hazard Ratio (ITT vs. PP)
  ",
  time.grid = tau_grid,
  casebase.var.hr = casebase_var_hr,
  titles = c("Intention-To-Treat", "Per-Protocol"),
  cols = colors[2:3]
)
```

\newpage
# Head-to-Head Comparison: Models 7 and 8
```{r}
plot_time_dep_hr(
  list(
    unadjusted_var_hr_casebase, ipw_var_hr_casebase_itt, ipw_var_hr_casebase_pp
  ),
  main.title = "Case Base: Variable Hazard Ratio",
  time.grid = tau_grid,
  casebase.var.hr = casebase_var_hr,
  titles = c("Unadjusted","Adjusted-Intention-To-Treat", "Adjusted-Per-Protocol"),
  cols = colors[2:4]
)
```

\newpage
# MS-Cox Counter-factual Survival Curves
```{r, cache=TRUE}
cox_csc_itt <- cox_csc(
  model = ms_cox_itt,
  new.data = survival_curves_data,
  time.grid = tau_grid
)

cox_csc_pp <- cox_csc(
  model = ms_cox_pp,
  new.data = survival_curves_data,
  time.grid = tau_grid
)
```

# IPW Case-Base Counter-factual Survival Curves
```{r, cache=TRUE, message=FALSE, warning=FALSE}
casebase_csc_itt <- casebase_csc(
  seed = seed,
  data.set = data_set,
  model = ipw_var_hr_casebase_itt,
  weights = iptw,
  new.data = survival_curves_data,
  time.grid = tau_grid,
  FSH = FSH
)

casebase_csc_pp <- casebase_csc(
  seed = seed,
  data.set = data_set,
  model = ipw_var_hr_casebase_pp,
  weights = combined_weights,
  new.data = survival_curves_data,
  time.grid = tau_grid,
  FSH = FSH
)
```

\newpage
# Head-to-Head Comparison of Counter-factual Survival Curves (Intention-To-Treat)
```{r}
# Counter-factual Survival Curves (MS-Cox)
plot_csc(
  time.grid = tau_grid,
  S0 = cox_csc_itt$S0,
  S0.lower = cox_csc_itt$S0.lower,
  S0.upper = cox_csc_itt$S0.upper,
  S1 = cox_csc_itt$S1,
  S1.lower = cox_csc_itt$S1.lower,
  S1.upper = cox_csc_itt$S1.upper,
  main.title = "Counter-factual Survival Curves (MS-Cox)",
  fills = fills[2:3], cols = colors[2:3]
)

# Counter-factual Survival Curves (Case Base)
plot_csc(
  time.grid = tau_grid,
  S0 = casebase_csc_itt$S0,
  S0.lower = casebase_csc_itt$S0.lower,
  S0.upper = casebase_csc_itt$S0.upper,
  S1 = casebase_csc_itt$S1,
  S1.lower = casebase_csc_itt$S1.lower,
  S1.upper = casebase_csc_itt$S1.upper,
  main.title = "Counter-factual Survival Curves (Case Base)",
  fills = fills[2:3], cols = colors[2:3]
)
```

\newpage
# Head-to-Head Comparison of Counter-factual Survival Curves (Per-Protocol)
```{r}
# Counter-factual Survival Curves (MS-Cox)
plot_csc(
  time.grid = tau_grid,
  S0 = cox_csc_pp$S0,
  S0.lower = cox_csc_pp$S0.lower,
  S0.upper = cox_csc_pp$S0.upper,
  S1 = cox_csc_pp$S1,
  S1.lower = cox_csc_pp$S1.lower,
  S1.upper = cox_csc_pp$S1.upper,
  main.title = "Counter-factual Survival Curves (MS-Cox)",
  fills = fills[2:3], cols = colors[2:3]
)

# Counter-factual Survival Curves (Case Base)
plot_csc(
  time.grid = tau_grid,
  S0 = casebase_csc_pp$S0,
  S0.lower = casebase_csc_pp$S0.lower,
  S0.upper = casebase_csc_pp$S0.upper,
  S1 = casebase_csc_pp$S1,
  S1.lower = casebase_csc_pp$S1.lower,
  S1.upper = casebase_csc_pp$S1.upper,
  main.title = "Counter-factual Survival Curves (Case Base)",
  fills = fills[2:3], cols = colors[2:3]
)
```

\newpage
# RMST Estimates and Standard Errors
```{r, cache=TRUE}
rmst_stats_itt <- rmst_stats(
  models.list = list(
    cox = ms_cox_itt,
    casebase = ipw_var_hr_casebase_itt
  ),
  time.grid = tau_grid,
  new.data = survival_curves_data,
  cox.rmst.list = cox_rmst_list,
  casebase.rmst.list = casebase_rmst_list,
  rmst.calculation = rmst_calculation,
  rmst.diff = rmst_diff,
  casebase.surv = casebase_surv
)
```

```{r}
rmst_stats_pp <- rmst_stats(
  models.list = list(
    cox = ms_cox_pp,
    casebase = ipw_var_hr_casebase_pp
  ),
  time.grid = tau_grid,
  new.data = survival_curves_data,
  cox.rmst.list = cox_rmst_list,
  casebase.rmst.list = casebase_rmst_list,
  rmst.calculation = rmst_calculation,
  rmst.diff = rmst_diff,
  casebase.surv = casebase_surv
)
```

\newpage
# RMST Comparison Plots
```{r}
plot_rmst_comparison(
  rmst.stats = rmst_stats_itt,
  fills = fills[2:3], cols = colors[2:3],
  additional.title = "(Intention-To-Treat)"
)

plot_rmst_comparison(
  rmst.stats = rmst_stats_pp,
  fills = fills[2:3], cols = colors[2:3],
  additional.title = "(Per-Protocol)"
)
```