---
title: "Causal RMST under case-base sampling with IPW: simulation study"
author: "Bilal Shaikh and Sumeet Kalia"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = FALSE, warning = FALSE, message = FALSE)
```

# Libraries

```{r}
library("survival")
library("casebase")
library("sandwich")
library("splines")
library("foreach")
library("doParallel")
library("doRNG")
library("ggplot2")
library("grid")
library("xtable")
```

# Global configuration

```{r}
SEED <- 20260811
N_REPLICATES <- 2000
N_REFERENCE <- 1e7
CB_RATIO <- 20
NU <- 1
TAU_GRID <- seq(0, 3, by = 0.02)
TAU_REPORT <- c(0.5, 1, 2, 3)
N_CORES <- max(1, parallel::detectCores() - 1)
ALPHA <- 0.05
Z_CRIT <- qnorm(1 - ALPHA / 2)

GAMMA_RAW <- c(0.7, -0.4, 0.5, 0.2, 0.6, -0.3, 0.1, -0.2, 0.4, -0.5)
BETA_RAW <- c(0.6, -0.3, 0.4, 0.1, 0.5, -0.2, 0.3, -0.1, 0.4, -0.4)

GAMMA_VEC <- GAMMA_RAW / sqrt(sum(GAMMA_RAW^2))
BETA_VEC <- BETA_RAW / sqrt(sum(BETA_RAW^2))

SCENARIOS <- data.frame(
  scenario = paste0("S", 1:8),
  label = c("Null treatment effect", "Moderate treatment effect",
            "Reference configuration", "Reduced sample size",
            "Small sample size", "Heavy censoring",
            "Strong confounding", "Weak prognostic covariate effect"),
  beta.a = c(0, log(1.5), log(2), log(2), log(2), log(2), log(2), log(2)),
  n = c(1000, 1000, 1000, 500, 250, 1000, 1000, 1000),
  c.max = c(3.0, 3.0, 3.0, 3.0, 3.0, 1.5, 3.0, 3.0),
  c.gamma = c(1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 2.0, 1.0),
  c.beta = c(1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 0.5),
  stringsAsFactors = FALSE
)

SCENARIOS
```

# Data-generating mechanism

Covariates are standardized to zero mean and unit variance so that `c.gamma` is
the standard deviation of the treatment linear predictor and `c.beta` scales the
prognostic index.

```{r}
standardize <- function(x, mu, sigma) (x - mu) / sigma

draw_covariates <- function(n) {
  x1 <- rnorm(n)
  x2 <- rnorm(n)
  x3 <- standardize(runif(n), 0.5, sqrt(1 / 12))
  x4 <- standardize(runif(n), 0.5, sqrt(1 / 12))
  m5 <- exp(0.125)
  s5 <- sqrt((exp(0.25) - 1) * exp(0.25))
  x5 <- standardize(rlnorm(n, meanlog = 0, sdlog = 0.5), m5, s5)
  x6 <- standardize(rgamma(n, shape = 2, rate = 1), 2, sqrt(2))
  x7 <- standardize(rexp(n, rate = 1), 1, 1)
  x8 <- standardize(rbinom(n, 1, 0.5), 0.5, sqrt(0.25))
  x9 <- standardize(rbinom(n, 1, 0.3), 0.3, sqrt(0.21))
  x10 <- standardize(rpois(n, lambda = 1.5), 1.5, sqrt(1.5))
  cbind(X1 = x1, X2 = x2, X3 = x3, X4 = x4, X5 = x5,
        X6 = x6, X7 = x7, X8 = x8, X9 = x9, X10 = x10)
}

generate_data <- function(n, beta.a, c.max, c.gamma, c.beta) {
  x.mat <- draw_covariates(n)
  lp.a <- c.gamma * as.vector(x.mat %*% GAMMA_VEC)
  e.score <- plogis(lp.a)
  a.vec <- rbinom(n, 1, e.score)
  lp.y <- c.beta * as.vector(x.mat %*% BETA_VEC) + beta.a * a.vec
  u.unif <- runif(n)
  t.event <- (-log(u.unif) / exp(lp.y))^(1 / NU)
  t.cens <- runif(n, 0, c.max)
  data.frame(time = pmin(t.event, t.cens),
             status = as.integer(t.event <= t.cens),
             A = a.vec,
             x.mat)
}
```

# Reference values

The counterfactual survival function is a covariate mixture and is evaluated by
Monte Carlo integration over the covariate distribution alone. No event times,
no censoring and no working model enter this computation.

```{r}
rmst_from_survival <- function(t.grid, s.vec) {
  d.t <- diff(t.grid)
  mids <- (head(s.vec, -1) + tail(s.vec, -1)) / 2
  c(0, cumsum(mids * d.t))
}

reference_values <- function(beta.a, c.beta, n.ref = N_REFERENCE, chunk = 1e5) {
  n.done <- 0
  next.report <- 0.1
  s0.sum <- rep(0, length(TAU_GRID))
  s1.sum <- rep(0, length(TAU_GRID))
  h0.num <- rep(0, length(TAU_GRID))
  h1.num <- rep(0, length(TAU_GRID))
  while (n.done < n.ref) {
    m <- min(chunk, n.ref - n.done)
    x.mat <- draw_covariates(m)
    eta <- c.beta * as.vector(x.mat %*% BETA_VEC)
    base.haz <- outer(TAU_GRID^NU, rep(1, m))
    lam0 <- exp(eta)
    lam1 <- exp(eta + beta.a)
    s0.mat <- exp(-base.haz * rep(lam0, each = length(TAU_GRID)))
    s1.mat <- exp(-base.haz * rep(lam1, each = length(TAU_GRID)))
    s0.sum <- s0.sum + rowSums(s0.mat)
    s1.sum <- s1.sum + rowSums(s1.mat)
    h0.num <- h0.num + rowSums(s0.mat * rep(lam0, each = length(TAU_GRID)))
    h1.num <- h1.num + rowSums(s1.mat * rep(lam1, each = length(TAU_GRID)))
    n.done <- n.done + m
    if (n.done / n.ref >= next.report) {
      cat(sprintf("  reference integration: %3.0f%%\n", 100 * n.done / n.ref))
      flush.console()
      next.report <- next.report + 0.1
    }
  }
  s0 <- s0.sum / n.ref
  s1 <- s1.sum / n.ref
  haz0 <- NU * TAU_GRID^(NU - 1) * (h0.num / n.ref) / pmax(s0, .Machine$double.eps)
  haz1 <- NU * TAU_GRID^(NU - 1) * (h1.num / n.ref) / pmax(s1, .Machine$double.eps)
  list(t.grid = TAU_GRID,
       s0 = s0,
       s1 = s1,
       rmst.diff = rmst_from_survival(TAU_GRID, s1) - rmst_from_survival(TAU_GRID, s0),
       hr.t = haz1 / haz0)
}
```

The constant-effect estimators are additionally compared against the large-sample
limit of the corresponding fitting criterion under an exogenous-treatment version
of the same mechanism. This limit is a projection attached to the criterion and
not a structural parameter.

```{r}
PROJ_N <- 5e5
PROJ_RATIO <- 5
proj.cache <- new.env(parent = emptyenv())

projection_limits <- function(beta.a, c.beta, c.max) {
  key <- paste(beta.a, c.beta, c.max, sep = "_")
  if (!is.null(proj.cache[[key]])) return(proj.cache[[key]])
  set.seed(SEED + 991)
  d.proj <- generate_data(PROJ_N, beta.a, c.max, c.gamma = 0, c.beta = c.beta)
  w.unit <- rep(1, nrow(d.proj))
  cox.fit <- coxph(Surv(time, status) ~ A, data = d.proj)
  cb.fit <- fit_casebase(d.proj, w.unit, basis = "spline", ratio = PROJ_RATIO)
  cox.ref <- cox_rmst(cox.fit)
  out <- list(
    psi = c(cox = unname(coef(cox.fit)["A"]),
            casebase = unname(coef(cb.fit$fit)["A"])),
    rmst.cox = cox.ref$rmst
  )
  proj.cache[[key]] <- out
  out
}
```

# Weighted case-base fitting

`sampleCaseBase` draws base moments proportionally to observed follow-up time.
Observation weights are attached before sampling so that each sampled moment
inherits the weight of the individual from which it was drawn, which makes the
weighted score equation target the weighted hazard.

```{r}
interior_knots <- function(event.times, bnd, probs = c(0.25, 0.5, 0.75)) {
  if (length(event.times) < 20) return(numeric(0))
  k <- unique(as.numeric(quantile(event.times, probs = probs, names = FALSE)))
  k <- k[k > bnd[1] + .Machine$double.eps & k < bnd[2] - .Machine$double.eps]
  sort(unique(k))
}

fit_casebase <- function(data.set, weights, basis = c("spline", "gompertz", "exponential"),
                         time.varying = FALSE, ratio = CB_RATIO, seed = NULL) {
  basis <- match.arg(basis)
  d.aug <- data.set
  d.aug$w <- weights
  if (!is.null(seed)) set.seed(seed)
  cb.data <- sampleCaseBase(d.aug, time = "time", event = "status", ratio = ratio)
  bnd <- range(cb.data$time)
  knot.q <- interior_knots(data.set$time[data.set$status == 1], bnd)
  time.term <- switch(
    basis,
    exponential = NULL,
    gompertz = "time",
    spline = if (length(knot.q) > 0) {
      sprintf("bs(time, knots = c(%s), Boundary.knots = c(%s), degree = 3)",
              paste(knot.q, collapse = ", "), paste(bnd, collapse = ", "))
    } else {
      sprintf("bs(time, df = 3, Boundary.knots = c(%s), degree = 3)",
              paste(bnd, collapse = ", "))
    }
  )
  rhs <- if (is.null(time.term)) {
    "A"
  } else if (time.varying) {
    paste0("A * ", time.term)
  } else {
    paste0("A + ", time.term)
  }
  form <- as.formula(paste("status ~", rhs, "+ offset(offset)"))
  fit <- glm(form, data = cb.data, family = binomial, weights = w)
  list(fit = fit,
       vcov = vcovHC(fit, type = "HC0"),
       knots = knot.q,
       bnd = bnd,
       basis = basis,
       time.varying = time.varying)
}

casebase_hazard_grid <- function(cb.obj, a.value, t.grid) {
  nd <- data.frame(time = pmax(t.grid, .Machine$double.eps), A = a.value, offset = 0)
  x.grid <- model.matrix(delete.response(terms(cb.obj$fit)), data = nd)
  x.grid <- x.grid[, names(coef(cb.obj$fit)), drop = FALSE]
  list(x = x.grid, hazard = as.vector(exp(x.grid %*% coef(cb.obj$fit))))
}
```

# RMST contrast and delta-method standard error

For the case-base model the gradient is available in closed form. Writing
$\lambda^{(a)}(u) = \exp\{z(u,a)^\top\beta\}$, the gradient of the survival
function is $-S^{(a)}(t)\int_0^t \lambda^{(a)}(u) z(u,a)\,du$, and the RMST
gradient integrates this over the horizon.

```{r}
casebase_rmst <- function(cb.obj, t.grid = TAU_GRID) {
  d.t <- diff(t.grid)
  grad.list <- list()
  surv.list <- list()
  for (a.value in c(0, 1)) {
    g <- casebase_hazard_grid(cb.obj, a.value, t.grid)
    lam <- g$hazard
    cum.haz <- c(0, cumsum(d.t * (head(lam, -1) + tail(lam, -1)) / 2))
    s.vec <- exp(-cum.haz)
    lam.x <- g$x * lam
    d.cum <- rbind(0, apply(lam.x, 2, function(col) {
      cumsum(d.t * (head(col, -1) + tail(col, -1)) / 2)
    }))
    d.surv <- -s.vec * d.cum
    d.rmst <- rbind(0, apply(d.surv, 2, function(col) {
      cumsum(d.t * (head(col, -1) + tail(col, -1)) / 2)
    }))
    surv.list[[as.character(a.value)]] <- s.vec
    grad.list[[as.character(a.value)]] <- d.rmst
  }
  rmst.diff <- rmst_from_survival(t.grid, surv.list[["1"]]) -
    rmst_from_survival(t.grid, surv.list[["0"]])
  grad.diff <- grad.list[["1"]] - grad.list[["0"]]
  v.mat <- cb.obj$vcov[colnames(grad.diff), colnames(grad.diff), drop = FALSE]
  se.vec <- sqrt(pmax(rowSums((grad.diff %*% v.mat) * grad.diff), 0))
  list(t.grid = t.grid, s0 = surv.list[["0"]], s1 = surv.list[["1"]],
       rmst = rmst.diff, se = se.vec)
}
```

For the marginal structural Cox model the baseline is held fixed and only the
treatment coefficient contributes, which gives
$\partial S^{(1)}/\partial\psi = S^{(1)}(t)\log S^{(1)}(t)$.

```{r}
cox_rmst <- function(cox.fit, t.grid = TAU_GRID) {
  sf <- survfit(cox.fit, newdata = data.frame(A = c(0, 1)))
  s0 <- approx(sf$time, sf$surv[, 1], t.grid, method = "constant", rule = 2)$y
  s1 <- approx(sf$time, sf$surv[, 2], t.grid, method = "constant", rule = 2)$y
  se.psi <- sqrt(diag(vcov(cox.fit))["A"])
  integrand <- ifelse(s1 > 0, s1 * log(pmax(s1, .Machine$double.eps)), 0)
  g.vec <- rmst_from_survival(t.grid, integrand)
  list(t.grid = t.grid, s0 = s0, s1 = s1,
       rmst = rmst_from_survival(t.grid, s1) - rmst_from_survival(t.grid, s0),
       se = abs(g.vec) * as.numeric(se.psi))
}
```

# One replicate

```{r}
run_replicate <- function(n, beta.a, c.max, c.gamma, c.beta) {
  d.set <- generate_data(n, beta.a, c.max, c.gamma, c.beta)
  ps.fit <- glm(A ~ X1 + X2 + X3 + X4 + X5 + X6 + X7 + X8 + X9 + X10,
                data = d.set, family = binomial)
  p.score <- pmin(pmax(fitted(ps.fit), 1e-8), 1 - 1e-8)
  p.marg <- mean(d.set$A)
  ipw <- ifelse(d.set$A == 1, p.marg / p.score, (1 - p.marg) / (1 - p.score))

  out <- list(ipw.max = max(ipw),
              ipw.mean = mean(ipw),
              ess = sum(ipw)^2 / sum(ipw^2) / n,
              event.rate = mean(d.set$status))

  m1 <- try(coxph(Surv(time, status) ~ A, data = d.set,
                  weights = ipw, robust = TRUE), silent = TRUE)
  if (!inherits(m1, "try-error")) {
    r1 <- cox_rmst(m1)
    out$M1.psi <- unname(coef(m1)["A"])
    out$M1.se <- unname(sqrt(diag(vcov(m1))["A"]))
    out$M1.rmst <- r1$rmst
    out$M1.rmst.se <- r1$se
  }

  for (spec in c("M2", "M3", "M4")) {
    basis <- switch(spec, M2 = "exponential", M3 = "gompertz", M4 = "spline")
    cb <- try(fit_casebase(d.set, ipw, basis = basis), silent = TRUE)
    if (!inherits(cb, "try-error") && cb$fit$converged) {
      rr <- casebase_rmst(cb)
      out[[paste0(spec, ".psi")]] <- unname(coef(cb$fit)["A"])
      out[[paste0(spec, ".se")]] <- unname(sqrt(diag(cb$vcov)["A"]))
      out[[paste0(spec, ".rmst")]] <- rr$rmst
      out[[paste0(spec, ".rmst.se")]] <- rr$se
    }
  }

  m5 <- try(coxph(Surv(time, status) ~ A + tt(A), data = d.set,
                  weights = ipw, robust = TRUE,
                  tt = function(x, t, ...) x * log(pmax(t, 1e-8))), silent = TRUE)
  if (!inherits(m5, "try-error")) {
    b5 <- coef(m5)
    v5 <- vcov(m5)
    log.t <- log(pmax(TAU_GRID, 1e-8))
    out$M5.loghr <- b5[1] + b5[2] * log.t
    out$M5.loghr.se <- sqrt(v5[1, 1] + log.t^2 * v5[2, 2] + 2 * log.t * v5[1, 2])
  }

  m6 <- try(fit_casebase(d.set, ipw, basis = "spline", time.varying = TRUE),
            silent = TRUE)
  if (!inherits(m6, "try-error") && m6$fit$converged) {
    g1 <- casebase_hazard_grid(m6, 1, TAU_GRID)$x
    g0 <- casebase_hazard_grid(m6, 0, TAU_GRID)$x
    d.x <- g1 - g0
    out$M6.loghr <- as.vector(d.x %*% coef(m6$fit))
    out$M6.loghr.se <- sqrt(pmax(rowSums((d.x %*% m6$vcov) * d.x), 0))
  }

  out
}
```

# Running the study

```{r}
dir.create("Figures/Simulation", recursive = TRUE, showWarnings = FALSE)

cl <- makeCluster(N_CORES)
registerDoParallel(cl)
registerDoRNG(SEED)

results.all <- list()

for (s.index in seq_len(nrow(SCENARIOS))) {
  s.row <- SCENARIOS[s.index, ]
  cat(sprintf("\n===== Scenario %s: %s =====\n", s.row$scenario, s.row$label))
  cat("computing reference values\n")
  ref <- reference_values(s.row$beta.a, s.row$c.beta)
  proj <- projection_limits(s.row$beta.a, s.row$c.beta, s.row$c.max)

  chunk.size <- ceiling(N_REPLICATES / 20)
  chunk.start <- seq(1, N_REPLICATES, by = chunk.size)
  reps <- list()
  for (k in seq_along(chunk.start)) {
    n.this <- min(chunk.size, N_REPLICATES - chunk.start[k] + 1)
    reps[[k]] <- foreach(
      i = seq_len(n.this),
      .packages = c("survival", "casebase", "sandwich", "splines"),
      .export = c("generate_data", "draw_covariates", "standardize",
                  "fit_casebase", "casebase_hazard_grid", "casebase_rmst",
                  "cox_rmst", "rmst_from_survival", "run_replicate",
                  "interior_knots", "GAMMA_VEC", "BETA_VEC", "NU",
                  "TAU_GRID", "CB_RATIO"),
      .errorhandling = "pass"
    ) %dopar% {
      run_replicate(s.row$n, s.row$beta.a, s.row$c.max, s.row$c.gamma, s.row$c.beta)
    }
    cat(sprintf("  %s: %3.0f%% complete\n", s.row$scenario,
                100 * min(chunk.start[k] + n.this - 1, N_REPLICATES) / N_REPLICATES))
    flush.console()
  }
  results.all[[s.row$scenario]] <- list(
    scenario = s.row,
    reference = ref,
    projection = proj,
    replicates = do.call(c, reps)
  )
}

stopCluster(cl)
saveRDS(results.all, file = "simulation-results.rds")
```

# Performance measures

```{r}
summarize_scalar <- function(est, se, truth) {
  ok <- is.finite(est) & is.finite(se)
  est <- est[ok]
  se <- se[ok]
  r <- length(est)
  bias <- mean(est) - truth
  mcsd <- sd(est)
  lower <- est - Z_CRIT * se
  upper <- est + Z_CRIT * se
  data.frame(
    n.used = r,
    truth = truth,
    mean.est = mean(est),
    bias = bias,
    bias.mcse = mcsd / sqrt(r),
    mcsd = mcsd,
    ase = mean(se),
    mse = mean((est - truth)^2),
    coverage = mean(lower <= truth & truth <= upper),
    coverage.mcse = sqrt(mean(lower <= truth & truth <= upper) *
                           (1 - mean(lower <= truth & truth <= upper)) / r),
    reject = mean(abs(est / se) > Z_CRIT)
  )
}

extract_grid <- function(reps, field, index) {
  vapply(reps, function(z) {
    v <- z[[field]]
    if (is.null(v) || length(v) < index) NA_real_ else v[index]
  }, numeric(1))
}

tau.index <- match(TAU_REPORT, TAU_GRID)

rmst.summary <- do.call(rbind, lapply(names(results.all), function(s.name) {
  obj <- results.all[[s.name]]
  reps <- Filter(function(z) is.list(z) && !inherits(z, "error"), obj$replicates)
  do.call(rbind, lapply(c("M1", "M2", "M3", "M4"), function(m) {
    do.call(rbind, lapply(seq_along(TAU_REPORT), function(j) {
      idx <- tau.index[j]
      est <- extract_grid(reps, paste0(m, ".rmst"), idx)
      se <- extract_grid(reps, paste0(m, ".rmst.se"), idx)
      cbind(data.frame(scenario = s.name, method = m, tau = TAU_REPORT[j]),
            summarize_scalar(est, se, obj$reference$rmst.diff[idx]))
    }))
  }))
}))

hr.summary <- do.call(rbind, lapply(names(results.all), function(s.name) {
  obj <- results.all[[s.name]]
  reps <- Filter(function(z) is.list(z) && !inherits(z, "error"), obj$replicates)
  do.call(rbind, lapply(c("M1", "M2", "M3", "M4"), function(m) {
    est <- vapply(reps, function(z) {
      v <- z[[paste0(m, ".psi")]]; if (is.null(v)) NA_real_ else v
    }, numeric(1))
    se <- vapply(reps, function(z) {
      v <- z[[paste0(m, ".se")]]; if (is.null(v)) NA_real_ else v
    }, numeric(1))
    pj <- if (is.list(obj$projection)) obj$projection$psi else obj$projection
    truth <- if (m == "M1") pj["cox"] else pj["casebase"]
    cbind(data.frame(scenario = s.name, method = m), summarize_scalar(est, se, truth))
  }))
}))

convergence.summary <- do.call(rbind, lapply(names(results.all), function(s.name) {
  reps <- results.all[[s.name]]$replicates
  errors <- sum(vapply(reps, function(z) inherits(z, "error"), logical(1)))
  ok <- Filter(function(z) is.list(z) && !inherits(z, "error"), reps)
  data.frame(
    scenario = s.name,
    replicates = length(reps),
    hard.failures = errors,
    M1.missing = sum(vapply(ok, function(z) is.null(z$M1.psi), logical(1))),
    M2.missing = sum(vapply(ok, function(z) is.null(z$M2.psi), logical(1))),
    M3.missing = sum(vapply(ok, function(z) is.null(z$M3.psi), logical(1))),
    M4.missing = sum(vapply(ok, function(z) is.null(z$M4.psi), logical(1))),
    mean.ipw = mean(vapply(ok, function(z) z$ipw.mean, numeric(1))),
    max.ipw = max(vapply(ok, function(z) z$ipw.max, numeric(1))),
    ess = mean(vapply(ok, function(z) z$ess, numeric(1))),
    event.rate = mean(vapply(ok, function(z) z$event.rate, numeric(1)))
  )
}))

rmst.summary
hr.summary
convergence.summary
```

# LaTeX tables

```{r}
print(xtable(rmst.summary[, c("scenario", "method", "tau", "truth", "mean.est",
                              "bias", "mcsd", "ase", "mse", "coverage")],
             digits = c(0, 0, 0, 2, 4, 4, 4, 4, 4, 4, 3)),
      include.rownames = FALSE, file = "table-sim-rmst.tex")

print(xtable(hr.summary[, c("scenario", "method", "truth", "mean.est", "bias",
                            "mcsd", "ase", "mse", "coverage", "reject")],
             digits = c(0, 0, 0, 4, 4, 4, 4, 4, 4, 3, 3)),
      include.rownames = FALSE, file = "table-sim-hr.tex")

print(xtable(convergence.summary, digits = c(0, 0, 0, 0, 0, 0, 0, 0, 3, 2, 3, 3)),
      include.rownames = FALSE, file = "table-sim-convergence.tex")
```

# Reference values: integration against a model-based alternative

This section quantifies what is lost by defining the reference value through a
fitted working model instead of by direct integration. It does not depend on the
replicates, only on the data-generating mechanism, so it can be run on its own
and is unaffected by objects left over from an earlier run.

```{r}
ref.cache <- new.env(parent = emptyenv())

cox_reference_rmst <- function(beta.a, c.beta, c.max, n.proj = PROJ_N) {
  key <- paste(beta.a, c.beta, c.max, sep = "_")
  if (!is.null(ref.cache[[key]])) return(ref.cache[[key]])
  set.seed(SEED + 991)
  d.proj <- generate_data(n.proj, beta.a, c.max, c.gamma = 0, c.beta = c.beta)
  cox.fit <- coxph(Surv(time, status) ~ A, data = d.proj)
  out <- cox_rmst(cox.fit)$rmst
  ref.cache[[key]] <- out
  out
}

integration_reference <- function(s.name, beta.a, c.beta) {
  obj <- if (exists("results.all")) results.all[[s.name]] else NULL
  if (!is.null(obj) && is.list(obj$reference)) return(obj$reference$rmst.diff)
  reference_values(beta.a, c.beta)$rmst.diff
}

reference.compare <- do.call(rbind, lapply(seq_len(nrow(SCENARIOS)), function(k) {
  s.row <- SCENARIOS[k, ]
  cat(sprintf("reference comparison: %s\n", s.row$scenario))
  flush.console()
  exact.curve <- integration_reference(s.row$scenario, s.row$beta.a, s.row$c.beta)
  model.curve <- cox_reference_rmst(s.row$beta.a, s.row$c.beta, s.row$c.max)
  do.call(rbind, lapply(c(1, 3), function(tau) {
    idx <- match(tau, TAU_GRID)
    exact <- exact.curve[idx]
    modelled <- model.curve[idx]
    data.frame(scenario = s.row$scenario, tau = tau,
               integration = exact, cox.working = modelled,
               difference = modelled - exact,
               pct.of.truth = ifelse(abs(exact) < 1e-10, NA_real_,
                                     100 * (modelled - exact) / abs(exact)))
  }))
}))
reference.compare

print(xtable(reference.compare, digits = c(0, 0, 0, 4, 4, 4, 1)),
      include.rownames = FALSE, file = "table-sim-reference.tex")
```

# Figures

```{r}
PAL <- c("red", "blue", "green4", "black")
LTY <- c("solid", "dashed", "dotdash", "dotted")
SHP <- c(16, 17, 15, 3)

theme_pub <- function(base_size = 10) {
  theme_minimal(base_size = base_size) +
    theme(
      panel.grid.minor = element_blank(),
      panel.grid.major = element_line(colour = "grey90", linewidth = 0.3),
      panel.border = element_rect(colour = "grey30", fill = NA, linewidth = 0.4),
      strip.background = element_blank(),
      strip.text = element_text(face = "bold", size = base_size),
      axis.text = element_text(colour = "grey20"),
      legend.position = "bottom",
      legend.direction = "horizontal",
      legend.title = element_blank(),
      legend.key.width = unit(1.4, "lines")
    )
}

ref.s3 <- results.all[["S3"]]$reference
reps.s3 <- Filter(function(z) is.list(z) && !inherits(z, "error"),
                  results.all[["S3"]]$replicates)

rmst.curve <- do.call(rbind, lapply(c("M1", "M3", "M4"), function(m) {
  mat <- do.call(rbind, lapply(reps.s3, function(z) z[[paste0(m, ".rmst")]]))
  data.frame(time = TAU_GRID, estimate = colMeans(mat, na.rm = TRUE), method = m)
}))
rmst.curve <- rbind(data.frame(time = TAU_GRID, estimate = ref.s3$rmst.diff,
                               method = "Truth"),
                    rmst.curve)
rmst.curve$method <- factor(rmst.curve$method, levels = c("Truth", "M1", "M3", "M4"))

ggplot(rmst.curve, aes(x = time, y = estimate, colour = method,
                       linetype = method)) +
  geom_line(linewidth = 0.6) +
  scale_colour_manual(values = c("black", PAL[1:3])) +
  scale_linetype_manual(values = c("solid", LTY[2:4])) +
  labs(x = "Restriction horizon", y = "RMST contrast") +
  theme_pub() +
  guides(colour = guide_legend(nrow = 1), linetype = guide_legend(nrow = 1))
ggsave("Figures/Simulation/rmst-curves.pdf", width = 7, height = 4.4)

hr.curve <- rbind(
  data.frame(time = TAU_GRID, hr = ref.s3$hr.t, method = "Truth"),
  data.frame(time = TAU_GRID, method = "MS-Cox (M5)",
             hr = exp(colMeans(do.call(rbind, lapply(reps.s3, function(z) z$M5.loghr)),
                               na.rm = TRUE))),
  data.frame(time = TAU_GRID, method = "Case-base (M6)",
             hr = exp(colMeans(do.call(rbind, lapply(reps.s3, function(z) z$M6.loghr)),
                               na.rm = TRUE)))
)
hr.curve$method <- factor(hr.curve$method,
                          levels = c("Truth", "MS-Cox (M5)", "Case-base (M6)"))

ggplot(subset(hr.curve, time > 0), aes(x = time, y = hr, colour = method,
                                       linetype = method)) +
  geom_hline(yintercept = 1, linetype = "dashed", colour = "grey40") +
  geom_line(linewidth = 0.6) +
  scale_colour_manual(values = c("black", PAL[1:2])) +
  scale_linetype_manual(values = c("solid", LTY[2:3])) +
  coord_cartesian(ylim = c(0.5, 3)) +
  labs(x = "Time", y = "Marginal hazard ratio") +
  theme_pub() +
  guides(colour = guide_legend(nrow = 1), linetype = guide_legend(nrow = 1))
ggsave("Figures/Simulation/hazard-ratio-time.pdf", width = 7, height = 4.4)

cov.frame <- subset(rmst.summary, method %in% c("M1", "M2", "M3", "M4"))
cov.frame$scenario <- factor(cov.frame$scenario, levels = paste0("S", 1:8))

ggplot(cov.frame, aes(x = factor(tau), y = coverage, colour = method,
                      shape = method, group = method)) +
  geom_hline(yintercept = 0.95, linetype = "dashed", colour = "grey40") +
  geom_line(linewidth = 0.4) +
  geom_point(size = 1.9) +
  facet_wrap(~ scenario, nrow = 2) +
  scale_colour_manual(values = PAL) +
  scale_shape_manual(values = SHP) +
  scale_y_continuous(limits = c(0.7, 1), breaks = seq(0.7, 1, by = 0.05)) +
  labs(x = "Restriction horizon", y = "Empirical coverage") +
  theme_pub() +
  guides(colour = guide_legend(nrow = 1), shape = guide_legend(nrow = 1))
ggsave("Figures/Simulation/coverage.pdf", width = 10, height = 5)
```

# Computation time

The benchmark below is independent of the main study and can be run on its own.
It times a single fit of each estimator at three sample sizes, repeated
`BENCH_REPS` times on one core, and writes the timing table used in the
appendix. Timings are wall-clock seconds per fit and exclude the construction of
the weights, which is common to all six estimators.

```{r}
BENCH_REPS <- 25
BENCH_N <- c(500, 1000, 2000)

time_once <- function(expr) {
  as.numeric(system.time(expr)["elapsed"])
}

benchmark_fits <- function(n, beta.a = log(2), c.max = 3, c.gamma = 1,
                           c.beta = 1) {
  d.set <- generate_data(n, beta.a, c.max, c.gamma, c.beta)
  ps.fit <- glm(A ~ X1 + X2 + X3 + X4 + X5 + X6 + X7 + X8 + X9 + X10,
                data = d.set, family = binomial)
  p.score <- pmin(pmax(fitted(ps.fit), 1e-8), 1 - 1e-8)
  p.marg <- mean(d.set$A)
  ipw <- ifelse(d.set$A == 1, p.marg / p.score, (1 - p.marg) / (1 - p.score))

  c(
    M1 = time_once({
      f <- coxph(Surv(time, status) ~ A, data = d.set, weights = ipw,
                 robust = TRUE)
      cox_rmst(f)
    }),
    M2 = time_once({
      f <- fit_casebase(d.set, ipw, basis = "exponential")
      casebase_rmst(f)
    }),
    M3 = time_once({
      f <- fit_casebase(d.set, ipw, basis = "gompertz")
      casebase_rmst(f)
    }),
    M4 = time_once({
      f <- fit_casebase(d.set, ipw, basis = "spline")
      casebase_rmst(f)
    }),
    M5 = time_once(coxph(Surv(time, status) ~ A + tt(A), data = d.set,
                         weights = ipw, robust = TRUE,
                         tt = function(x, t, ...) x * log(pmax(t, 1e-8)))),
    M6 = time_once(fit_casebase(d.set, ipw, basis = "spline",
                                time.varying = TRUE)),
    events = sum(d.set$status)
  )
}

set.seed(SEED + 7)
timing.raw <- do.call(rbind, lapply(BENCH_N, function(n) {
  cat(sprintf("benchmark n = %d\n", n))
  flush.console()
  m <- do.call(rbind, lapply(seq_len(BENCH_REPS), function(i) benchmark_fits(n)))
  data.frame(n = n, t(colMeans(m)))
}))

timing.summary <- data.frame(
  n = timing.raw$n,
  events = round(timing.raw$events),
  M1 = timing.raw$M1, M2 = timing.raw$M2, M3 = timing.raw$M3,
  M4 = timing.raw$M4, M5 = timing.raw$M5, M6 = timing.raw$M6,
  ratio.M4.M1 = timing.raw$M4 / timing.raw$M1
)
timing.summary

print(xtable(timing.summary, digits = c(0, 0, 0, 3, 3, 3, 3, 3, 3, 2)),
      include.rownames = FALSE, file = "table-sim-timing.tex")
```

# Session information

```{r}
sessionInfo()
```
