---
title: "Causal RMST under case-base sampling with IPW: Simulation Study"
author: "Bilal Shaikh and Sumeet Kalia"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = FALSE, warning = FALSE, message = FALSE)
```

# Global configuration

```{r}
SEED <- 20260811
N_REPLICATES <- 2000
N_REFERENCE <- 1e7
CB_RATIO <- 20
WEIBULL_SHAPE <- 1
TAU_GRID <- seq(0, 3, by = 0.02)
TAU_REPORT <- c(0.5, 1, 2, 3)
N_CORES <- max(1, parallel::detectCores() - 1)
ALPHA <- 0.05
Z_CRIT <- qnorm(1 - ALPHA / 2)

TREATMENT_COEF_RAW <- c(0.7, -0.4, 0.5, 0.2, 0.6, -0.3, 0.1, -0.2, 0.4, -0.5)
OUTCOME_COEF_RAW <- c(0.6, -0.3, 0.4, 0.1, 0.5, -0.2, 0.3, -0.1, 0.4, -0.4)

TREATMENT_COEF <- TREATMENT_COEF_RAW / sqrt(sum(TREATMENT_COEF_RAW^2))
OUTCOME_COEF <- OUTCOME_COEF_RAW / sqrt(sum(OUTCOME_COEF_RAW^2))

SCENARIOS <- data.frame(
  scenario = paste0("S", 1:8),
  label = c(
    "Null treatment effect",
    "Moderate treatment effect",
    "Reference configuration",
    "Reduced sample size",
    "Small sample size",
    "Heavy censoring",
    "Strong confounding",
    "Weak prognostic covariate effect"
  ),
  trt.effect = c(0, log(1.5), log(2), log(2), log(2), log(2), log(2), log(2)),
  n = c(1000, 1000, 1000, 500, 250, 1000, 1000, 1000),
  cens.max = c(3.0, 3.0, 3.0, 3.0, 3.0, 1.5, 3.0, 3.0),
  trt.lin.pred.scale = c(1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 2.0, 1.0),
  outcome.lin.pred.scale = c(1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 0.5),
  stringsAsFactors = FALSE
)

SCENARIOS

dir.create("Figures", recursive = TRUE, showWarnings = FALSE)
dir.create("Tables", recursive = TRUE, showWarnings = FALSE)
```

# Data-generating mechanism

Covariates are standardized to zero mean and unit variance so that 
`trt.lin.pred.scale` is the standard deviation of the treatment linear predictor
and `outcome.lin.pred.scale` scales the prognostic index.

```{r}
standardize <- function(value, mu, sigma) {
  return((value - mu) / sigma)
}

draw_covariates <- function(n) {
  x1 <- rnorm(n)
  x2 <- rnorm(n)
  x3 <- standardize(runif(n), 0.5, sqrt(1 / 12))
  x4 <- standardize(runif(n), 0.5, sqrt(1 / 12))
  
  x5.mean <- exp(0.125)
  x5.sd <- sqrt((exp(0.25) - 1) * exp(0.25))
  x5 <- standardize(rlnorm(n, meanlog = 0, sdlog = 0.5), x5.mean, x5.sd)
  
  x6 <- standardize(rgamma(n, shape = 2, rate = 1), 2, sqrt(2))
  x7 <- standardize(rexp(n, rate = 1), 1, 1)
  x8 <- standardize(rbinom(n, 1, 0.5), 0.5, sqrt(0.25))
  x9 <- standardize(rbinom(n, 1, 0.3), 0.3, sqrt(0.21))
  x10 <- standardize(rpois(n, lambda = 1.5), 1.5, sqrt(1.5))
  
  return(cbind(
    X1 = x1, X2 = x2, X3 = x3, X4 = x4, X5 = x5,
    X6 = x6, X7 = x7, X8 = x8, X9 = x9, X10 = x10
  ))
}

generate_data <- function(
    n, trt.effect, cens.max,
    trt.lin.pred.scale,
    outcome.lin.pred.scale  
  ) {
  
  covariate.matrix <- draw_covariates(n)
  
  treat.linear.pred <- trt.lin.pred.scale * as.vector(
    covariate.matrix %*% TREATMENT_COEF
  )
  
  e.score.treat <- plogis(treat.linear.pred)
  treatment.vec <- rbinom(n, 1, e.score.treat)
  
  outcome.linear.pred <-
    outcome.lin.pred.scale * as.vector(covariate.matrix %*% OUTCOME_COEF) +
    trt.effect * treatment.vec
  
  unif.draw <- runif(n)
  event.time <- (-log(unif.draw) / exp(outcome.linear.pred))^(1 / WEIBULL_SHAPE)
  censor.time <- runif(n, 0, cens.max)
  
  return(data.frame(
    time = pmin(event.time, censor.time),
    status = as.integer(event.time <= censor.time),
    A = treatment.vec,
    covariate.matrix
  ))
}
```

# Reference values

The counterfactual survival function is a covariate mixture and is evaluated by
Monte Carlo integration over the covariate distribution alone. No event times,
no censoring and no working model enter this computation.

```{r}
rmst_from_survival <- function(time.grid, surv.vec) {
  d.t <- diff(time.grid)
  surv.vec.mids <- (head(surv.vec, -1) + tail(surv.vec, -1)) / 2
  return(c(0, cumsum(surv.vec.mids * d.t)))
}

reference_values <- function(
    trt.effect, outcome.lin.pred.scale,
    n.ref = N_REFERENCE, chunk = 1e5  
  ) {
  
  n.completed <- 0
  next.report <- 0.1
  
  s0.total <- rep(0, length(TAU_GRID))
  s1.total <- rep(0, length(TAU_GRID))
  
  h0.numerator <- rep(0, length(TAU_GRID))
  h1.numerator <- rep(0, length(TAU_GRID))
  
  while (n.completed < n.ref) {
    chunk.n <- min(chunk, n.ref - n.completed)
    
    covariate.matrix <- draw_covariates(chunk.n)
    outcome.linear.pred <- outcome.lin.pred.scale * as.vector(
      covariate.matrix %*% OUTCOME_COEF
    )
    
    base.cum.haz.grid <- outer(TAU_GRID^WEIBULL_SHAPE, rep(1, chunk.n))
    
    hazard0 <- exp(outcome.linear.pred)
    hazard1 <- exp(outcome.linear.pred + trt.effect)
    
    s0.grid <- exp(-base.cum.haz.grid * rep(hazard0, each = length(TAU_GRID)))
    s1.grid <- exp(-base.cum.haz.grid * rep(hazard1, each = length(TAU_GRID)))
    
    s0.total <- s0.total + rowSums(s0.grid)
    s1.total <- s1.total + rowSums(s1.grid)
    
    h0.numerator <- 
      h0.numerator +
      rowSums(s0.grid * rep(hazard0, each = length(TAU_GRID)))
    h1.numerator <-
      h1.numerator +
      rowSums(s1.grid * rep(hazard1, each = length(TAU_GRID)))
    
    n.completed <- n.completed + chunk.n
    
    if (n.completed / n.ref >= next.report) {
      cat(
        sprintf("  reference integration: %3.0f%%\n", 100 * n.completed / n.ref)
      )
      flush.console()
      next.report <- next.report + 0.1
    }
  }
  
  s0 <- s0.total / n.ref
  s1 <- s1.total / n.ref
  
  hazard0.estimate <- 
    WEIBULL_SHAPE *
    TAU_GRID^(WEIBULL_SHAPE - 1) *
    (h0.numerator / n.ref) / pmax(s0, .Machine$double.eps)
  hazard1.estimate <-
    WEIBULL_SHAPE *
    TAU_GRID^(WEIBULL_SHAPE - 1) *
    (h1.numerator / n.ref) / pmax(s1, .Machine$double.eps)
  
  return(list(
    time.grid = TAU_GRID,
    s0 = s0, s1 = s1,
    rmst.diff =
      rmst_from_survival(TAU_GRID, s1) -
      rmst_from_survival(TAU_GRID, s0),
    hr.t = hazard1.estimate / hazard0.estimate
  ))
}
```

The constant-effect estimators are additionally compared against the large-sample
limit of the corresponding fitting criterion under an exogenous-treatment version
of the same mechanism. This limit is a projection attached to the criterion and
not a structural parameter.

```{r}
PROJ_N <- 5e5
PROJ_RATIO <- 5
proj.cache <- new.env(parent = emptyenv())

projection_limits <- function(trt.effect, outcome.lin.pred.scale, cens.max) {
  
  cache.key <- paste(trt.effect, outcome.lin.pred.scale, cens.max, sep = "_")
  
  if (!is.null(proj.cache[[cache.key]])) {
    return(proj.cache[[cache.key]])
  }
  
  set.seed(SEED + 991)
  proj.data <- generate_data(
    PROJ_N, trt.effect, cens.max,
    trt.lin.pred.scale = 0,
    outcome.lin.pred.scale = outcome.lin.pred.scale
  )
  unit.weights <- rep(1, nrow(proj.data))
  
  cox.fit <- survival::coxph(Surv(time, status) ~ A, data = proj.data)
  cb.object <- fit_casebase(
    proj.data, unit.weights, basis = "spline", ratio = PROJ_RATIO
  )
  
  cox.ref.rmst <- cox_rmst(cox.fit)
  
  result <- list(
    psi = c(
      cox = unname(coef(cox.fit)["A"]),
      casebase = unname(coef(cb.object$fit)["A"])
    ),
    rmst.cox = cox.ref.rmst$rmst
  )
  proj.cache[[cache.key]] <- result
  return(result)
}
```

# Weighted case-base fitting

`sampleCaseBase` draws base moments proportionally to observed follow-up time.
Observation weights are attached before sampling so that each sampled moment
inherits the weight of the individual from which it was drawn, which makes the
weighted score equation target the weighted hazard.

```{r}
interior_knots <- function(event.times, boundary, probs = c(0.25, 0.5, 0.75)) {
  if (length(event.times) < 20) {
    return(numeric(0))
  }
  
  knots <- unique(
    as.numeric(quantile(event.times, probs = probs, names = FALSE))
  )
  knots <- knots[
    knots > boundary[1] +
      .Machine$double.eps & knots < boundary[2] -
      .Machine$double.eps
  ]
  
  return(sort(unique(knots)))
}

fit_casebase <- function(
    data.set, weights, basis = c("spline", "gompertz", "exponential"),
    time.varying = FALSE, ratio = CB_RATIO, seed = NULL  
  ) {
  
  basis <- match.arg(basis)
  
  data.set.augmented <- data.set
  data.set.augmented$w <- weights
  
  if (!is.null(seed)) {
    set.seed(seed)
  }
  
  cb.data <- casebase::sampleCaseBase(
    data.set.augmented, time = "time", event = "status", ratio = ratio
  )
  boundary <- range(cb.data$time)
  interior.knots <- interior_knots(data.set$time[data.set$status == 1], boundary)
  time.term <- switch(
    basis,
    exponential = NULL,
    gompertz = "time",
    spline = if (length(interior.knots) > 0) {
      sprintf(
        "splines::bs(time, knots = c(%s), Boundary.knots = c(%s), degree = 3)",
        paste(interior.knots, collapse = ", "), paste(boundary, collapse = ", ")
      )
      
    } else {
      sprintf(
        "splines::bs(time, df = 3, Boundary.knots = c(%s), degree = 3)",
        paste(boundary, collapse = ", ")
      )
    }
  )
  
  right.hand.side <- if (is.null(time.term)) {
    "A"
  } else if (time.varying) {
    paste0("A * ", time.term)
  } else {
    paste0("A + ", time.term)
  }
  
  form <- as.formula(paste("status ~", right.hand.side, "+ offset(offset)"))
  fit <- glm(form, data = cb.data, family = binomial, weights = w)
  
  return(list(
    fit = fit,
    vcov = sandwich::vcovHC(fit, type = "HC0"),
    knots = interior.knots,
    boundary = boundary,
    basis = basis,
    time.varying = time.varying
  ))
}

casebase_hazard_grid <- function(cb.object, a.value, time.grid) {
  new.data <- data.frame(
    time = pmax(time.grid, .Machine$double.eps), A = a.value, offset = 0
  )
  design.matrix <- model.matrix(
    delete.response(terms(cb.object$fit)), data = new.data
  )
  design.matrix <- design.matrix[, names(coef(cb.object$fit)), drop = FALSE]
  
  return(list(
    design.matrix = design.matrix,
    hazard = as.vector(exp(design.matrix %*% coef(cb.object$fit)))
  ))
}
```

# RMST contrast and delta-method standard error

For the case-base model the gradient is available in closed form. Writing
$\lambda^{(a)}(u) = \exp\{z(u,a)^\top\boutcome.linear.pred\}$, the gradient of the survival
function is $-S^{(a)}(t)\int_0^t \lambda^{(a)}(u) z(u,a)\,du$, and the RMST
gradient integrates this over the horizon.

```{r}
casebase_rmst <- function(cb.object, time.grid = TAU_GRID) {
  d.t <- diff(time.grid)
  
  grad.list <- list()
  surv.list <- list()
  
  for (a.value in c(0, 1)) {
    hazard.grid <- casebase_hazard_grid(cb.object, a.value, time.grid)
    hazard <- hazard.grid$hazard
    cum.haz <- c(0, cumsum(d.t * (head(hazard, -1) + tail(hazard, -1)) / 2))
    surv.vec <- exp(-cum.haz)
    hazard.weighted.design <- hazard.grid$design.matrix * hazard
    grad.cum.hazard <- rbind(
      0, apply(
        hazard.weighted.design, 2,
        function(integrand.column) {
          return(cumsum(
            d.t * (head(integrand.column, -1) + tail(integrand.column, -1)) / 2
          ))
        }
      )
    )
    
    grad.surv <- -surv.vec * grad.cum.hazard
    grad.rmst <- rbind(0, apply(grad.surv, 2, function(integrand.column) {
      return(cumsum(
        d.t * (head(integrand.column, -1) + tail(integrand.column, -1)) / 2
      ))
    }))
    
    surv.list[[as.character(a.value)]] <- surv.vec
    grad.list[[as.character(a.value)]] <- grad.rmst
  }
  
  rmst.diff <- 
    rmst_from_survival(time.grid, surv.list[["1"]]) -
    rmst_from_survival(time.grid, surv.list[["0"]])
  
  rmst.grad.diff <- grad.list[["1"]] - grad.list[["0"]]
  var.matrix <- cb.object$vcov[
    colnames(rmst.grad.diff), colnames(rmst.grad.diff), drop = FALSE
  ]
  
  rmst.se <- sqrt(
    pmax(rowSums((rmst.grad.diff %*% var.matrix) * rmst.grad.diff), 0)
  )
  
  return(list(
    time.grid = time.grid,
    s0 = surv.list[["0"]], s1 = surv.list[["1"]],
    rmst = rmst.diff, se = rmst.se
  ))
}
```

For the marginal structural Cox model the baseline is held fixed and only the
treatment coefficient contributes, which gives
$\partial S^{(1)}/\partial\psi = S^{(1)}(t)\log S^{(1)}(t)$.

```{r}
cox_rmst <- function(cox.fit, time.grid = TAU_GRID) {
  survival.fit <- survival::survfit(cox.fit, newdata = data.frame(A = c(0, 1)))
  
  s0 <- approx(
    survival.fit$time, survival.fit$surv[, 1], time.grid,
    method = "constant", rule = 2
  )$y
  s1 <- approx(
    survival.fit$time, survival.fit$surv[, 2], time.grid,
    method = "constant", rule = 2
  )$y
  
  se.psi <- sqrt(diag(vcov(cox.fit))["A"])
  integrand <- ifelse(s1 > 0, s1 * log(pmax(s1, .Machine$double.eps)), 0)
  gradient.vec <- rmst_from_survival(time.grid, integrand)
  
  return(list(
    time.grid = time.grid,
    s0 = s0, s1 = s1,
    rmst = rmst_from_survival(time.grid, s1) - rmst_from_survival(time.grid, s0),
    se = abs(gradient.vec) * as.numeric(se.psi)
  ))
}
```

# One replicate

```{r}
run_replicate <- function(
    n, trt.effect, cens.max,
    trt.lin.pred.scale,
    outcome.lin.pred.scale
  ) {
  
  data.set <- generate_data(
    n, trt.effect, cens.max,
    trt.lin.pred.scale,
    outcome.lin.pred.scale
  )
  
  p.score.fit <- glm(
    A ~ X1 + X2 + X3 + X4 + X5 + X6 + X7 + X8 + X9 + X10,
    data = data.set, family = binomial
  )
  
  p.score.treat <- pmin(pmax(fitted(p.score.fit), 1e-8), 1 - 1e-8)
  p.marginal <- mean(data.set$A)
  
  iptw <- ifelse(
    data.set$A == 1,
    p.marginal / p.score.treat,
    (1 - p.marginal) / (1 - p.score.treat)
  )

  result <- list(
    iptw.max = max(iptw),
    iptw.mean = mean(iptw),
    relative.ess = sum(iptw)^2 / sum(iptw^2) / n,
    event.rate = mean(data.set$status)
  )

  cox.iptw.fit <- try(survival::coxph(
    Surv(time, status) ~ A, data = data.set, weights = iptw, robust = TRUE
  ), silent = TRUE)
  
  if (!inherits(cox.iptw.fit, "try-error")) {
    cox.iptw.rmst <- cox_rmst(cox.iptw.fit)
    result$M1.psi <- unname(coef(cox.iptw.fit)["A"])
    result$M1.se <- unname(sqrt(diag(vcov(cox.iptw.fit))["A"]))
    result$M1.rmst <- cox.iptw.rmst$rmst
    result$M1.rmst.se <- cox.iptw.rmst$se
  }

  for (model.label in c("M2", "M3", "M4")) {
    basis <- switch(
      model.label,
      M2 = "exponential", M3 = "gompertz", M4 = "spline"
    )
    
    cb.object <- try(fit_casebase(data.set, iptw, basis = basis), silent = TRUE)
    if (!inherits(cb.object, "try-error") && cb.object$fit$converged) {
      cb.rmst <- casebase_rmst(cb.object)
      result[[paste0(model.label, ".psi")]] <- unname(
        coef(cb.object$fit)["A"]
      )
      result[[paste0(model.label, ".se")]] <- unname(
        sqrt(diag(cb.object$vcov)["A"])
      )
      result[[paste0(model.label, ".rmst")]] <- cb.rmst$rmst
      result[[paste0(model.label, ".rmst.se")]] <- cb.rmst$se
    }
  }

  cox.time.varying.fit <- try(survival::coxph(
    Surv(time, status) ~ A + tt(A), data = data.set,
    weights = iptw, robust = TRUE,
    tt = function(x, t, ...) x * log(pmax(t, 1e-8))
  ), silent = TRUE)
  
  if (!inherits(cox.time.varying.fit, "try-error")) {
    cox.time.varying.coef <- coef(cox.time.varying.fit)
    cox.time.varying.vcov <- vcov(cox.time.varying.fit)
    log.t <- log(pmax(TAU_GRID, 1e-8))
    result$M5.loghr <-
      cox.time.varying.coef[1] +
      cox.time.varying.coef[2] * log.t
    result$M5.loghr.se <-sqrt(
      cox.time.varying.vcov[1, 1] +
        log.t^2 * cox.time.varying.vcov[2, 2] +
        2 * log.t * cox.time.varying.vcov[1, 2]
    )
  }

  cb.time.varying.fit <- try(fit_casebase(
    data.set, iptw, basis = "spline", time.varying = TRUE
  ), silent = TRUE)
  
  if (
    !inherits(cb.time.varying.fit, "try-error") &&
    cb.time.varying.fit$fit$converged
  ) {
    
    hazard.design.1 <- casebase_hazard_grid(
      cb.time.varying.fit, 1, TAU_GRID
    )$design.matrix
    
    hazard.design.0 <- casebase_hazard_grid(
      cb.time.varying.fit, 0, TAU_GRID
    )$design.matrix
    
    design.diff <- hazard.design.1 - hazard.design.0
    result$M6.loghr <- as.vector(design.diff %*% coef(cb.time.varying.fit$fit))
    
    result$M6.loghr.se <- sqrt(
      pmax(rowSums((design.diff %*% cb.time.varying.fit$vcov) * design.diff), 0)
    )
  }

  return(result)
}
```

# Running the study

```{r}
cluster <- parallel::makeCluster(N_CORES)
doParallel::registerDoParallel(cluster)
doRNG::registerDoRNG(SEED)
`%dopar%` <- foreach::`%dopar%`

results.all <- list()

for (scenario.idx in seq_len(nrow(SCENARIOS))) {
  scenario.row <- SCENARIOS[scenario.idx, ]
  
  cat(sprintf(
    "\n===== Scenario %s: %s =====\n", scenario.row$scenario, scenario.row$label
  ))
  cat("computing reference values\n")
  
  reference <- reference_values(
    scenario.row$trt.effect,
    scenario.row$outcome.lin.pred.scale
  )
  projection <- projection_limits(
    scenario.row$trt.effect,
    scenario.row$outcome.lin.pred.scale,
    scenario.row$cens.max
  )

  chunk.size <- ceiling(N_REPLICATES / 20)
  chunk.start <- seq(1, N_REPLICATES, by = chunk.size)
  reps <- list()
  
  for (curr.chunk in seq_along(chunk.start)) {
    chunk.n <- min(chunk.size, N_REPLICATES - chunk.start[curr.chunk] + 1)
    reps[[curr.chunk]] <- foreach::foreach(
      i = seq_len(chunk.n),
      .packages = c("survival", "casebase", "sandwich", "splines"),
      .export = c(
        "generate_data", "draw_covariates", "standardize",
        "fit_casebase", "casebase_hazard_grid", "casebase_rmst",
        "cox_rmst", "rmst_from_survival", "run_replicate",
        "interior_knots", "TREATMENT_COEF", "OUTCOME_COEF", "WEIBULL_SHAPE",
        "TAU_GRID", "CB_RATIO"
      ),
      .errorhandling = "pass"
      
    ) %dopar% {
      run_replicate(
        scenario.row$n,
        scenario.row$trt.effect,
        scenario.row$cens.max,
        scenario.row$trt.lin.pred.scale,
        scenario.row$outcome.lin.pred.scale
      )
    }
    cat(sprintf(
      "  %s: %3.0f%% complete\n",
      scenario.row$scenario,
      100 *
        min(chunk.start[curr.chunk] + chunk.n - 1, N_REPLICATES) /
        N_REPLICATES
    ))
    
    flush.console()
  }
  
  results.all[[scenario.row$scenario]] <- list(
    scenario = scenario.row,
    reference = reference,
    projection = projection,
    replicates = do.call(c, reps)
  )
}

parallel::stopCluster(cluster)
saveRDS(results.all, file = "simulation-results.rds")
```

# Performance measures

```{r}
summarize_scalar <- function(estimate, se, truth) {
  
  finite.idx <- is.finite(estimate) & is.finite(se)
  estimate <- estimate[finite.idx]
  se <- se[finite.idx]
  
  n.valid <- length(estimate)
  
  bias <- mean(estimate) - truth
  mcsd <- sd(estimate)
  
  lower <- estimate - Z_CRIT * se
  upper <- estimate + Z_CRIT * se
  
  return(data.frame(
    n.used = n.valid,
    truth = truth,
    mean.est = mean(estimate),
    bias = bias,
    bias.mcse = mcsd / sqrt(n.valid),
    mcsd = mcsd,
    ase = mean(se),
    mse = mean((estimate - truth)^2),
    coverage = mean(lower <= truth & truth <= upper),
    coverage.mcse = sqrt(
      mean(lower <= truth & truth <= upper) *
        (1 - mean(lower <= truth & truth <= upper)) /
        n.valid
    ),
    reject = mean(abs(estimate / se) > Z_CRIT)
  ))
}

extract_grid <- function(valid.reps, field, index) {
  return(vapply(valid.reps, function(replicate.result) {
    field.value <- replicate.result[[field]]
    return(ifelse(
      is.null(field.value) || length(field.value) < index,
      NA_real_,
      field.value[index]
    ))
  }, numeric(1)))
}

tau.index <- match(TAU_REPORT, TAU_GRID)

rmst.summary <- do.call(rbind, lapply(
  names(results.all),
  function(scenario.name) {
    
  scenario.results <- results.all[[scenario.name]]
  valid.reps <- Filter(
    function(replicate.result) {
      return(is.list(replicate.result) &&
      !inherits(replicate.result, "error"))
    },
    scenario.results$replicates
  )
  
  return(do.call(rbind, lapply(
    c("M1", "M2", "M3", "M4"),
    function(model.name) {
      return(do.call(rbind, lapply(
        seq_along(TAU_REPORT),
        function(tau.report.idx) {
          grid.idx <- tau.index[tau.report.idx]
          estimate <- extract_grid(
            valid.reps, paste0(model.name, ".rmst"), grid.idx
          )
          se <- extract_grid(
            valid.reps, paste0(model.name, ".rmst.se"), grid.idx
          )
          return(cbind(
            data.frame(
              scenario = scenario.name,
              method = model.name,
              tau = TAU_REPORT[tau.report.idx]
            ),
            summarize_scalar(
              estimate, se, scenario.results$reference$rmst.diff[grid.idx]
            )
          ))
        }
      )))
    }
  )))
  }
))

hr.summary <- do.call(rbind, lapply(
  names(results.all),
  function(scenario.name) {
    scenario.results <- results.all[[scenario.name]]
    valid.reps <- Filter(
      function(replicate.result) {
        is.list(replicate.result) && !inherits(replicate.result, "error")
      },
      scenario.results$replicates
    )
    return(do.call(rbind, lapply(
      c("M1", "M2", "M3", "M4"),
      function(model.name) {
        estimate <- vapply(
          valid.reps,
          function(replicate.result) {
            field.value <- replicate.result[[paste0(model.name, ".psi")]]
            return(ifelse(is.null(field.value), NA_real_, field.value))
          }, numeric(1)
        )
        se <- vapply(
          valid.reps,
          function(replicate.result) {
            field.value <- replicate.result[[paste0(model.name, ".se")]]
            return(ifelse(is.null(field.value), NA_real_, field.value))
          }, numeric(1)
        )
        scenario.projection <- ifelse(
          is.list(scenario.results$projection),
          scenario.results$projection$psi,
          scenario.results$projection
        )
        truth <- ifelse(
          model.name == "M1",
          scenario.projection["cox"],
          scenario.projection["casebase"]
        )
        return(cbind(
          data.frame(scenario = scenario.name, method = model.name),
          summarize_scalar(estimate, se, truth)
        ))
      }
    )))
  }
))

convergence.summary <- do.call(rbind, lapply(
  names(results.all),
  function(scenario.name) {
    replicates <- results.all[[scenario.name]]$replicates
    error.count <- sum(vapply(
      replicates,
      function(replicate.result) inherits(replicate.result, "error"),
      logical(1)
    ))
    valid.reps <- Filter(
      function(replicate.result) {
        return(is.list(replicate.result) && !inherits(replicate.result, "error"))
      },
      replicates
    )
    
    return(data.frame(
      scenario = scenario.name,
      replicates = length(replicates),
      hard.failures = error.count,
      
      M1.missing = sum(vapply(
        valid.reps,
        function(replicate.result) is.null(replicate.result$M1.psi),
        logical(1)
      )),
      M2.missing = sum(vapply(
        valid.reps,
        function(replicate.result) is.null(replicate.result$M2.psi),
        logical(1)
      )),
      M3.missing = sum(vapply(
        valid.reps,
        function(replicate.result) is.null(replicate.result$M3.psi),
        logical(1)
      )),
      M4.missing = sum(vapply(
        valid.reps,
        function(replicate.result) is.null(replicate.result$M4.psi),
        logical(1)
        )),
      mean.ipw = mean(vapply(
        valid.reps,
        function(replicate.result) replicate.result$iptw.mean,
        numeric(1)
      )),
      max.ipw = max(vapply(
        valid.reps,
        function(replicate.result) replicate.result$iptw.max,
        numeric(1)
      )),
      relative.ess = mean(vapply(
        valid.reps,
        function(replicate.result) replicate.result$relative.ess,
        numeric(1)
      )),
      event.rate = mean(vapply(
        valid.reps,
        function(replicate.result) replicate.result$event.rate,
        numeric(1)
        ))
    ))
  }
))

rmst.summary
hr.summary
convergence.summary
```

# LaTeX tables

```{r}
print(
  xtable::xtable(
    rmst.summary[, c(
      "scenario", "method", "tau", "truth", "mean.est",
      "bias", "mcsd", "ase", "mse", "coverage"
    )],
    digits = c(0, 0, 0, 2, 4, 4, 4, 4, 4, 4, 3)
  ),
  include.rownames = FALSE, file = "Tables/rmst.tex"
)

print(
  xtable::xtable(
    hr.summary[, c(
      "scenario", "method", "truth", "mean.est", "bias",
      "mcsd", "ase", "mse", "coverage", "reject"
    )],
    digits = c(0, 0, 0, 4, 4, 4, 4, 4, 4, 3, 3)
  ),
  include.rownames = FALSE, file = "Tables/hr.tex"
)

print(
  xtable::xtable(
    convergence.summary, digits = c(0, 0, 0, 0, 0, 0, 0, 0, 3, 2, 3, 3)
  ),
  include.rownames = FALSE, file = "Tables/convergence.tex"
)
```

# Reference values: integration against a model-based alternative

This section quantifies what is lost by defining the reference value through a
fitted working model instead of by direct integration. It does not depend on the
replicates, only on the data-generating mechanism, so it can be run on its own
and is unaffected by objects left over from an earlier run.

```{r}
ref.cache <- new.env(parent = emptyenv())

cox_reference_rmst <- function(
    trt.effect, outcome.lin.pred.scale, cens.max, n.proj = PROJ_N
  ) {
  
  cache.key <- paste(trt.effect, outcome.lin.pred.scale, cens.max, sep = "_")
  
  if (!is.null(ref.cache[[cache.key]])) {
    return(ref.cache[[cache.key]])
  }
  
  set.seed(SEED + 991)
  proj.data <- generate_data(
    n.proj, trt.effect, cens.max,
    trt.lin.pred.scale = 0,
    outcome.lin.pred.scale = outcome.lin.pred.scale
  )
  cox.fit <- survival::coxph(Surv(time, status) ~ A, data = proj.data)
  
  result <- cox_rmst(cox.fit)$rmst
  ref.cache[[cache.key]] <- result
  
  return(result)
}

integration_reference <- function(
    scenario.name, trt.effect, outcome.lin.pred.scale  
  ) {
  
  scenario.results <- ifelse(
    exists("results.all"), results.all[[scenario.name]], NULL
  )
  
  if (!is.null(scenario.results) && is.list(scenario.results$reference)) {
    return(scenario.results$reference$rmst.diff)
  }
  
  return(reference_values(trt.effect, outcome.lin.pred.scale)$rmst.diff)
}

reference.compare <- do.call(rbind, lapply(
  seq_len(nrow(SCENARIOS)),
  function(scenario.idx) {
    scenario.row <- SCENARIOS[scenario.idx, ]
    
    cat(sprintf("reference comparison: %s\n", scenario.row$scenario))
    flush.console()
    
    exact.curve <- integration_reference(
      scenario.row$scenario,
      scenario.row$trt.effect,
      scenario.row$outcome.lin.pred.scale
    )
    model.curve <- cox_reference_rmst(
      scenario.row$trt.effect,
      scenario.row$outcome.lin.pred.scale,
      scenario.row$cens.max
    )
    
    return(do.call(rbind, lapply(
      c(1, 3),
      function(tau) {
        grid.idx <- match(tau, TAU_GRID)
        exact <- exact.curve[grid.idx]
        modelled <- model.curve[grid.idx]
        
        return(data.frame(
          scenario = scenario.row$scenario, tau = tau,
          integration = exact, cox.working = modelled,
          difference = modelled - exact,
          pct.of.truth = ifelse(
            abs(exact) < 1e-10,
            NA_real_,
            100 * (modelled - exact) / abs(exact)
          )
        ))
      }
    )))
  }
))

reference.compare

print(
  xtable::xtable(reference.compare, digits = c(0, 0, 0, 4, 4, 4, 1)),
  include.rownames = FALSE, file = "Tables/reference.tex"
)
```

# Figures

```{r}
palette <- c("red", "blue", "green4", "black")
line_types <- c("solid", "dashed", "dotdash", "dotted")
shapes <- c(16, 17, 15, 3)

plot_theme <- function(base.size = 10) {
  return(ggplot2::theme_minimal(base_size = base.size) +
    ggplot2::theme(
      panel.grid.minor = ggplot2::element_blank(),
      panel.grid.major = ggplot2::element_line(
        colour = "grey90", linewidth = 0.3
      ),
      panel.border = ggplot2::element_rect(
        colour = "grey30", fill = NA, linewidth = 0.4
      ),
      strip.background = ggplot2::element_blank(),
      strip.text = ggplot2::element_text(face = "bold", size = base.size),
      axis.text = ggplot2::element_text(colour = "grey20"),
      legend.position = "bottom",
      legend.direction = "horizontal",
      legend.title = ggplot2::element_blank(),
      legend.key.width = ggplot2::unit(1.4, "lines")
    ))
}

s3.reference <- results.all[["S3"]]$reference
s3.valid.reps <- Filter(
  function(replicate.result) {
    return(is.list(replicate.result) && !inherits(replicate.result, "error"))
  },
  results.all[["S3"]]$replicates
)

rmst.curve <- do.call(rbind, lapply(
  c("M1", "M3", "M4"),
  function(model.name) {
    rmst.matrix <- do.call(rbind, lapply(
      s3.valid.reps,
      function(replicate.result) {
        return(replicate.result[[paste0(model.name, ".rmst")]])
      }
    ))
  
    return(data.frame(
      time = TAU_GRID,
      estimate = colMeans(rmst.matrix, na.rm = TRUE),
      method = model.name
    ))
  }
))

rmst.curve <- rbind(
  data.frame(
    time = TAU_GRID, estimate = s3.reference$rmst.diff, method = "Truth"
  ),
  rmst.curve
)

rmst.curve$method <- factor(
  rmst.curve$method, levels = c("Truth", "M1", "M3", "M4")
)

ggplot2::ggplot(
  rmst.curve, ggplot2::aes(
    x = time, y = estimate, colour = method, linetype = method
  )
) +
  ggplot2::geom_line(linewidth = 0.6) +
  ggplot2::scale_colour_manual(values = c("black", palette[1:3])) +
  ggplot2::scale_linetype_manual(values = c("solid", line_types[2:4])) +
  ggplot2::labs(x = "Restriction horizon", y = "RMST contrast") +
  plot_theme() +
  ggplot2::guides(
    colour = ggplot2::guide_legend(nrow = 1),
    linetype = ggplot2::guide_legend(nrow = 1)
  )

ggplot2::ggsave("Figures/rmst-curves.pdf", width = 7, height = 4.4)

hr.curve <- rbind(
  data.frame(time = TAU_GRID, hr = s3.reference$hr.t, method = "Truth"),
  data.frame(
    time = TAU_GRID, method = "MS-Cox (M5)",
    hr = exp(colMeans(
      do.call(rbind, lapply(
        s3.valid.reps, function(replicate.result) replicate.result$M5.loghr
      )), na.rm = TRUE
    ))
  ),
  data.frame(
    time = TAU_GRID, method = "Case-base (M6)",
    hr = exp(colMeans(
      do.call(rbind, lapply(
        s3.valid.reps, function(replicate.result) replicate.result$M6.loghr
      )), na.rm = TRUE
    ))
  )
)

hr.curve$method <- factor(
  hr.curve$method,
  levels = c("Truth", "MS-Cox (M5)", "Case-base (M6)")
)

ggplot2::ggplot(
  subset(hr.curve, time > 0), ggplot2::aes(
    x = time, y = hr, colour = method, linetype = method
  )
) +
  ggplot2::geom_hline(yintercept = 1, linetype = "dashed", colour = "grey40") +
  ggplot2::geom_line(linewidth = 0.6) +
  ggplot2::scale_colour_manual(values = c("black", palette[1:2])) +
  ggplot2::scale_linetype_manual(values = c("solid", line_types[2:3])) +
  ggplot2::coord_cartesian(ylim = c(0.5, 3)) +
  ggplot2::labs(x = "Time", y = "Marginal hazard ratio") +
  plot_theme() +
  ggplot2::guides(
    colour = ggplot2::guide_legend(nrow = 1),
    linetype = ggplot2::guide_legend(nrow = 1)
  )

ggplot2::ggsave("Figures/hazard-ratio-time.pdf", width = 7, height = 4.4)

coverage.frame <- subset(rmst.summary, method %in% c("M1", "M2", "M3", "M4"))
coverage.frame$scenario <- factor(
  coverage.frame$scenario, levels = paste0("S", 1:8)
)

ggplot2::ggplot(
  coverage.frame, ggplot2::aes(
    x = factor(tau), y = coverage, colour = method,
    shape = method, group = method
  )
) +
  ggplot2::geom_hline(yintercept = 0.95, linetype = "dashed", colour = "grey40") +
  ggplot2::geom_line(linewidth = 0.4) +
  ggplot2::geom_point(size = 1.9) +
  ggplot2::facet_wrap(~ scenario, nrow = 2) +
  ggplot2::scale_colour_manual(values = palette) +
  ggplot2::scale_shape_manual(values = shapes) +
  ggplot2::scale_y_continuous(
    limits = c(0.7, 1), breaks = seq(0.7, 1, by = 0.05)
  ) +
  ggplot2::labs(x = "Restriction horizon", y = "Empirical coverage") +
  plot_theme() +
  ggplot2::guides(
    colour = ggplot2::guide_legend(nrow = 1),
    shape = ggplot2::guide_legend(nrow = 1)
  )

ggplot2::ggsave("Figures/coverage.pdf", width = 10, height = 5)
```

# Computation time

The benchmark below is independent of the main study and can be run on its own.
It times a single fit of each estimator at three sample sizes, repeated
`BENCH_REPS` times on one core, and writes the timing table used in the
appendix. Timings are wall-clock seconds per fit and exclude the construction of
the weights, which is common to all six estimators.

```{r}
BENCH_REPS <- 25
BENCH_N <- c(500, 1000, 2000)

time_once <- function(expr) {
  return(as.numeric(system.time(expr)["elapsed"]))
}

benchmark_fits <- function(
    n, trt.effect = log(2), cens.max = 3,
    trt.lin.pred.scale = 1,
    outcome.lin.pred.scale = 1  
  ) {
  
  data.set <- generate_data(
    n, trt.effect, cens.max, trt.lin.pred.scale, outcome.lin.pred.scale
  )
  
  p.score.fit <- glm(
    A ~ X1 + X2 + X3 + X4 + X5 + X6 + X7 + X8 + X9 + X10,
    data = data.set, family = binomial
  )
  
  p.score.treat <- pmin(pmax(fitted(p.score.fit), 1e-8), 1 - 1e-8)
  p.marginal <- mean(data.set$A)
  
  iptw <- ifelse(
    data.set$A == 1,
    p.marginal / p.score.treat,
    (1 - p.marginal) / (1 - p.score.treat)
  )

  return(c(
    M1 = time_once({
      model.fit <- survival::coxph(
        Surv(time, status) ~ A, data = data.set, weights = iptw, robust = TRUE
      )
      cox_rmst(model.fit)
    }),
    M2 = time_once({
      model.fit <- fit_casebase(data.set, iptw, basis = "exponential")
      casebase_rmst(model.fit)
    }),
    M3 = time_once({
      model.fit <- fit_casebase(data.set, iptw, basis = "gompertz")
      casebase_rmst(model.fit)
    }),
    M4 = time_once({
      model.fit <- fit_casebase(data.set, iptw, basis = "spline")
      casebase_rmst(model.fit)
    }),
    M5 = time_once(survival::coxph(
      Surv(time, status) ~ A + tt(A), data = data.set,
      weights = iptw, robust = TRUE,
      tt = function(x, t, ...) x * log(pmax(t, 1e-8))
    )),
    M6 = time_once(fit_casebase(
      data.set, iptw, basis = "spline", time.varying = TRUE
    )),
    events = sum(data.set$status)
  ))
}

set.seed(SEED + 7)

timing.raw <- do.call(rbind, lapply(BENCH_N, function(n) {
  cat(sprintf("benchmark n = %d\n", n))
  flush.console()
  timing.matrix <- do.call(rbind, lapply(
    seq_len(BENCH_REPS),
    function(i) benchmark_fits(n)
  ))
  return(data.frame(n = n, t(colMeans(timing.matrix))))
}))

timing.summary <- data.frame(
  n = timing.raw$n,
  events = round(timing.raw$events),
  M1 = timing.raw$M1,
  M2 = timing.raw$M2,
  M3 = timing.raw$M3,
  M4 = timing.raw$M4,
  M5 = timing.raw$M5,
  M6 = timing.raw$M6,
  ratio.M4.M1 = timing.raw$M4 / timing.raw$M1
)

timing.summary

print(
  xtable::xtable(timing.summary, digits = c(0, 0, 0, 3, 3, 3, 3, 3, 3, 2)),
  include.rownames = FALSE, file = "Tables/timing.tex"
)
```

# Session information

```{r}
sessionInfo()
```
