---
title: "Causal Survival - Case Base"
output: pdf_document
date: "2026-05-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newpage
# Imports

### EDA
```{r, warning=FALSE, message=FALSE}
library("tableone")
library("xtable")
```

### Cox PH Model
```{r, warning=FALSE, message=FALSE}
library("survival")
```

### SMD Balance
```{r, warning=FALSE, message=FALSE}
library("survey")
library("cobalt")
```

### Case Base
```{r, warning=FALSE, message=FALSE}
library("casebase")
```

\newpage
# EDA

## Data Set
```{r}
#set seed
set.seed(2026)
```

```{r}
#re-factor the data set

#main variables of interest
time <- veteran$time
Delta <- veteran$status
A <- (veteran$trt - 1)

#covriates
CellType <- veteran$celltype
Karno <- veteran$karno
DiagTime <- veteran$diagtime
Age <- veteran$age
Prior <- (veteran$prior == 10)

#data set design
data.set <- data.frame(
  time = time,
  Delta = Delta,
  A = A,
  
  CellType = CellType,
  Karno = Karno,
  DiagTime = DiagTime,
  Age = Age,
  Prior = Prior
)
```

\newpage
## Table One
```{r, warning=FALSE, message=FALSE}
covariates = c(
  "CellType",
  "Karno",
  "DiagTime",
  "Age",
  "Prior"
)

table.one.original <- CreateTableOne(
  vars = covariates,
  strata = "A",
  data = data.set,
  test = FALSE
)

print(table.one.original, smd = TRUE)
```
\newpage
## Propensity Scores
```{r}
#generate propensity scores
propensity.score <- fitted(
  glm(
    A ~ CellType +
      Karno +
      DiagTime +
      Age +
      Prior,
    data = data.set,
    family = binomial(link = "logit")
  )
)

#separate propensity scores for exposed & unexposed groups
propensity.score.0 <- propensity.score[A == 0]
propensity.score.1 <- propensity.score[A == 1]

#visualize the propensity scores overlap
par(
  mar=c(2,2,2,0)
)

hist(
  propensity.score.0,
  col=rgb(0,0,1,0.5),
  breaks=seq(0,1,by=0.02),
  ylim=c(0,8),
  main="Propensity Score Overlap",
  xlab="Propensity Scores"
)

hist(
  propensity.score.1,
  col=rgb(1,0,0,0.5),
  breaks=seq(0,1,by=0.02),
  add=T
)

box()

legend(
  "topright",
  legend=c('A = 0', 'A = 1'),
  fill=c(rgb(0,0,1,0.5),rgb(1,0,0,0.5)),
  border="black", bty="n"
)
```

\newpage
## Stabilized IPW
```{r}
p.A1 <- mean(A)

#calculate stabilized ipw
ipw <- ifelse(
  A == 1,
  p.A1 / propensity.score,
  (1 - p.A1) / (1 - propensity.score)
)

hist(
  ipw,
  breaks=seq(0,3,by=0.02),
  ylim = c(0, 15),
  main="IPW Distribution",
  xlab="IPWs"
)
box()
```

\newpage
## Covariate Balance
```{r}
#calculate the covariate balance before & after IPW
balance <- bal.tab(
  A ~ CellType +
      Karno +
      DiagTime +
      Age +
      Prior,
  data = data.set,
  weights = ipw,
  method = "weighting",
  estimand = "Average Treatment Effect",
  un = TRUE
)

#visualize the covariate balance before & after IPW
love.plot(
  balance,
  stats = "mean.diffs",
  threshold = 0.1,
  abs = TRUE,
  line = TRUE,
  var.order = "unadjusted",
  colors = c("red", "blue"),
  shapes = c(17, 16),
  limits = c(m = c(0, 0.7)), #hard-coded 0.7??
  title = "Covariate Balance Before and After IPW",
  labels = TRUE
)
```

\newpage
## Table Two
```{r, warning=FALSE, message=FALSE}
data.set.ipw <- svydesign(
  ids = ~ 0,
  data = data.set,
  weights = ipw
)

table.one.ipw <- svyCreateTableOne(
  vars = covariates,
  strata = "A",
  data = data.set.ipw,
  test = FALSE
)

print(table.one.ipw, smd = TRUE)
```
\newpage
# Estimation

## Assumptions
*Non-Informative Censoring*

\newpage
## Model 1: Unadjusted Cox
```{r}
#fit model
model.1 <- coxph(
  Surv(time, Delta) ~ A,
  data = data.set,
  robust = TRUE
)

#generate summary
summary(model.1)
```
\newpage
## Model 2: Covariate Adjusted Cox
```{r}
#fit model
model.2 <- coxph(
  Surv(time, Delta) ~ A +
    CellType +
    Karno +
    DiagTime +
    Age +
    Prior,
  data = data.set,
  robust = TRUE
)

#generate summary
summary(model.2)
```
\newpage
## Model 3: Marginal Structural Cox
```{r}
#fit model
model.3 <- coxph(
  Surv(time, Delta) ~ A,
  data = data.set,
  weights = ipw,
  robust = TRUE
)

#generate summary
summary(model.3)
```
\newpage
## Model 4: IPW Case Base PH
```{r, message=FALSE}
#fit model
model.4 <- fitSmoothHazard(
  Delta ~ time * A,
  data = data.set,
  ratio = 10,
  weights = ipw
)

#generate summary
summary(model.4)
```