---
title: "Causal Survival - Simulation Study"
output: pdf_document
date: "2026-07-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newpage
# Imports
```{r, warning=FALSE, message=FALSE}
library("survival")
library("casebase")
library("sandwich")
library("ggplot2")
library("patchwork")
library("gt")
library("foreach")
library("doParallel")
library("doRNG")
```

\newpage
# Functions

## Data Generation
```{r}
## Generate a simulated survival dataset with baseline confounding, a
## treatment effect, and independent random censoring.
generate.data <- function(
    n,
    r,
    l,
    treatment.predictor.coefs,
    intercept,
    outcome.predictor.coefs,
    beta.A,
    censoring.limit,
    constant.HR = TRUE,
    shape = 1 
) {
  
  # Baseline Covariates
  
  ## Generate 10 baseline covariates from a range of continuous and discrete
  ## distributions. These covariates are subsequently used to generate both
  ## treatment assignment and the event-time distribution.
  X1 <- rnorm(n)
  X2 <- rnorm(n)
  X3 <- runif(n)
  X4 <- runif(n)
  X5 <- rlnorm(n)
  X6 <- rlnorm(n)
  X7 <- rexp(n, rate = r)
  X8 <- rbinom(n, 1, 0.6)
  X9 <- rbinom(n, 1, 0.4)
  X10 <- rpois(n, lambda = l)
  
  ## Combine the baseline covariates into a single design matrix for the
  ## treatment and outcome models
  X <- cbind(
    X1, X2, X3, X4, X5,
    X6, X7, X8, X9, X10
  )
  
  # Input Validation
  
  ## Ensure that the treatment and outcome coefficient vectors contain one
  ## coefficient for each simulated baseline covariate
  stopifnot(length(treatment.predictor.coefs) == ncol(X))
  stopifnot(length(outcome.predictor.coefs) == ncol(X))
  
  ## Ensure that all supplied model coefficients are finite
  stopifnot(all(is.finite(treatment.predictor.coefs)))
  stopifnot(all(is.finite(outcome.predictor.coefs)))
  
  # Treatment Assignment
  
  ## Generate treatment assignment from a logistic model in which all baseline
  ## covariates contribute to treatment assignment, thereby inducing
  ## confounding between treatment and the outcome
  weighted.covariates <- as.vector(
    X %*% treatment.predictor.coefs
  )
  
  linear.predictor.A <- intercept + weighted.covariates
  
  ## Convert the treatment linear predictor to a propensity score and generate
  ## the binary treatment indicator
  A <- rbinom(
    n,
    1,
    plogis(linear.predictor.A)
  )
  
  # Event-Time Model and Censoring
  
  ## Construct the covariate component of the event-time linear predictor.
  ## The treatment coefficient beta.A represents the true causal treatment
  ## effect in the constant-hazard-ratio setting.
  weighted.covariates <- as.vector(
    X %*% outcome.predictor.coefs
  )
  
  ## Generate independent random censoring times from a uniform distribution
  ## bounded by the specified censoring limit
  C <- runif(
    n,
    0,
    censoring.limit
  )
  
  ## Set the baseline event-time linear predictor. The intercept controls the
  ## baseline event-time distribution, while the covariates induce prognostic
  ## heterogeneity across individuals.
  linear.predictor.Y <- -1 + weighted.covariates
  
  T.star <- numeric(n)
  
  # Constant Hazard-Ratio Outcome Model
  
  if (constant.HR) {
    
    ## Add the treatment effect to the event-time linear predictor. Under this
    ## specification, beta.A determines the constant log hazard ratio
    ## associated with treatment.
    linear.predictor.Y <- linear.predictor.Y + (beta.A * A)
    
    ## Generate event times from a Weibull distribution. The scale parameter
    ## depends on the individual-specific linear predictor, allowing both
    ## treatment and baseline covariates to affect event-time risk.
    T.star <- rweibull(
      n,
      shape = shape,
      scale = exp(-linear.predictor.Y / shape)
    )
    
    # Time-Varying Hazard-Ratio Outcome Model
    
  } else {
    
    ## Placeholder for the time-varying treatment-effect specification
    ## (implemented separately when constant.HR = FALSE)
    linear.predictor.Y <- NULL
    
    T.star <- NULL
  }
  
  # Observed Survival Data
  
  ## Determine the observed follow-up time as the minimum of the event and
  ## censoring times
  time <- pmin(
    T.star,
    C
  )
  
  ## Define the event indicator: 1 if the event occurs before or at the
  ## censoring time, and 0 otherwise
  Delta <- as.integer(
    T.star <= C
  )
  
  # Return Simulated Dataset
  
  ## Combine treatment, observed survival outcomes, and baseline covariates
  ## into the final simulated analysis dataset
  data_set <- data.frame(
    time = as.numeric(time),
    Delta = Delta,
    A = A,
    X
  )
  
  return(data_set)
}
```

\newpage
# Modified Case-Base Smooth Hazard Model
```{r}
## Modified implementation of fitSmoothHazard() from the casebase package.
## The function extends the standard case-base fitting procedure to allow
## observation-level weights, including inverse probability weights (IPWs).
FSH <- function(
    formula,
    data,
    time,
    w = rep(1, nrow(data)),
    ratio = 100, ...  
) {
  
  ## Preserve the original function call for reproducibility and method
  ## dispatch in downstream model summaries
  cl <- match.call()
  
  ## Explicitly expand "." in the model formula so that all covariates in the
  ## supplied dataset are represented in the model specification
  formula <- expand.dot.formula(
    formula,
    data = data
  )
  
  ## Extract the event variable from the left-hand side of the model formula
  eventVar <- all.vars(formula[[2]])
  
  ## If the follow-up time variable is not explicitly supplied, infer it from
  ## the data using the event variable and the standard case-base structure
  if (missing(time)) {
    varNames <- checkArgsTimeEvent(
      data = data,
      event = eventVar
    )
    
    timeVar <- varNames$time
    
  } else {
    ## Otherwise, use the time variable explicitly supplied by the user
    timeVar <- time
  }
  
  ## Identify the distinct event types present in the data to determine whether
  ## the analysis involves a single event or competing risks
  typeEvents <- sort(unique(data[[eventVar]]))
  
  
  # Case-Base Sampling
  
  ## Generate the case-base sample when the input data have not already been
  ## sampled. The original dataset is retained so that the supplied observation
  ## weights can be incorporated into the case-base sampling procedure.
  if (!inherits(data, "cbData")) {
    
    originalData <- as.data.frame(data)
    originalData$w <- w
    
    ## Sample cases and controls from the person-time data. The sampling ratio
    ## controls the number of person-moments sampled relative to observed cases.
    sampleData <- sampleCaseBase(
      originalData,
      timeVar,
      eventVar,
      comprisk = (length(typeEvents) > 2),
      ratio
    )
    
  } else {
    ## If the input is already a case-base dataset, retain it directly.
    ## The original unsampled data are unavailable at this stage.
    originalData <- data.frame()
    sampleData <- data
  }
  
  # Weighted Case-Base Model
  
  ## Include the case-base sampling offset in the model formula. This offset
  ## accounts for the sampling mechanism when estimating the hazard function.
  formula <- update(formula, ~ . + offset(offset))
  
  ## Fit a weighted binomial regression model to the case-base sample.
  ## Observation-level weights allow the model to incorporate IPWs in addition
  ## to the case-base sampling structure.
  fittingFunction <- function(formula) glm(
    formula,
    data = sampleData,
    family = binomial,
    weights = w
  )
  
  ## Estimate the smooth hazard model using the weighted case-base sample
  out <- fittingFunction(formula)
  
  # Model Object and Metadata
  
  ## Preserve the lower-level glm call while replacing the displayed call with
  ## the original FSH call for consistency with the custom model interface
  out$lower_call <- out$call 
  out$call <- cl
  
  ## Store the original data and key variable names for downstream methods
  ## such as prediction and plotting
  out$originalData <- originalData
  out$typeEvents <- typeEvents
  out$timeVar <- timeVar
  out$eventVar <- eventVar
  
  ## Record the number of case and background person-moments in the sampled
  ## case-base dataset
  num_pm <- table(sampleData[[eventVar]])
  out$num_cm <- num_pm[2]
  out$num_bm <- num_pm[1]
  out$formula <- formula
  
  ## Store the case-base sampling offset separately, then reset the model's
  ## offset to zero so that absolute risk estimation can be performed without
  ## retaining the sampling correction in the prediction stage
  out$offset <- out$data$offset
  out$data$offset <- 0
  
  ## Assign a custom class so that methods defined for single-event case-base
  ## models can be applied to the fitted object
  class(out) <- c("singleEventCB", class(out))
  
  return(out)
}
```

\newpage
# Formula Expansion Helper
```{r}
## Expand "." in a model formula into the complete set of covariates available
## in the supplied dataset. This ensures that the modified FSH function can
## explicitly evaluate formulas containing "." before fitting the model.
expand.dot.formula <- function(formula, data = NULL) {
  
  ## Check whether the formula contains the "." shorthand for all available
  ## predictors
  if (isTRUE("." %in% all.vars(formula))) {
    
    ## Preserve the original formula attributes while expanding the terms
    att <- attributes(formula)
    
    try.terms <- try(
      stats::terms(formula, data = data),
      silent = TRUE
    )
    
    ## Replace the shorthand formula with its explicitly expanded version when
    ## the terms object can be constructed successfully
    if (!is(try.terms, "try-error")) {
      formula <- formula(try.terms)
    }
    
    ## Restore the original formula attributes after expansion
    attributes(formula) <- att
  }
  
  ## Return the expanded formula
  return(formula)
}
```

\newpage
# Restricted Mean Survival Time (RMST)
```{r}
## Numerically integrate a survival curve over time using the trapezoidal rule.
## The returned vector gives the cumulative RMST evaluated at each time point.
rmst.fn <- function(time, surv) {
  dt <- diff(time)
  mids <- (head(surv, -1) + tail(surv, -1)) / 2
  c(0, cumsum(mids * dt))
}
```

\newpage
# Refactor Simulation Results
```{r}
## Convert the list of simulation results into a variable-wise structure.
## Each element of the returned list contains the corresponding result
## across all simulation replicates.
refactor <- function(result) {
  
  store.by.var <- lapply(
    names(result[[1]]),
    function(x) {
      sapply(result, `[[`, x)
    }
  )
  
  names(store.by.var) <- names(result[[1]])
  
  return(store.by.var)
}
```

\newpage
# Large-Sample Marginal Effect Estimates
```{r}
## Generate a large simulated dataset and estimate the marginal treatment
## effect using an unadjusted Cox model. The resulting estimates provide
## large-sample benchmarks for evaluating the performance of the estimators
## used in the simulation study.
get.marginal.estimates <- function(
    r, l,
    intercept,
    outcome.predictor.coefs,
    beta.A,
    censoring.limit,
    constant.HR
) {
  
  # Generate Large Simulated Dataset
  
  ## Generate a large dataset to approximate the marginal treatment effect.
  ## Setting all treatment-model coefficients to zero removes confounding from
  ## treatment assignment while retaining the covariate effects in the outcome
  ## model.
  dat <- generate.data(
    n                         = 10000000,
    r                         = r,
    l                         = l,
    treatment.predictor.coefs = rep(0, 10),
    intercept                 = intercept,
    outcome.predictor.coefs   = outcome.predictor.coefs,
    beta.A                    = beta.A,
    censoring.limit           = censoring.limit,
    constant.HR               = constant.HR
  )
  
  # Marginal Cox Model
  
  ## Fit an unadjusted Cox model to estimate the marginal treatment effect.
  ## Under the constant-HR specification, the coefficient of A provides the
  ## large-sample marginal log hazard ratio used as the simulation benchmark.
  ms.cox <- NULL
  psi <- NULL
  
  if (constant.HR) {
    
    ms.cox <- coxph(
      Surv(time, Delta) ~ A,
      data = dat
    )
    
    ## Extract the estimated marginal log hazard ratio for treatment
    psi <- coef(ms.cox)["A"]
    
    # Time-Varying Treatment Effect
    
  } else {
    
    ## Under the time-varying specification, allow the treatment effect to
    ## vary with follow-up time through the A-by-time interaction
    ms.cox <- coxph(
      Surv(time, Delta) ~ A * time,
      data = dat
    )
    
    ## No single scalar marginal log hazard ratio is extracted when the
    ## treatment effect varies over time
    psi <- NULL
  }
  
  # Counterfactual Survival Curves
  
  ## Estimate the marginal counterfactual survival curves under no treatment
  ## and treatment by evaluating the fitted model at A = 0 and A = 1
  sf.cox <- survfit(
    ms.cox,
    newdata = data.frame(A = c(0, 1))
  )
  
  ## Interpolate the counterfactual survival curve under A = 0 onto the common
  ## simulation time grid
  s0.cox <- approx(
    sf.cox$time,
    sf.cox$surv[, 1],
    xout   = t.grid,
    method = "constant",
    rule   = 2
  )$y
  
  ## Interpolate the counterfactual survival curve under A = 1 onto the common
  ## simulation time grid
  s1.cox <- approx(
    sf.cox$time,
    sf.cox$surv[, 2],
    xout   = t.grid,
    method = "constant",
    rule   = 2
  )$y
  
  # Marginal RMST Contrast
  
  ## Calculate the marginal RMST contrast by integrating the difference between
  ## the two counterfactual survival curves over the simulation time grid
  output <- list(
    psi  = psi,
    rmst = rmst.fn(t.grid, s1.cox) - rmst.fn(t.grid, s0.cox)
  )
  
  return(output)
}
```

\newpage
# Intermediate Simulation Step
```{r}
## Repeatedly estimate the large-sample marginal treatment effect and RMST
## contrast across independent simulated datasets. The computation is
## parallelized across multiple worker processes to reduce runtime.
intermediate.step <- function(
    r, l, 
    intercept,
    outcome.predictor.coefs,
    beta.A,
    censoring.limit,
    constant.HR
) {
  
  # Parallel Computing Setup
  
  ## Create a two-worker parallel cluster and register it for use by foreach
  cl <- parallel::makeCluster(2)
  doParallel::registerDoParallel(cl)
  
  ## Ensure that the cluster is stopped when the function exits, including if
  ## an error occurs during the simulation
  on.exit(
    parallel::stopCluster(cl),
    add = TRUE
  )
  
  # Repeated Marginal Effect Estimation
  
  ## Repeat the marginal-effect calculation over 100 independently generated
  ## datasets. The random-number-generation option ensures reproducible
  ## results across parallel workers.
  result <- foreach(
    i = 1:100,
    .packages = "survival",
    .export = c(
      "generate.data",
      "get.marginal.estimates",
      "t.grid",
      "rmst.fn"
    ),
    .options.RNG = seed
  ) %dorng% {
    
    ## Generate the large simulated dataset and obtain the corresponding
    ## marginal log hazard ratio and RMST contrast
    get.marginal.estimates(
      r, l,
      intercept,
      outcome.predictor.coefs,
      beta.A,
      censoring.limit,
      constant.HR
    )
  }
  
  # Result Restructuring
  
  ## Convert the list of simulation results into the required structured
  ## format for subsequent simulation analyses
  return(
    refactor(result)
  )
}
```

\newpage
# MS-Cox RMST and Standard Error
```{r}
## Calculate the RMST contrast and its standard error from the marginal
## structural Cox model. The contrast compares restricted mean survival under
## treatment (A = 1) with restricted mean survival under control (A = 0).
cox.rmst.list <- function(
    ms.cox, t.grid, cox.psi.se, survival.curves.data 
) {
  
  ## Estimate the counterfactual survival curves under A = 0 and A = 1
  sf <- survfit(
    ms.cox,
    newdata = survival.curves.data
  )
  
  ## Interpolate the counterfactual survival curve under A = 0 onto the
  ## specified time grid
  S0 <- approx(
    sf$time,
    sf$surv[, 1],
    t.grid,
    method = "constant",
    rule = 2
  )$y
  
  ## Interpolate the counterfactual survival curve under A = 1 onto the
  ## specified time grid
  S1 <- approx(
    sf$time,
    sf$surv[, 2],
    t.grid,
    method = "constant",
    rule = 2
  )$y
  
  # RMST Contrast
  
  ## Calculate the RMST contrast as the difference between the restricted mean
  ## survival under treatment and control
  cox.rmst <- rmst.fn(t.grid, S1) - rmst.fn(t.grid, S0)
  
  # Delta-Method Standard Error
  
  ## Construct the gradient of the RMST functional with respect to the Cox
  ## treatment coefficient. The gradient involves the integral of
  ## S1(t) * log{S1(t)} over the restriction period.
  integrand <- ifelse(
    S1 > 0,
    S1 * log(S1),
    0
  )
  
  g.cox <- rmst.fn(t.grid, integrand)
  
  ## Propagate uncertainty in the estimated treatment coefficient to the RMST
  ## contrast using the delta method
  cox.rmst.se <- abs(g.cox) * cox.psi.se
  
  return(list(
    cox.rmst = cox.rmst,
    cox.rmst.se = cox.rmst.se
  ))
}
```

\newpage
# IPW Case-Base RMST and Standard Error
```{r}
## Calculate the RMST contrast and its standard error from the IPW case-base
## model. The RMST standard error is obtained using a numerical delta method.
cb.rmst.list <- function(
    cb.ipw, cb.var, t.grid, constant.HR
) {
  
  ## Extract the fitted model coefficients and the model terms used to construct
  ## the corresponding design matrix
  beta.hat <- coef(cb.ipw)
  term.form <- delete.response(terms(cb.ipw))
  
  # Counterfactual Survival Function
  
  ## Construct the counterfactual survival function under treatment level a
  ## directly from the fitted case-base model
  cb.surv <- function(beta, a, grid) {
    
    ## Construct prediction data over the specified time grid for treatment
    ## level a
    nd <- data.frame(
      time = grid,
      A = a,
      offset = 0
    )
    
    ## Generate the model matrix using the fitted model's own terms and align
    ## its columns with the estimated coefficient vector
    X <- model.matrix(term.form, data = nd)
    X <- X[, names(beta), drop = FALSE]
    
    ## Obtain the estimated hazard function from the model's linear predictor
    lambda <- exp(as.vector(X %*% beta))
    
    ## Numerically integrate the hazard function using the trapezoidal rule to
    ## obtain the cumulative hazard
    Lambda <- c(
      0,
      cumsum(
        diff(grid) *
          (head(lambda, -1) + tail(lambda, -1)) / 2
      )
    )
    
    ## Convert the cumulative hazard to the corresponding survival function
    exp(-Lambda)
  }
  
  # RMST Contrast
  
  ## Define the RMST contrast as a function of the model coefficient vector.
  ## This function is evaluated at the fitted coefficients and is also used
  ## when calculating the numerical gradient.
  rmst.diff <- function(beta) {
    rmst.fn(t.grid, cb.surv(beta, 1, t.grid)) -
      rmst.fn(t.grid, cb.surv(beta, 0, t.grid))
  }
  
  ## Evaluate the RMST contrast at the fitted model coefficients
  cb.rmst <- rmst.diff(beta.hat)
  
  # Numerical Gradient
  
  ## Calculate the numerical gradient of the RMST contrast with respect to each
  ## fitted model coefficient using a central finite-difference approximation
  p <- length(beta.hat)
  m <- length(t.grid)
  
  grad <- matrix(
    NA_real_,
    nrow = m,
    ncol = p
  )
  
  colnames(grad) <- names(beta.hat)
  
  ## Finite-difference step size
  eps <- 1e-6
  
  ## Perturb each coefficient in turn to approximate its contribution to the
  ## RMST contrast
  for (j in seq_len(p)) {
    bp <- bm <- beta.hat
    
    bp[j] <- bp[j] + eps
    bm[j] <- bm[j] - eps
    
    grad[, j] <- (
      rmst.diff(bp) - rmst.diff(bm)
    ) / (2 * eps)
  }
  
  # Delta-Method Standard Error
  
  ## Align the covariance matrix with the coefficient ordering used by the
  ## numerical gradient before applying the delta-method variance formula
  cb.var <- cb.var[
    colnames(grad),
    colnames(grad),
    drop = FALSE
  ]
  
  ## Calculate the RMST standard error at each restriction time using the
  ## delta-method variance g' V g
  cb.rmst.se <- vapply(
    seq_len(m),
    function(i) {
      g <- grad[i, ]
      
      sqrt(
        drop(
          t(g) %*% cb.var %*% g
        )
      )
    },
    numeric(1)
  )
  
  return(list(
    cb.rmst = cb.rmst,
    cb.rmst.se = cb.rmst.se
  ))
}
```

\newpage
# Single Simulation Replicate
```{r}
## Run one simulation replicate by generating a dataset, estimating stabilized
## inverse probability weights, fitting the MS-Cox and IPW case-base models,
## and calculating the corresponding treatment effects and RMST contrasts.
run.one.sim <- function(
    n, r, l,
    treatment.predictor.coefs,
    intercept,
    outcome.predictor.coefs,
    beta.A,
    censoring.limit,
    constant.HR,
    i
) {
  
  # Generate Simulated Dataset
  
  ## Generate one dataset under the specified data-generating mechanism
  data.set <- generate.data(
    n                         = n,
    r                         = r,
    l                         = l,
    treatment.predictor.coefs = treatment.predictor.coefs,
    intercept                 = intercept,
    outcome.predictor.coefs   = outcome.predictor.coefs,
    beta.A                    = beta.A,
    censoring.limit           = censoring.limit,
    constant.HR               = constant.HR
  )
  
  # Propensity Score Estimation
  
  ## Estimate the probability of treatment conditional on all baseline
  ## covariates using logistic regression
  p.score <- fitted(
    glm(
      A ~ X1 + X2 + X3 + X4 + X5 +
        X6 + X7 + X8 + X9 + X10,
      data = data.set,
      family = binomial(link = "logit")
    )
  )
  
  ## Extract the observed treatment indicator and estimate the marginal
  ## probability of receiving treatment
  A <- data.set$A
  p.A1 <- mean(A)
  
  # Stabilized Inverse Probability Weights
  
  ## Construct stabilized inverse probability weights using the estimated
  ## propensity scores and the marginal probability of treatment
  ipw.stabilized <- ifelse(
    A == 1,
    p.A1 / p.score,
    (1 - p.A1) / (1 - p.score)
  )
  

  # Marginal Structural Cox Model

  ## Initialize objects used to store the Cox model estimates
  ms.cox <- NULL
  cox.psi <- NULL
  cox.psi.se <- NULL
  
  if (constant.HR) {
    
    # Constant Hazard Ratio
    
    ## Fit the marginal structural Cox model using the stabilized IPWs
    ms.cox <- coxph(
      Surv(time, Delta) ~ A,
      data = data.set,
      weights = ipw.stabilized,
      robust = TRUE
    )
    
    ## Extract the estimated marginal log hazard ratio and its robust standard
    ## error
    cox.psi <- coef(ms.cox)["A"]
    cox.psi.se <- sqrt(diag(vcov(ms.cox))["A"])
    
  } else {
    
    # Time-Varying Hazard Ratio
    
    ## Allow the treatment effect to vary over time through the treatment-by-time
    ## interaction
    ms.cox <- coxph(
      Surv(time, Delta) ~ A * time,
      data = data.set,
      weights = ipw.stabilized,
      robust = TRUE
    )
    
    ## A single scalar treatment-effect estimate is not extracted under the
    ## time-varying specification
    cox.psi <- NULL
    cox.psi.se <- NULL
  }
  
  # MS-Cox Counterfactual Survival Curves and RMST
  
  ## Define the two treatment profiles used to obtain counterfactual survival
  ## curves under A = 0 and A = 1
  survival.curves.data <- data.frame(
    A = c(0, 1)
  )
  
  ## Calculate the RMST contrast and its standard error from the fitted
  ## marginal structural Cox model
  cox.rmst.list <- cox.rmst.list(
    ms.cox = ms.cox,
    t.grid = t.grid,
    cox.psi.se = cox.psi.se,
    survival.curves.data = survival.curves.data
  )
  
  ## Extract the RMST contrast and its standard error
  cox.rmst <- cox.rmst.list$cox.rmst
  cox.rmst.se <- cox.rmst.list$cox.rmst.se
  

  # IPW Case-Base Model
  
  ## Set the number of controls sampled per case in the case-base design
  fsh.ratio <- 100
  
  ## Initialize objects used to store the case-base model estimates
  cb.ipw <- NULL
  cb.psi <- NULL
  cb.psi.se <- NULL
  cb.var <- NULL
  
  if (constant.HR) {
    
    # Constant Hazard Ratio
    
    ## Fit the IPW case-base model using the stabilized inverse probability
    ## weights
    cb.ipw <- FSH(
      Delta ~ A,
      data = data.set,
      ratio = fsh.ratio,
      w = ipw.stabilized
    )
    
    ## Extract the estimated marginal log hazard ratio and its
    ## heteroskedasticity-consistent standard error
    cb.psi <- coef(cb.ipw)["A"]
    cb.var <- vcovHC(
      cb.ipw,
      type = "HC0"
    )
    cb.psi.se <- sqrt(diag(cb.var)["A"])
    
  } else {
    
    # Time-Varying Hazard Ratio
    
    ## Fit the IPW case-base model with a treatment-by-time interaction to
    ## allow the treatment effect to vary over follow-up
    cb.ipw <- FSH(
      Delta ~ A * pspline(time, df = 4),
      data = data.set,
      ratio = fsh.ratio,
      w = ipw.stabilized
    )
    
    ## A single scalar treatment-effect estimate and standard error are not
    ## extracted under the time-varying specification
    cb.psi <- NULL
    cb.psi.se <- NULL
    cb.var <- NULL
  }
  
  # IPW Case-Base RMST
  
  ## Calculate the RMST contrast and its standard error from the fitted IPW
  ## case-base model
  cb.rmst.list <- cb.rmst.list(
    cb.ipw,
    cb.var,
    t.grid,
    constant.HR
  )
  
  ## Extract the RMST contrast and its standard error
  cb.rmst <- cb.rmst.list$cb.rmst
  cb.rmst.se <- cb.rmst.list$cb.rmst.se

  # Store Simulation Results
  
  ## Return all treatment-effect and RMST estimates from the current simulation
  ## replicate, together with the replicate identifier
  output <- list(
    simulation.number = i,
    cox.psi = cox.psi,
    cox.psi.se = cox.psi.se,
    cb.psi = cb.psi,
    cb.psi.se = cb.psi.se,
    cox.rmst = cox.rmst,
    cox.rmst.se = cox.rmst.se,
    cb.rmst = cb.rmst,
    cb.rmst.se = cb.rmst.se
  )
  
  return(output)
}
```

\newpage
# Run a Simulation Study
```{r}
## Run the complete simulation study for a specified data-generating mechanism.
## Each simulation replicate independently generates data, estimates stabilized
## IPWs, fits the MS-Cox and IPW case-base models, and calculates the treatment
## effect and RMST estimates.
simulate <- function(
    n, r, l,
    treatment.predictor.coefs,
    intercept,
    outcome.predictor.coefs,
    beta.A,
    censoring.limit,
    constant.HR
) {
  
  # Parallel Processing Setup
  
  ## Allocate approximately 60% of the available CPU cores to the parallel
  ## simulation to balance computational speed with system resource usage
  n.cores <- round(parallel::detectCores() * 0.6)
  
  cl <- parallel::makeCluster(n.cores)
  doParallel::registerDoParallel(cl)
  
  ## Ensure that the parallel cluster is stopped when the function exits,
  ## including if an error occurs during the simulation
  on.exit(
    parallel::stopCluster(cl),
    add = TRUE
  )
  
  # Run Simulation Replicates
  
  ## Run 1,000 independent simulation replicates in parallel. The RNG option
  ## ensures reproducible random-number generation across parallel workers.
  result <- foreach(
    i = 1:1000,
    .packages = c("survival", "casebase"),
    .export = c(
      "run.one.sim",
      "generate.data",
      "vcovHC", "vcov",
      "FSH", "expand.dot.formula",
      "t.grid", "rmst.fn",
      "cox.rmst.list",
      "cb.rmst.list",
      "shape"
    ),
    .options.RNG = seed
  ) %dorng% {
    
    ## Execute a single simulation replicate using the specified
    ## data-generating mechanism and estimation settings
    run.one.sim(
      n,
      r,
      l,
      treatment.predictor.coefs,
      intercept,
      outcome.predictor.coefs,
      beta.A,
      censoring.limit,
      constant.HR,
      i
    )
  }
  
  # Organize Simulation Results
  
  ## Convert the list of replicate-specific results into a compact structure
  ## suitable for summarizing estimator performance across simulations
  return(
    refactor(result)
  )
}
```

\newpage
# RMST Plotting Functions
```{r}
## Plot one or more RMST curves on a common set of axes.
## The function automatically selects the plot title, y-axis limits, and legend
## placement based on the supplied curves.
plot.multiple.rmst <- function(
    titles, rmst.list, cols, main.title = ""
) {
  
  num.types <- length(rmst.list)
  
  # Plot Configuration
  
  ## Generate a default title when one is not explicitly supplied. A single
  ## curve uses its corresponding title, while multiple curves are presented
  ## as an RMST comparison.
  if (main.title == "") {
    main.title <- paste(titles[1], "RMST")
    
    if (num.types > 1) {
      main.title <- "RMST Comparison"
    }
  }
  
  ## Proceed only when the number of curves, labels, and colors agree and the
  ## number of curves is within the supported plotting range
  if (
    num.types == length(titles) &&
    num.types == length(cols) &&
    num.types > 0 &&
    num.types <= 5
  ) {
    
    # Determine Plot Limits and Legend Placement
    
    ## Use the final RMST value to determine an appropriate y-axis range and
    ## position the legend away from the region containing the plotted curves.
    avg.last.point <- mean(
      rmst.list[[1]][-nrow(rmst.list[[1]])]
    )
    
    legend.placement <- "topright"
    
    if (avg.last.point < -0.3) {
      ylim <- c(
        min(rmst.list[[1]][-nrow(rmst.list[[1]])]),
        0
      )
      
      legend.placement <- "bottomleft"
      
    } else if (avg.last.point > 0.3) {
      ylim <- c(
        0,
        max(rmst.list[[1]][-nrow(rmst.list[[1]])])
      )
      
    } else {
      ylim <- c(
        min(rmst.list[[1]][-nrow(rmst.list[[1]])]),
        max(rmst.list[[1]][-nrow(rmst.list[[1]])])
      )
    }
    
    ## Apply transparency to the plotted curves so that overlapping RMST
    ## trajectories remain visually distinguishable
    alpha.cols <- adjustcolor(
      cols,
      alpha.f = 0.05
    )
    
    # Draw RMST Curves
    
    ## Initialize the plot using the first RMST curve
    matplot(
      t.grid,
      rmst.list[[1]],
      type = "l",
      lty = 1,
      col = alpha.cols[1],
      xlab = "Time",
      ylab = "RMST",
      main = main.title
    )
    
    ## Overlay any additional RMST curves on the same axes
    if (length(rmst.list) > 1) {
      for (i in 2:num.types) {
        matplot(
          t.grid,
          rmst.list[[i]],
          type = "l",
          lty = 1,
          col = alpha.cols[i],
          add = TRUE
        )
      }
    }
    
    ## Add a horizontal reference line at zero RMST contrast
    abline(
      h = 0,
      col = "black",
      lwd = 3,
      lty = 2
    )
    
    # Add Legend
    
    legend(
      legend.placement,
      legend = titles,
      col = cols[1:num.types],
      lwd = 2,
      lty = 1
    )
    
  } else {
    
    ## Report an error when the supplied plotting arguments are inconsistent
    cat("error")
    
    if (length(titles) != num.types) {
      cat("mismatch")
    }
  }
}
```

\newpage
# Generate RMST Plots
```{r}
## Generate the complete set of RMST figures used to compare the estimated
## treatment effects from MS-Cox and case-base models against the marginal
## (true) RMST curve.
##
## When save.plots = TRUE, each figure is written to a separate PDF file.
plot.all.rmst <- function(
    true.rmst, cox.rmst, cb.rmst, save.plots = FALSE
) {
  
  # Marginal RMST
  
  ## Plot the marginal RMST curve alone to display the target causal estimand
  if (save.plots) {
    pdf(
      "true-rmst.pdf",
      width = 8,
      height = 5
    )
  }
  
  plot.multiple.rmst(
    titles = c("Marginal"),
    rmst.list = list(true.rmst),
    cols = "purple"
  )
  
  if (save.plots) {
    dev.off()
  }
  
  # MS-Cox vs. Marginal RMST
  
  ## Compare the MS-Cox RMST estimate with the marginal RMST target
  if (save.plots) {
    pdf(
      "cox-rmst.pdf",
      width = 8,
      height = 5
    )
  }
  
  plot.multiple.rmst(
    titles = c(
      "MS-Cox",
      "Marginal"
    ),
    rmst.list = list(
      cox.rmst,
      true.rmst
    ),
    cols = c(
      "red",
      "black"
    ),
    main.title = "MS-Cox RMST"
  )
  
  if (save.plots) {
    dev.off()
  }
  
  # Case-Base vs. Marginal RMST
  
  ## Compare the IPW case-base RMST estimate with the marginal RMST target
  if (save.plots) {
    pdf(
      "cb-rmst.pdf",
      width = 8,
      height = 5
    )
  }
  
  plot.multiple.rmst(
    titles = c(
      "Case Base",
      "Marginal"
    ),
    rmst.list = list(
      cb.rmst,
      true.rmst
    ),
    cols = c(
      "blue",
      "black"
    ),
    main.title = "Case Base RMST"
  )
  
  if (save.plots) {
    dev.off()
  }
  
  # Overall RMST Comparison
  
  ## Compare both estimators simultaneously against the marginal RMST target
  if (save.plots) {
    pdf(
      "all-rmst.pdf",
      width = 8,
      height = 5
    )
  }
  
  plot.multiple.rmst(
    titles = c(
      "MS-Cox",
      "Case Base",
      "Marginal"
    ),
    rmst.list = list(
      cox.rmst,
      cb.rmst,
      true.rmst
    ),
    cols = c(
      "red",
      "blue",
      "black"
    )
  )
  
  if (save.plots) {
    dev.off()
  }
}
```

\newpage
# RMST Simulation Performance Summary
```{r}
## Summarize the finite-sample performance of an RMST estimator across
## simulation replicates. The function calculates the Monte Carlo mean,
## empirical standard deviation, mean estimated standard error, bias, RMSE,
## and confidence-interval coverage at the specified restriction times.
##
## The marginal RMST values are treated as the target estimand against which
## the simulation estimates are evaluated.
get.rmst.summary <- function(
    tau.points,
    rmst.matrix,
    rmst.se.matrix,
    marginal.rmst.means,
    alpha = 0.05
) {
  
  ## Number of simulation replicates and standard-normal critical value for
  ## the nominal (1 - alpha) confidence intervals
  n <- ncol(rmst.matrix)
  z <- qnorm(1 - alpha / 2)
  
  # Point Estimation and Variability
  
  ## Monte Carlo mean of the estimated RMST contrast across simulation
  ## replicates at each selected restriction time
  rmst.means <- rowMeans(
    rmst.matrix[tau.points, ]
  )
  
  ## Empirical standard deviation of the RMST estimates across simulation
  ## replicates, representing the observed Monte Carlo variability
  rmst.sd <- apply(
    rmst.matrix[tau.points, ],
    1,
    sd
  )
  
  ## Average model-based standard error reported by the estimator across
  ## simulation replicates
  rmst.se.mean <- rowMeans(
    rmst.se.matrix[tau.points, ]
  )
  
  # Bias and RMSE
  
  ## Monte Carlo bias relative to the marginal RMST target
  rmst.biases <- rmst.means - marginal.rmst.means
  
  ## Monte Carlo standard error of the estimated bias
  rmst.bias.mcse <- rmst.sd / sqrt(n)
  
  ## Root mean squared error relative to the marginal RMST target, combining
  ## both systematic bias and sampling variability
  rmst.rmse <- sqrt(
    rowMeans(
      (
        rmst.matrix[tau.points, ] -
          marginal.rmst.means
      )^2
    )
  )
  
  # Confidence-Interval Coverage
  
  ## Construct nominal (1 - alpha) Wald confidence intervals for the RMST
  ## contrast using the estimated standard errors from each simulation
  lower <- rmst.matrix[tau.points, ] -
    z * rmst.se.matrix[tau.points, ]
  
  upper <- rmst.matrix[tau.points, ] +
    z * rmst.se.matrix[tau.points, ]
  
  ## Estimate empirical coverage as the proportion of simulation replicates
  ## whose confidence interval contains the marginal RMST target
  coverage <- rowMeans(
    lower <= marginal.rmst.means &
      upper >= marginal.rmst.means
  )
  
  ## Monte Carlo standard error of the estimated coverage probability, treating
  ## coverage as a binomial proportion across simulation replicates
  coverage.mcse <- sqrt(
    coverage * (1 - coverage) / n
  )
  
  # Assemble Summary Table
  
  ## Initialize the summary matrix with one row per restriction time and one
  ## column for each simulation performance metric
  k <- length(tau.points)
  
  rmst.summary <- matrix(
    NA_real_,
    nrow = k,
    ncol = 8
  )
  
  colnames(rmst.summary) <- c(
    "mean",
    "sd",
    "se.mean",
    "bias",
    "bias.mcse",
    "rmse",
    "coverage",
    "coverage.mcse"
  )
  
  ## Populate the summary matrix with the calculated performance measures
  rmst.summary[, "mean"] <- rmst.means
  rmst.summary[, "sd"] <- rmst.sd
  rmst.summary[, "se.mean"] <- rmst.se.mean
  rmst.summary[, "bias"] <- rmst.biases
  rmst.summary[, "bias.mcse"] <- rmst.bias.mcse
  rmst.summary[, "rmse"] <- rmst.rmse
  rmst.summary[, "coverage"] <- coverage
  rmst.summary[, "coverage.mcse"] <- coverage.mcse
  
  return(rmst.summary)
}
```

\newpage
# Print RMST Simulation Summary
```{r}
## Print the simulation performance summary for an RMST estimator in a
## human-readable format. Results are displayed separately for each selected
## restriction time, including the Monte Carlo mean, bias, RMSE, estimated
## standard error, empirical standard deviation, and confidence-interval
## coverage.
print.rmst.summary <- function(
    rmst.summary,
    tau.values,
    model
) {
  
  # Print Model Header
  
  ## Display the name of the estimator being summarized
  cat("\n========================================\n")
  cat(model, "\n")
  cat("========================================\n")
  
  cat("\nRestricted Mean Survival Time Summary\n")
  cat("----------------------------------------\n")
  
  # Print Results at Each Restriction Time
  
  ## Display the simulation performance metrics separately at each selected
  ## restriction time
  for (i in seq_along(tau.values)) {
    
    cat("\nTau =", tau.values[i], "\n")
    cat("----------------------------------------\n")
    
    ## Monte Carlo mean of the estimated RMST contrast
    cat(
      "Mean RMST:",
      rmst.summary[i, "mean"],
      "\n"
    )
    
    ## Monte Carlo bias and its Monte Carlo standard error
    cat(
      "Bias:",
      rmst.summary[i, "bias"],
      " | MC-SE:",
      rmst.summary[i, "bias.mcse"],
      "\n"
    )
    
    ## Root mean squared error of the RMST estimator
    cat(
      "RMSE:",
      rmst.summary[i, "rmse"],
      "\n"
    )
    
    ## Compare the average model-based standard error with the empirical
    ## standard deviation across simulation replicates
    cat(
      "Mean Estimated SE:",
      rmst.summary[i, "se.mean"],
      " | Empirical SD:",
      rmst.summary[i, "sd"],
      "\n"
    )
    
    ## Empirical confidence-interval coverage and its Monte Carlo standard error
    cat(
      "Coverage:",
      rmst.summary[i, "coverage"],
      " | MC-SE:",
      rmst.summary[i, "coverage.mcse"],
      "\n"
    )
  }
  
  cat("\n")
}
```

\newpage
# Print True Marginal RMST
```{r}
## Print the true marginal RMST contrast at each selected restriction time.
## The true value is obtained by averaging the simulated marginal RMST values
## across the corresponding time point.
print.true.rmst <- function(
    true.rmst,
    tau.points,
    tau.values
) {
  
  # Print Header
  
  ## Identify the values being reported as the true marginal RMST
  cat("\n========================================\n")
  cat("\nTrue Mean Restricted Mean Survival Time\n")
  cat("========================================\n")
  
  
  # Print True RMST at Each Restriction Time
  
  ## Display the true marginal RMST corresponding to each selected restriction
  ## time
  for (i in seq_along(tau.values)) {
    cat(
      tau.values[i],
      ":",
      mean(true.rmst[tau.points[i], ]),
      "\n"
    )
  }
  
  cat("\n")
}
```

\newpage
# Plot Hazard Ratio Distributions
```{r}
## Plot the sampling distributions of the estimated hazard ratios from the
## MS-Cox and Case-Base estimators and compare them with the true marginal HR.
plot.hr <- function(
    true.hr,
    cox.hr,
    cb.hr,
    save.plot = FALSE
) {
  
  # Set Up Optional PDF Output
  
  ## Open a PDF device when the plot is requested to be saved
  if (save.plot) {
    pdf("hr.pdf", width = 8, height = 5)
  }
  
  # Plot Estimated HR Distributions
  
  ## Plot the sampling distribution of the MS-Cox HR estimates
  hist(
    cox.hr,
    breaks = 50,
    col = rgb(1, 0, 0, 0.5),
    border = "red",
    xlab = "Hazard Ratio",
    main = "Hazard Ratio Comparison"
  )
  
  ## Overlay the sampling distribution of the Case-Base HR estimates
  hist(
    cb.hr,
    breaks = 50,
    col = rgb(0, 0, 1, 0.5),
    border = "blue",
    add = TRUE
  )
  
  # Add True HR Reference
  
  ## Mark the true marginal HR with a vertical reference line
  abline(
    v = true.hr,
    col = "black",
    lwd = 3,
    lty = 2
  )
  
  # Add Legend
  
  ## Identify the two estimator distributions and the true marginal HR
  legend(
    "topright",
    legend = c(
      "MS-Cox HR",
      "Case Base HR",
      "True HR"
    ),
    fill = c(
      rgb(1, 0, 0, 0.5),
      rgb(0, 0, 1, 0.5),
      NA
    ),
    border = c(
      "red",
      "blue",
      NA
    ),
    lty = c(
      NA,
      NA,
      2
    ),
    lwd = c(
      NA,
      NA,
      3
    ),
    col = c(
      NA,
      NA,
      "black"
    )
  )
  
  # Close Optional PDF Output
  
  ## Close the PDF device after the plot has been written
  if (save.plot) {
    dev.off()
  }
}
```

\newpage
# Plot Hazard Ratio Estimation Errors
```{r}
## Plot the sampling distributions of the estimation errors for the MS-Cox and
## Case-Base HR estimators. An error of zero indicates an estimate equal to the
## true marginal HR.
plot.hr.errors <- function(
    true.hr,
    cox.hr,
    cb.hr,
    save.plot = FALSE
) {
  
  # Calculate HR Estimation Errors
  
  ## Calculate the deviation of each estimated HR from the true marginal HR
  cox.hr.error <- cox.hr - true.hr
  cb.hr.error <- cb.hr - true.hr
  
  # Set Up Optional PDF Output
  
  ## Open a PDF device when the plot is requested to be saved
  if (save.plot) {
    pdf("hr-errors.pdf", width = 8, height = 5)
  }
  
  # Plot HR Error Distributions
  
  ## Plot the sampling distribution of the MS-Cox HR estimation errors
  hist(
    cox.hr.error,
    breaks = 50,
    col = rgb(1, 0, 0, 0.5),
    border = "red",
    xlab = "HR Errors",
    main = "HR Errors Comparison"
  )
  
  ## Overlay the sampling distribution of the Case-Base HR estimation errors
  hist(
    cb.hr.error,
    breaks = 50,
    col = rgb(0, 0, 1, 0.5),
    border = "blue",
    add = TRUE
  )
  
  # Add Zero-Error Reference
  
  ## Mark zero error, corresponding to an estimate exactly equal to the true
  ## marginal HR
  abline(
    v = 0,
    col = "black",
    lwd = 3,
    lty = 2
  )
  
  # Add Legend
  
  ## Identify the estimation-error distributions for the two estimators
  legend(
    "topleft",
    legend = c(
      "MS-Cox HR Error",
      "Case Base HR Error"
    ),
    fill = c(
      rgb(1, 0, 0, 0.5),
      rgb(0, 0, 1, 0.5)
    ),
    border = c(
      "red",
      "blue"
    )
  )
  
  # Close Optional PDF Output
  
  ## Close the PDF device after the plot has been written
  if (save.plot) {
    dev.off()
  }
}
```

\newpage
# Summarize Hazard Ratio Simulation Performance
```{r}
## Calculate simulation performance measures for a hazard ratio estimator,
## including bias, RMSE, average estimated standard error, confidence-interval
## coverage, and either Type I error or power.
get.hr.summary <- function(
    true.hr,
    psi.vector,
    se.vector,
    alpha = 0.05,
    power = TRUE
) {
  
  # Set Up Simulation Quantities
  
  ## Number of simulation replicates and normal critical value for the
  ## confidence intervals
  n <- length(psi.vector)
  z <- qnorm(1 - alpha / 2)
  
  ## Transform log-hazard-ratio estimates to the hazard-ratio scale
  hr.vector <- exp(psi.vector)
  
  # Point-Estimation Performance
  
  ## Monte Carlo mean, empirical standard deviation, and bias of the estimated
  ## hazard ratios
  hr.mean <- exp(mean(psi.vector))
  hr.emp.sd <- sd(hr.vector)
  hr.bias <- hr.mean - true.hr
  
  ## Monte Carlo standard error of the estimated bias
  hr.bias.mcse <- hr.emp.sd / sqrt(n)
  
  ## Mean squared error and root mean squared error relative to the true HR
  hr.mse <- mean((hr.vector - true.hr)^2)
  hr.rmse <- sqrt(hr.mse)
  
  # Standard Error and Confidence-Interval Performance
  
  ## Average model-based standard error across simulation replicates
  hr.se.mean <- mean(se.vector)
  
  ## Construct Wald confidence intervals on the HR scale
  hr.ci.lower <- exp(psi.vector - z * se.vector)
  hr.ci.upper <- exp(psi.vector + z * se.vector)
  
  # Type I Error / Power and Coverage
  
  ## Calculate the proportion of confidence intervals excluding the null HR of
  ## one. This corresponds to Type I error when the true HR equals one and to
  ## empirical power when the true HR differs from one.
  t1.error.or.power <- mean(
    hr.ci.lower > 1 | hr.ci.upper < 1
  )
  
  ## Calculate empirical coverage of the confidence intervals for the true HR
  hr.coverage <- mean(
    hr.ci.lower <= true.hr &
      true.hr <= hr.ci.upper
  )
  
  ## Monte Carlo standard error of the empirical coverage probability
  hr.coverage.mcse <- sqrt(
    hr.coverage * (1 - hr.coverage) / n
  )
  
  # Store Summary Statistics and Confidence Intervals
  
  ## Store the simulation performance measures and individual confidence
  ## intervals in a list for downstream analysis and reporting
  results <- list(
    summary = c(
      mean               = hr.mean,
      emp.sd             = hr.emp.sd,
      bias               = hr.bias,
      bias.mcse          = hr.bias.mcse,
      mse                = hr.mse,
      rmse               = hr.rmse,
      se.mean            = hr.se.mean,
      t1.error.or.power  = t1.error.or.power,
      coverage           = hr.coverage,
      coverage.mcse      = hr.coverage.mcse
    ),
    ci = cbind(
      lower = hr.ci.lower,
      upper = hr.ci.upper
    )
  )
  
  return(results)
}
```

# Print Hazard Ratio Simulation Summary
```{r}
## Print the main simulation performance measures for a hazard ratio estimator
## in a human-readable format.
print.hr.summary <- function(
    hr.stats.summary,
    model
) {
  
  # Print Model Header
  
  ## Identify the estimator corresponding to the reported results
  cat("\n========================================\n")
  cat(model, "\n")
  cat("========================================\n")
  
  cat("\nHazard Ratio Summary\n")
  cat("----------------------------------------\n")
  
  # Print Performance Measures
  
  ## Monte Carlo mean and bias of the estimated hazard ratio
  cat(
    "Mean HR:",
    hr.stats.summary["mean"],
    "\n"
  )
  
  ## Bias and its Monte Carlo standard error
  cat(
    "Bias:",
    hr.stats.summary["bias"],
    " | MC-SE:",
    hr.stats.summary["bias.mcse"],
    "\n"
  )
  
  ## Mean squared error and root mean squared error
  cat(
    "MSE:",
    hr.stats.summary["mse"],
    "\n"
  )
  
  cat(
    "RMSE:",
    hr.stats.summary["rmse"],
    "\n"
  )
  
  ## Compare the average estimated standard error with the empirical standard
  ## deviation across simulation replicates
  cat(
    "Mean Estimated SE:",
    hr.stats.summary["se.mean"],
    " | Empirical SD:",
    hr.stats.summary["emp.sd"],
    "\n"
  )
  
  ## Report the proportion of confidence intervals excluding the null HR of one
  cat(
    "Type I Error/Power:",
    hr.stats.summary["t1.error.or.power"],
    "\n"
  )
  
  ## Report empirical confidence-interval coverage and its Monte Carlo standard
  ## error
  cat(
    "Coverage:",
    hr.stats.summary["coverage"],
    " | MC-SE:",
    hr.stats.summary["coverage.mcse"],
    "\n"
  )
  
  cat("\n")
}
```

\newpage
# Print True Hazard Ratio
```{r}
## Print the true marginal hazard ratio used as the reference value in the
## simulation study.
print.true.hr <- function(true.hr) {
  
  cat("\n========================================\n")
  cat("True HR:", true.hr, "\n")
  cat("========================================\n")
}
```

\newpage
# Calculate and Print HR Simulation Summaries
```{r}
## Calculate and print simulation performance summaries for both the MS-Cox and
## Case-Base estimators using their estimated log-HR coefficients and standard
## errors.
summarize.print.hr <- function(
    true.hr,
    cox.psi, cox.psi.se,
    cb.psi, cb.psi.se
) {
  
  # Calculate Estimator-Specific Summaries
  
  ## Summarize the simulation performance of the MS-Cox hazard ratio estimator
  cox.psi.stats <- get.hr.summary(
    true.hr,
    cox.psi,
    cox.psi.se
  )
  
  ## Summarize the simulation performance of the Case-Base hazard ratio
  ## estimator
  cb.psi.stats <- get.hr.summary(
    true.hr,
    cb.psi,
    cb.psi.se
  )
  
  # Print Estimator Results
  
  ## Print the MS-Cox simulation summary
  print.hr.summary(
    cox.psi.stats$summary,
    "MS-Cox"
  )
  
  ## Print the Case-Base simulation summary
  print.hr.summary(
    cb.psi.stats$summary,
    "Case Base"
  )
  
  # Return Summary Objects
  
  ## Return both estimator-specific results for downstream analysis
  return(list(
    cox.psi.stats,
    cb.psi.stats
  ))
}
```

\newpage
# Plot Hazard Ratio Confidence Intervals
```{r}
## Plot a random sample of simulation-specific hazard ratio estimates and their
## corresponding confidence intervals. The true HR and null value are included
## as reference lines to visually assess estimator accuracy and interval
## coverage.
plot.hr.ci <- function(
    true.value,
    title,
    ci.list,
    hr.list,
    colour = "red",
    coverage = 0,
    seed = 123
) {
  
  # Select Simulation Replicates
  
  ## Set the seed to ensure that the same simulation replicates are selected
  ## each time the plot is generated
  set.seed(seed)
  x = 1:100
  sampled.indices <- sample(1:length(hr.list), 100)
  
  # Construct Confidence-Interval Plot
  
  ## Plot the confidence intervals as shaded bands, with the corresponding
  ## hazard ratio estimates overlaid as individual simulation trajectories
  plot <- ggplot() +
    geom_ribbon(
      aes(
        x = x,
        ymin = ci.list[[1]][sampled.indices],
        ymax = ci.list[[2]][sampled.indices]
      ),
      fill = colour,
      alpha = 0.5,
      show.legend = FALSE
    ) +
    geom_line(
      aes(
        x = x,
        y = hr.list[sampled.indices],
        colour = title
      )
    ) +
    
    ## Add the true marginal HR as a horizontal reference line
    geom_hline(
      aes(
        yintercept = true.value,
        colour = "True HR",
        linetype = "True HR"
      )
    ) +
    
    ## Add the null HR of one as a reference for hypothesis testing
    geom_hline(
      aes(
        yintercept = 1,
        colour = "Null value",
        linetype = "Null value"
      )
    ) +
    
  # Configure Plot Appearanc
  
  ## Assign distinct colours to the estimator, true HR, and null value
  scale_colour_manual(
    name = "",
    values = c(
      setNames(colour, title),
      "True HR" = "black",
      "Null value" = "purple"
    )
  ) +
    
    ## Use distinct line types to differentiate the estimator from the
    ## reference values
    scale_linetype_manual(
      name = "",
      values = c(
        title = "solid",
        "True HR" = "dashed",
        "Null value" = "dotted"
      )
    ) +
    
    ## Label the axes and include the empirical coverage in the plot title
    labs(
      x = "Simulation",
      y = "Hazard ratio",
      title = paste(
        title,
        "HR Confidence Intervals - Coverage:",
        coverage
      )
    ) +
    
    theme_bw()
  
  # Return Plot
  
  ## Return the ggplot object so that it can be displayed, modified, or combined
  ## with other plots
  return(plot)
}
```

\newpage
# Plot HR Confidence Intervals for Both Estimators
```{r}
## Generate and combine confidence-interval plots for the MS-Cox and Case-Base
## estimators, allowing direct visual comparison of their interval coverage and
## estimated hazard ratios across simulation replicates.
plot.all.hr.ci <- function(
    true.hr,
    cox.hr, cox.hr.stats,
    cb.hr, cb.hr.stats,
    seed = 123,
    save.plot = FALSE
) {
  
  # Generate MS-Cox Confidence-Interval Plot
  
  ## Extract the MS-Cox confidence intervals and estimated HRs for plotting
  cox.hr.ci.plot <- plot.hr.ci(
    true.hr,
    title = "MS-Cox",
    list(
      cox.hr.stats$ci[, "lower"],
      cox.hr.stats$ci[, "upper"]
    ),
    cox.hr,
    colour = "red",
    coverage = cox.hr.stats$summary["coverage"],
    seed = seed
  )
  
  # Generate Case-Base Confidence-Interval Plot
  
  ## Extract the Case-Base confidence intervals and estimated HRs for plotting
  cb.hr.ci.plot <- plot.hr.ci(
    true.hr,
    title = "Case Base",
    list(
      cb.hr.stats$ci[, "lower"],
      cb.hr.stats$ci[, "upper"]
    ),
    cb.hr,
    colour = "blue",
    coverage = cb.hr.stats$summary["coverage"],
    seed = seed
  )
  
  # Combine Estimator-Specific Plots
  
  ## Stack the two estimator-specific plots vertically and combine their legends
  combined_plot <- (cox.hr.ci.plot / cb.hr.ci.plot) +
    plot_layout(guides = "collect")
  
  # Optionally Save Combined Plot
  
  ## Save the combined figure as a PDF when requested
  if (save.plot) {
    ggsave(
      "hr-ci.pdf",
      combined_plot,
      width = 8,
      height = 8
    )
  }
}
```

\newpage
# Run a Complete Simulation Study
```{r}
## Run the complete simulation study for a specified sample size, treatment
## effect, and censoring level. The function evaluates both RMST and hazard
## ratio performance for the MS-Cox and Case-Base estimators, with optional
## generation and saving of diagnostic plots.
run.complete.ss <- function(
    n,
    beta.A,
    censoring.limit,
    marginal.estimates,
    generate.plots = FALSE,
    save.plots = FALSE
) {
  
  # Determine Whether Plots Are Required
  
  ## Generate plots when either display or file output has been requested
  need.plots <- generate.plots || save.plots
  
  # Run Simulation Replicates
  
  ## Generate the simulation results for the specified data-generating scenario
  ss.out <- simulate(
    n, r, l,
    treatment.predictor.coefs,
    intercept,
    outcome.predictor.coefs,
    beta.A,
    censoring.limit,
    constant.HR
  )

  # Restricted Mean Survival Time (RMST)

  
  # Plot RMST Results
  
  ## Generate RMST plots comparing the two estimators with the true marginal
  ## RMST curve when plotting has been requested
  if (need.plots) {
    plot.all.rmst(
      marginal.estimates$rmst,
      ss.out$cox.rmst,
      ss.out$cb.rmst,
      save.plots = save.plots
    )
  }
  
  # Select Restriction Times for Summary
  
  ## Identify integer restriction times greater than zero for tabular summaries
  tau.points <- which(
    (t.grid != 0) &
      (t.grid == floor(t.grid))
  )
  
  # Calculate True Marginal RMST Values
  
  ## Calculate the Monte Carlo mean and standard deviation of the marginal RMST
  ## values used as the true reference values
  marginal.rmst.means <- rowMeans(
    marginal.estimates$rmst[tau.points, ]
  )
  
  marginal.rmst.sd <- apply(
    marginal.estimates$rmst[tau.points, ],
    1,
    sd
  )
  
  # Summarize RMST Simulation Performance
  
  ## Calculate simulation performance measures for the MS-Cox RMST estimator
  cox.rmst.summary <- get.rmst.summary(
    tau.points,
    ss.out$cox.rmst,
    ss.out$cox.rmst.se,
    marginal.rmst.means
  )
  
  ## Calculate simulation performance measures for the Case-Base RMST estimator
  cb.rmst.summary <- get.rmst.summary(
    tau.points,
    ss.out$cb.rmst,
    ss.out$cb.rmst.se,
    marginal.rmst.means
  )
  
  ## Define the restriction times reported in the simulation summary
  tau.values <- seq(1, 4, by = 1)
  
  # Print RMST Results
  
  ## Print the true marginal RMST values at the selected restriction times
  print.true.rmst(
    marginal.estimates$rmst,
    tau.points,
    tau.values
  )
  
  ## Print the MS-Cox RMST simulation performance summary
  print.rmst.summary(
    cox.rmst.summary,
    tau.values,
    "MS-Cox RMST"
  )
  
  ## Print the Case-Base RMST simulation performance summary
  print.rmst.summary(
    cb.rmst.summary,
    tau.values,
    "Case Base RMST"
  )

  # Hazard Ratio (HR)

  
  # Determine the True Marginal HR
  
  ## Calculate the true marginal hazard ratio by averaging the simulated
  ## marginal log-HR values and transforming to the HR scale
  true.hr <- exp(mean(marginal.estimates$psi))
  
  # Plot HR Results
  
  if (need.plots) {
    
    ## Plot the distribution of the marginal HR values used to determine the
    ## simulation reference value
    if (save.plots) {
      pdf("true-hr.pdf", width = 8, height = 5)
    }
    
    hist(
      exp(marginal.estimates$psi),
      breaks = 30,
      xlab = "HR",
      main = "Marginal HR"
    )
    
    if (save.plots) {
      dev.off()
    }
    
    ## Compare the sampling distributions of the MS-Cox and Case-Base HR
    ## estimates against the true marginal HR
    plot.hr(
      true.hr,
      exp(ss.out$cox.psi),
      exp(ss.out$cb.psi),
      save.plot = save.plots
    )
    
    ## Compare the estimation errors of the two HR estimators relative to the
    ## true marginal HR
    plot.hr.errors(
      true.hr,
      exp(ss.out$cox.psi),
      exp(ss.out$cb.psi),
      save.plot = save.plots
    )
  }
  
  # Print True HR
  
  ## Report the true marginal HR used as the reference value for evaluating
  ## estimator performance
  print.true.hr(true.hr)
  
  # Summarize HR Simulation Performance
  
  ## Calculate and print simulation performance measures for both HR estimators
  all.hr.stats <- summarize.print.hr(
    true.hr,
    ss.out$cox.psi,
    ss.out$cox.psi.se,
    ss.out$cb.psi,
    ss.out$cb.psi.se
  )
  
  # Plot HR Confidence Intervals
  
  ## Visualize confidence intervals from both estimators across simulation
  ## replicates, including the true HR and null value
  if (need.plots) {
    plot.all.hr.ci(
      true.hr,
      exp(ss.out$cox.psi),
      all.hr.stats[[1]],
      exp(ss.out$cb.psi),
      all.hr.stats[[2]],
      seed = seed,
      save.plot = save.plots
    )
  }
}
```


\newpage
# Simulation Study

## Data Generation Parameters
```{r}
seed <- 123

tau.max <- 4
t.grid <- seq(0, tau.max, by = 0.1)

r = 2
l = 1.5

treatment.predictor.coefs <- c(
  0.7,  -0.4,  0.5,  0.2, 0.6,
  -0.3,  0.1,  -0.2, 0.4, -0.5
)

outcome.predictor.coefs <- c(
  0.6,  -0.3,  0.4,  0.1, 0.5,
  -0.2,  0.05,  -0.1, 0.3, -0.4
)

intercept <- 0.3
constant.HR <- TRUE
shape = 1
```

\newpage
### Setting 1: beta.A = 0, n = 1000 (Null)
```{r}
n1 = 1000
beta.A1 <- 0
censoring.limit.1 <- 5
```

### Setting 2: beta.A = 1, n = 1000 (Strong)
```{r}
n2 = 1000
beta.A2 <- 1
censoring.limit.2 <- 5
```

### Setting 3: beta.A = 0.5, n = 1000 (Moderate)
```{r}
n3 = 1000
beta.A3 <- 0.5
censoring.limit.3 <- 5
```

### Setting 4: beta.A = 0.5, n = 500 (Moderate)
```{r}
n4 = 500
beta.A4 <- 0.5
censoring.limit.4 <- 5
```

### Setting 5: beta.A = 0.5, n = 250 (Moderate)
```{r}
n5 = 250
beta.A5 <- 0.5
censoring.limit.5 <- 5
```

### Setting 6: beta.A = 0.5, n = 1000 (Heavy Censoring)
```{r}
n6 = 1000
beta.A6 <- 0.5
censoring.limit.6 <- 3
```

\newpage
# Extract the marginal estimates for all settings

## Moderate Beta
```{r, cache=TRUE}
true.estimates.moderate <- intermediate.step(
  r, l, 
  intercept,
  outcome.predictor.coefs,
  beta.A3,
  censoring.limit.1,
  constant.HR
)
```

## Null Beta
```{r, cache=TRUE}
true.estimates.null <- intermediate.step(
  r, l, 
  intercept,
  outcome.predictor.coefs,
  beta.A1,
  censoring.limit.1,
  constant.HR
)
```

## Strong Beta
```{r, cache=TRUE}
true.estimates.strong <- intermediate.step(
  r, l, 
  intercept,
  outcome.predictor.coefs,
  beta.A2,
  censoring.limit.1,
  constant.HR
)
```

\newpage
# Run Simulation Study: Setting 1
```{r}
run.complete.ss(
  n1, beta.A1,
  censoring.limit.1,
  true.estimates.null,
  generate.plots = TRUE
)
```

\newpage
# Run Simulation Study: Setting 2
```{r}
run.complete.ss(
  n2, beta.A2,
  censoring.limit.2,
  true.estimates.strong,
  generate.plots = TRUE
)
```

\newpage
# Run Simulation Study: Setting 3
```{r}
run.complete.ss(
  n3, beta.A3, 
  censoring.limit.3, 
  true.estimates.moderate, 
  generate.plots = TRUE
)
```

\newpage
# Run Simulation Study: Setting 4
```{r}
run.complete.ss(
  n4, beta.A4,
  censoring.limit.4,
  true.estimates.moderate
)
```

\newpage
# Run Simulation Study: Setting 5
```{r}
run.complete.ss(
  n5, beta.A5,
  censoring.limit.5,
  true.estimates.moderate
)
```

\newpage
# Run Simulation Study: Setting 6
```{r}
run.complete.ss(
  n6, beta.A6,
  censoring.limit.6,
  true.estimates.moderate
)
```