---
title: "Causal Survival - Case Base"
output: pdf_document
date: "2026-05-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newpage
# Imports

### EDA
```{r, warning=FALSE, message=FALSE}
library("Hmisc")
library("tableone")
library("xtable")
library("ggplot2")
```

### Cox PH Model
```{r, warning=FALSE, message=FALSE}
library("survival")
```

### SMD Balance
```{r, warning=FALSE, message=FALSE}
library("survey")
library("cobalt")
```

### Case Base
```{r, warning=FALSE, message=FALSE}
library("casebase")
```

\newpage
# EDA

## Data Set
```{r}
#set seed
seed = 123

#load the dataset
getHdata(rhc)
```

```{r}
#re-factor the data set

##main variables of interest

###time: last contact/death date - start date
time <- pmax(rhc$dthdte, rhc$lstctdte, na.rm = TRUE) - rhc$sadmdte

###event: death
Delta <- ifelse(
  rhc$death == "Yes",
  1,
  0
)

###treatment: rhc performed or not
A <- ifelse(
  rhc$swang1 == "RHC",
  1,
  0
)

##covriates

###demographics
Gender <- rhc$sex
Age <- as.numeric(rhc$age)
Weight <- as.numeric(rhc$wtkilo1)
Race <- rhc$race
Education <- as.numeric(rhc$edu)
Income <- rhc$income
InsuranceClass <- rhc$ninsclas

###comorbidities
hx_vars <- c(
  "chfhx",
  "cardiohx",
  "dementhx",
  "psychhx",
  "chrpulhx",
  "renalhx",
  "liverhx",
  "gibledhx",
  "malighx",
  "immunhx",
  "transhx",
  "amihx"
)

###convert labelled data to factors
rhc[hx_vars] <- lapply(rhc[hx_vars], function(x) {
  x <- as.numeric(x)

  factor(ifelse(x == 1, "Yes", "No"),
         levels = c("No", "Yes"))
})

CHFhx <- rhc$chfhx
CARDIOhx <- rhc$cardiohx
DEMENThx <- rhc$dementhx
PSYCHhx <- rhc$psychhx
CHRPULhx <- rhc$chrpulhx
RENALhx <- rhc$renalhx
LIVERhx <- rhc$liverhx
GIBLEDhx <- rhc$gibledhx
MALIGhx <- rhc$malighx
IMMUNhx <- rhc$immunhx
TRANShx <- rhc$transhx
AMIhx <- rhc$amihx

#severities
Cancer <- rhc$ca
Category <- rhc$cat1
Severity <- as.numeric(rhc$das2d3pc)

#data set design
data.set <- data.frame(
  time = time,
  Delta = Delta,
  A = A,
  
  Gender = Gender,
  Age = Age,
  Weight = Weight,
  Race = Race,
  Education = Education,
  Income = Income,
  InsuranceClass = InsuranceClass,
  
  CHFhx = CHFhx,
  CARDIOhx = CARDIOhx,
  DEMENThx = DEMENThx,
  PSYCHhx = PSYCHhx,
  CHRPULhx = CHRPULhx,
  RENALhx = RENALhx,
  LIVERhx = LIVERhx,
  GIBLEDhx = GIBLEDhx,
  MALIGhx = MALIGhx,
  IMMUNhx = IMMUNhx,
  TRANShx = TRANShx,
  AMIhx = AMIhx,
  
  Cancer = Cancer,
  Category = Category,
  Severity = Severity
)
```

\newpage
## Table One
```{r}
covariates = c(
  "Gender",
  "Age",
  "Weight",
  "Race",
  "Education",
  "Income",
  "InsuranceClass",
  
  "CHFhx",
  "CARDIOhx",
  "DEMENThx",
  "PSYCHhx",
  "CHRPULhx",
  "RENALhx",
  "LIVERhx",
  "GIBLEDhx",
  "MALIGhx",
  "IMMUNhx",
  "TRANShx",
  "AMIhx",
  
  "Cancer",
  "Category",
  "Severity"
)

table.one.original <- CreateTableOne(
  vars = covariates,
  strata = "A",
  data = data.set,
  test = FALSE
)
print(table.one.original, smd = TRUE)
```


\newpage
## Propensity Scores
```{r}
#set seed for output reproducibility
set.seed(seed)

#generate propensity scores
propensity.score <- fitted(
  glm(
    A ~ Gender +
      Age +
      Weight +
      Race +
      Education +
      Income +
      InsuranceClass +
      CHFhx +
      CARDIOhx +
      DEMENThx +
      PSYCHhx +
      CHRPULhx +
      RENALhx +
      LIVERhx +
      GIBLEDhx +
      MALIGhx +
      IMMUNhx +
      TRANShx +
      AMIhx + 
      Cancer + 
      Category + 
      Severity,
    data = data.set,
    family = binomial(link = "logit")
  )
)

#separate propensity scores for exposed & unexposed groups
propensity.score.0 <- propensity.score[A == 0]
propensity.score.1 <- propensity.score[A == 1]

#visualize the propensity scores overlap
par(
  mar=c(2,2,2,0)
)

hist(
  propensity.score.0,
  col=rgb(0,0,1,0.5),
  breaks=seq(0,1,by=0.02),
  ylim=c(0,350),
  main="Propensity Score Overlap",
  xlab="Propensity Scores"
)

hist(
  propensity.score.1,
  col=rgb(1,0,0,0.5),
  breaks=seq(0,1,by=0.02),
  add=T
)

box()

legend(
  "topright",
  legend=c('A = 0', 'A = 1'),
  fill=c(rgb(0,0,1,0.5),rgb(1,0,0,0.5)),
  border="black", bty="n"
)
```

\newpage
## Stabilized IPW
```{r}
p.A1 <- mean(A)

#calculate unstabilized ipws
ipw.unstabilized <- ifelse(
  A == 1,
  1 / propensity.score,
  1 / (1 - propensity.score)
)

#calculate stabilized ipws
ipw.stabilized <- ifelse(
  A == 1,
  p.A1 / propensity.score,
  (1 - p.A1) / (1 - propensity.score)
)

#visualize the ipws
par(
  mar=c(2,2,2,0)
)

hist(
  ipw.unstabilized,
  col=rgb(0,0,1,0.5),
  breaks=seq(0, max(ipw.unstabilized) + 0.5,by=0.1),
  ylim=c(0,1000),
  main="Unstablized Vs. Stabilized IPWs",
  xlab="IPWs"
)

hist(
  ipw.stabilized,
  col=rgb(1,0,0,0.5),
  breaks=seq(0, max(ipw.stabilized) + 0.5,by=0.1),
  add=T
)

box()

legend(
  "topright",
  legend=c('Unstabilized', 'Stabilized'),
  fill=c(rgb(0,0,1,0.5),rgb(1,0,0,0.5)),
  border="black", bty="n"
)
```

\newpage
## Covariate Balance
```{r, warning=FALSE, message=FALSE}
#calculate the covariate balance before & after IPW
balance <- bal.tab(
  A ~ Gender +
      Age +
      Weight +
      Race +
      Education +
      Income +
      InsuranceClass +
      CHFhx +
      CARDIOhx +
      DEMENThx +
      PSYCHhx +
      CHRPULhx +
      RENALhx +
      LIVERhx +
      GIBLEDhx +
      MALIGhx +
      IMMUNhx +
      TRANShx +
      AMIhx + 
      Cancer + 
      Category + 
      Severity,
  data = data.set,
  weights = ipw.stabilized,
  method = "weighting",
  estimand = "Average Treatment Effect",
  un = TRUE
)

#visualize the covariate balance before & after IPW
love.plot(
  balance,
  stats = "mean.diffs",
  threshold = 0.1,
  abs = TRUE,
  line = TRUE,
  var.order = "unadjusted",
  colors = c("red", "blue"),
  shapes = c(17, 16),
  limits = c(m = c(0, 0.3)),
  title = "Covariate Balance Before and After IPW",
  labels = TRUE
)
```
*Note that covariate balance improved substantially after the inverse probability weighting (IPW) procedure was applied to create the pseudo-population.
*

\newpage
## Table Two
```{r, warning=FALSE, message=FALSE}
data.set.ipw <- svydesign(
  ids = ~ 0,
  data = data.set,
  weights = ipw.stabilized
)

table.one.ipw <- svyCreateTableOne(
  vars = covariates,
  strata = "A",
  data = data.set.ipw,
  test = FALSE
)

print(table.one.ipw, smd = TRUE)
```
\newpage
# Estimation

## Assumptions
*Non-Informative Censoring*

\newpage
## Model 1: Unadjusted Cox
```{r}
#set seed for output reproducibility
set.seed(seed)

#fit model
model.1 <- coxph(
  Surv(time, Delta) ~ A,
  data = data.set,
  robust = TRUE
)

#generate summary
summary(model.1)
```

\newpage
## Model 2: Unadjusted Case Base
```{r}
#set seed for output reproducibility
set.seed(seed)

#fit model
model.2 <- fitSmoothHazard(
  Delta ~ time * A,
  data = data.set,
  ratio = 10
)

#generate summary
summary(model.2)

#HR
cat("HR-hat:", exp(-model.2$coefficients[3]))
```


\newpage
## Model 3: Covariate Adjusted Cox
```{r}
#set seed for output reproducibility
set.seed(seed)

#fit model
model.3 <- coxph(
  Surv(time, Delta) ~ A +
    Gender +
    Age +
    Weight +
    Race +
    Education +
    Income +
    InsuranceClass +
    CHFhx +
    CARDIOhx +
    DEMENThx +
    PSYCHhx +
    CHRPULhx +
    RENALhx +
    LIVERhx +
    GIBLEDhx +
    MALIGhx +
    IMMUNhx +
    TRANShx +
    AMIhx + 
    Cancer + 
    Category + 
    Severity,
  data = data.set,
  robust = TRUE
)

#generate summary
summary(model.3)
```

\newpage
## Model 4: Covariate Adjusted Case Base
```{r}
#set seed for output reproducibility
set.seed(seed)

#fit model
model.4 <- fitSmoothHazard(
  Delta ~ time * A +
    Gender +
    Age +
    Weight +
    Race +
    Education +
    Income +
    InsuranceClass +
    CHFhx +
    CARDIOhx +
    DEMENThx +
    PSYCHhx +
    CHRPULhx +
    RENALhx +
    LIVERhx +
    GIBLEDhx +
    MALIGhx +
    IMMUNhx +
    TRANShx +
    AMIhx + 
    Cancer + 
    Category + 
    Severity,
  data = data.set,
  ratio = 10
)

#generate summary
summary(model.4)

#HR
cat("HR-hat:", exp(-model.4$coefficients[3]))
```

\newpage
## Model 5: Marginal Structural Cox
```{r}
#set seed for output reproducibility
set.seed(seed)

#fit model
model.5 <- coxph(
  Surv(time, Delta) ~ A,
  data = data.set,
  weights = ipw.stabilized,
  robust = TRUE
)

#generate summary
summary(model.5)
```

\newpage
## Model 6: IPW Case Base PH
```{r, message=FALSE}
#set seed for output reproducibility
set.seed(seed)

#fit model
model.6 <- fitSmoothHazard(
  Delta ~ time * A,
  data = data.set,
  ratio = 10,
  family = "glm",
  weights = ipw.stabilized
)

#generate summary
summary(model.6)

#HR
cat("HR-hat:", exp(-model.6$coefficients[3]))
```

\newpage
## Model 7: IPW Case Base (Variable Hazard over Time)
```{r, message=FALSE}
#set seed for output reproducibility
set.seed(seed)

#fit model
model.7 <- fitSmoothHazard(
  Delta ~ pspline(time, df = 2) * A,
  data = data.set,
  ratio = 10,
  weights = ipw.stabilized
)

#generate summary
summary(model.7)

#plot hazard-time
plot.time <- seq(
  min(data.set$time, na.rm = TRUE),
  max(data.set$time, na.rm = TRUE),
  by = 0.1
)

plot.data <- data.frame(
  A = factor(
    "No RHC",
    levels = c("No RHC", "RHC")
  ),
  time = plot.time
)

plot(
  model.7,
  type = "hr",
  newdata = plot.data,
  var = "A",
  xvar = "time",
  ci = TRUE,
  ylim = c(0,2)
)
```