---
title: "Causal Survival - Case Base"
output: pdf_document
date: "2026-05-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newpage
# Imports

### EDA
```{r, warning=FALSE, message=FALSE}
library("Hmisc")
library("tableone")
library("xtable")
library("ggplot2")
library("survey")
library("cobalt")
```

### Estimation
```{r, warning=FALSE, message=FALSE}
library("survival")
library("casebase")
library("sandwich")
library("foreach")
library("doParallel")
```

\newpage
# EDA

## Data Set
```{r}
# set seed
s = 123

# load the dataset
getHdata(rhc)
```

```{r}
# re-factor the data set

## main variables of interest

### time: last contact/death date - start date
time <- as.numeric(pmax(rhc$dthdte, rhc$lstctdte, na.rm = TRUE) - rhc$sadmdte)

### event: death
Delta <- ifelse(
  rhc$death == "Yes",
  1,
  0
)

### treatment: rhc performed or not
A <- ifelse(
  rhc$swang1 == "RHC",
  1,
  0
)

## covariates

### demographics
Gender <- rhc$sex
Age <- as.numeric(rhc$age)

# contains 0 instead of NA
# changing 0 to NA completely messes up propensity scores 
# and covariate balance after weighting
Weight <- as.numeric(rhc$wtkilo1)

Race <- droplevels(rhc$race)
Education <- as.numeric(rhc$edu)
Income <- rhc$income
InsuranceClass <- rhc$ninsclas

### comorbidities
hx.vars <- c(
  "chfhx",
  "cardiohx",
  "dementhx",
  "psychhx",
  "chrpulhx",
  "renalhx",
  "liverhx",
  "gibledhx",
  "malighx",
  "immunhx",
  "transhx",
  "amihx"
)

### convert labelled data to factors
rhc[hx.vars] <- lapply(rhc[hx.vars], function(x) {
  x <- as.numeric(x)

  factor(ifelse(x == 1, "Yes", "No"),
         levels = c("No", "Yes"))
})

CHFhx <- rhc$chfhx
CARDIOhx <- rhc$cardiohx
DEMENThx <- rhc$dementhx
PSYCHhx <- rhc$psychhx
CHRPULhx <- rhc$chrpulhx
RENALhx <- rhc$renalhx
LIVERhx <- rhc$liverhx
GIBLEDhx <- rhc$gibledhx
MALIGhx <- rhc$malighx
IMMUNhx <- rhc$immunhx
TRANShx <- rhc$transhx
AMIhx <- rhc$amihx

# severities
Cancer <- rhc$ca
Category <- droplevels(rhc$cat1)
Severity <- as.numeric(rhc$das2d3pc)

# data set design
data.set <- data.frame(
  time = time,
  Delta = Delta,
  A = A,
  
  Gender = Gender,
  Age = Age,
  Weight = Weight,
  Race = Race,
  Education = Education,
  Income = Income,
  InsuranceClass = InsuranceClass,
  
  CHFhx = CHFhx,
  CARDIOhx = CARDIOhx,
  DEMENThx = DEMENThx,
  PSYCHhx = PSYCHhx,
  CHRPULhx = CHRPULhx,
  RENALhx = RENALhx,
  LIVERhx = LIVERhx,
  GIBLEDhx = GIBLEDhx,
  MALIGhx = MALIGhx,
  IMMUNhx = IMMUNhx,
  TRANShx = TRANShx,
  AMIhx = AMIhx,
  
  Cancer = Cancer,
  Category = Category,
  Severity = Severity
)
```

\newpage
## Table One
```{r}
covariates = c(
  "Gender",
  "Age",
  "Weight",
  "Race",
  "Education",
  "Income",
  "InsuranceClass",
  
  "CHFhx",
  "CARDIOhx",
  "DEMENThx",
  "PSYCHhx",
  "CHRPULhx",
  "RENALhx",
  "LIVERhx",
  "GIBLEDhx",
  "MALIGhx",
  "IMMUNhx",
  "TRANShx",
  "AMIhx",
  
  "Cancer",
  "Category",
  "Severity"
)

table.one.original <- CreateTableOne(
  vars = covariates,
  strata = "A",
  data = data.set,
  test = FALSE
)
print(table.one.original, smd = TRUE)
```


\newpage
## Propensity Scores
```{r}
# set seed for output reproducibility
set.seed(s)

# generate propensity scores
propensity.score <- fitted(
  glm(
    A ~ Gender +
      Age +
      Weight +
      Race +
      Education +
      Income +
      InsuranceClass +
      CHFhx +
      CARDIOhx +
      DEMENThx +
      PSYCHhx +
      CHRPULhx +
      RENALhx +
      LIVERhx +
      GIBLEDhx +
      MALIGhx +
      IMMUNhx +
      TRANShx +
      AMIhx + 
      Cancer + 
      Category + 
      Severity,
    data = data.set,
    family = binomial(link = "logit")
  )
)

# separate propensity scores for exposed & unexposed groups
propensity.score.0 <- propensity.score[A == 0]
propensity.score.1 <- propensity.score[A == 1]

# visualize the propensity scores overlap
par(
  mar=c(2,2,2,0)
)

hist(
  propensity.score.0,
  col=rgb(0,0,1,0.5),
  breaks=seq(0,1,by=0.02),
  ylim=c(0,200),
  main="Propensity Score Overlap",
  xlab="Propensity Scores"
)

hist(
  propensity.score.1,
  col=rgb(1,0,0,0.5),
  breaks=seq(0,1,by=0.02),
  add=T
)

box()

legend(
  "topright",
  legend=c('A = 0', 'A = 1'),
  fill=c(rgb(0,0,1,0.5),rgb(1,0,0,0.5)),
  border="black",
  bty="n"
)
```

\newpage
## Stabilized IPW
```{r}
p.A1 <- mean(A)

# calculate unstabilized ipws
ipw.unstabilized <- ifelse(
  A == 1,
  1 / propensity.score,
  1 / (1 - propensity.score)
)

# calculate stabilized ipws
ipw.stabilized <- ifelse(
  A == 1,
  p.A1 / propensity.score,
  (1 - p.A1) / (1 - propensity.score)
)

# visualize the ipws
par(
  mar=c(2,2,2,0)
)

hist(
  ipw.unstabilized,
  col=rgb(0,0,1,0.5),
  breaks=seq(0, max(ipw.unstabilized) + 0.5,by=0.1),
  ylim=c(0,1000),
  main="Unstablized Vs. Stabilized IPWs",
  xlab="IPWs"
)

hist(
  ipw.stabilized,
  col=rgb(1,0,0,0.5),
  breaks=seq(0, max(ipw.stabilized) + 0.5,by=0.1),
  add=T
)

box()

legend(
  "topright",
  legend=c('Unstabilized', 'Stabilized'),
  fill=c(rgb(0,0,1,0.5),rgb(1,0,0,0.5)),
  border="black", bty="n"
)

# ipw.stabilized is better so we augment that
data.set$ipw <- ipw.stabilized
```

\newpage
## Covariate Balance
```{r, warning=FALSE, message=FALSE}
# calculate the covariate balance before & after IPW
balance <- bal.tab(
  A ~ Gender +
      Age +
      Weight +
      Race +
      Education +
      Income +
      InsuranceClass +
      CHFhx +
      CARDIOhx +
      DEMENThx +
      PSYCHhx +
      CHRPULhx +
      RENALhx +
      LIVERhx +
      GIBLEDhx +
      MALIGhx +
      IMMUNhx +
      TRANShx +
      AMIhx + 
      Cancer + 
      Category + 
      Severity,
  data = data.set,
  weights = ipw.stabilized,
  method = "weighting",
  estimand = "Average Treatment Effect",
  un = TRUE
)

# visualize the covariate balance before & after IPW
love.plot(
  balance,
  stats = "mean.diffs",
  threshold = 0.1,
  abs = TRUE,
  line = TRUE,
  var.order = "unadjusted",
  colors = c("red", "blue"),
  shapes = c(17, 16),
  limits = c(m = c(0, 0.3)),
  title = "Covariate Balance Before and After IPW",
  labels = TRUE
)
```
*Note that covariate balance improved substantially after the inverse probability weighting (IPW) procedure was applied to create the pseudo-population.
*

\newpage
## Table Two
```{r, warning=FALSE, message=FALSE}
data.set.ipw <- svydesign(
  ids = ~ 0,
  data = data.set,
  weights = ipw.stabilized
)

table.one.ipw <- svyCreateTableOne(
  vars = covariates,
  strata = "A",
  data = data.set.ipw,
  test = FALSE
)

print(table.one.ipw, smd = TRUE)
```
\newpage
# Estimation

## Assumptions
*Non-Informative Censoring*

\newpage
##FSH: Modified Definition
```{r, include=FALSE}
FSH <- function(
    formula,
    data,
    time,
    w = rep(1, nrow(data)),
    ratio = 100, ...  
  ) {
  
  cl <- match.call()
  
  formula <- expand.dot.formula(
    formula,
    data = data
  )
  
  # Infer name of event variable from LHS of formula
  eventVar <- all.vars(formula[[2]])

  if (missing(time)) {
    varNames <- checkArgsTimeEvent(
      data = data,
      event = eventVar
    )
    
    timeVar <- varNames$time
  
  } else {
    timeVar <- time
  }

  typeEvents <- sort(unique(data[[eventVar]]))
  
  # Call sampleCaseBase if class is not cbData
  if (!inherits(data, "cbData")) {
  
    originalData <- as.data.frame(data)
    originalData$w <- w
    
    sampleData <- sampleCaseBase(
      originalData,
      timeVar,
      eventVar,
      comprisk = (length(typeEvents) > 2),
      ratio
    )
  
  } else {
    # If class is cbData we no longer have the original data
    originalData <- data.frame() # Empty data.frame
    sampleData <- data
  }

  formula <- update(formula, ~ . + offset(offset))

  # Fit a binomial model because there are no competing risks
  fittingFunction <- function(formula) glm(
    formula,
    data = sampleData,
    family = binomial,
    weights = w
  )

  out <- fittingFunction(formula)
  
  # Save lower call for plot method
  out$lower_call <- out$call 
  out$call <- cl
  
  out$originalData <- originalData
  out$typeEvents <- typeEvents
  out$timeVar <- timeVar
  out$eventVar <- eventVar
  
  # Count person-moment types
  num_pm <- table(sampleData[[eventVar]])
  out$num_cm <- num_pm[2]
  out$num_bm <- num_pm[1]
  out$formula <- formula
  
  # Reset offset for absolute risk estimation, but keep track of it
  out$offset <- out$data$offset
  out$data$offset <- 0
  
  # Add new class
  class(out) <- c("singleEventCB", class(out))
  
  return(out)
}



expand.dot.formula <- function(formula, data = NULL) {
  if (isTRUE("." %in% all.vars(formula))) {
    att <- attributes(formula)
    try.terms <- try(
      stats::terms(formula, data = data),
      silent = TRUE
    )
    if (!is(try.terms, "try-error")) {
      formula <- formula(try.terms)
    }
    attributes(formula) <- att
  }
  formula
}
```

\newpage
## Model 1: Unadjusted Cox
```{r}
# set seed for output reproducibility
set.seed(s)

# fit model
model.1 <- coxph(
  Surv(time, Delta) ~ A,
  data = data.set,
  robust = TRUE
)

# generate summary
summary(model.1)
```

\newpage
## Case Base Setup
```{r}
# set a high ratio for all casebase models
fsh.ratio <- 100
```

## Model 2: Unadjusted Case Base
```{r, warning=FALSE, message=FALSE}
# set seed for output reproducibility
set.seed(s)

# fit model
model.2 <- fitSmoothHazard(
  Delta ~  A,
  data = data.set,
  ratio = fsh.ratio
)

# generate summary
summary(model.2)

# HR
beta.2 <- coef(model.2)["A"]
se.2 <- sqrt(diag(vcovHC(model.2))["A"])
cat("HR-hat:", exp(beta.2), "\n") #glm
cat("SE of HR-hat:", se.2, "\n")
cat("95% CI: (", exp(beta.2 + c(-1,1)*1.96*se.2), ")\n")
```


\newpage
## Model 3: Covariate Adjusted Cox
```{r}
# set seed for output reproducibility
set.seed(s)

# fit model
model.3 <- coxph(
  Surv(time, Delta) ~ A +
    Gender +
    Age +
    Weight +
    Race +
    Education +
    Income +
    InsuranceClass +
    CHFhx +
    CARDIOhx +
    DEMENThx +
    PSYCHhx +
    CHRPULhx +
    RENALhx +
    LIVERhx +
    GIBLEDhx +
    MALIGhx +
    IMMUNhx +
    TRANShx +
    AMIhx + 
    Cancer + 
    Category + 
    Severity,
  data = data.set,
  robust = TRUE
)

# generate summary
summary(model.3)
```

\newpage
## Model 4: Covariate Adjusted Case Base
```{r, warning=FALSE, message=FALSE}
# set seed for output reproducibility
set.seed(s)

# fit model
model.4 <- fitSmoothHazard(
  Delta ~ A +
    Gender +
    Age +
    Weight +
    Race +
    Education +
    Income +
    InsuranceClass +
    CHFhx +
    CARDIOhx +
    DEMENThx +
    PSYCHhx +
    CHRPULhx +
    RENALhx +
    LIVERhx +
    GIBLEDhx +
    MALIGhx +
    IMMUNhx +
    TRANShx +
    AMIhx + 
    Cancer + 
    Category + 
    Severity,
  data = data.set,
  ratio = fsh.ratio
)

# generate summary
summary(model.4)

# HR
beta.4 <- coef(model.4)["A"]
se.4 <- sqrt(diag(vcovHC(model.4))["A"])
cat("HR-hat:", exp(beta.4), "\n") #glm
cat("SE of HR-hat:", se.4, "\n")
cat("95% CI: (", exp(beta.4 + c(-1,1)*1.96*se.4), ")\n")
```

\newpage
## Model 5: Marginal Structural Cox
```{r}
# set seed for output reproducibility
set.seed(s)

# fit model
model.5 <- coxph(
  Surv(time, Delta) ~ A,
  data = data.set,
  weights = ipw.stabilized,
  robust = TRUE
)

# generate summary
summary(model.5)
```

\newpage
## Model 6: IPW Case Base PH
```{r, warning=FALSE, message=FALSE}
# set seed for output reproducibility
set.seed(s)

# fit model
model.6 <- FSH(
  Delta ~ A,
  data = data.set,
  ratio = fsh.ratio,
  w = ipw.stabilized
)

# generate summary
summary(model.6)

# HR
beta.6 <- coef(model.6)["A"]
se.6 <- sqrt(diag(vcovHC(model.6))["A"])
cat("HR-hat:", exp(beta.6), "\n") #glm
cat("SE of HR-hat:", se.6, "\n")
cat("95% CI: (", exp(beta.6 + c(-1,1)*1.96*se.6), ")\n")
```

\newpage
## Model 7: Unadjusted Case Base (Variable HR)

### Generate Model Summary
```{r, warning=FALSE, message=FALSE}
# set seed for output reproducibility
set.seed(s)

# fit model
model.7 <- FSH(
  Delta ~ pspline(time, df = 2) * A,
  data = data.set,
  ratio = fsh.ratio
)

# generate summary
summary(model.7)
```

### Setup for Survival Curve
```{r}
trt.types <- c(0, 1)
individual.profile.data <- data.frame(A = 1)
survival.curves.data <- data.frame(A = trt.types)

tau.max <- max(data.set$time, na.rm = TRUE)
t.max <- 150 # First 4 months

tau.grid <- seq(0, tau.max, by = 1)
t.grid <- seq(0, t.max, by = 1)
```

### Plot variable HR
```{r, message=FALSE}
# visualize the hazard ratio over time
par(mfrow = c(2, 1))

plot(
  model.7,
  type = "hr",
  newdata = expand.grid(
    time = tau.grid,
    A = trt.types
  ),
  var = "A",
  xvar = "time",
  ci = TRUE,
  ylim = c(0,2),
  xlab="Time",
  ylab="Hazard Ratio",
  main="Model 7: Variable Hazard Ratio"
)
abline(h = 1, col = "red", lty = 2)


plot(
  model.7,
  type = "hr",
  newdata = expand.grid(
    time = t.grid,
    A = trt.types
  ),
  var = "A",
  xvar = "time",
  ci = TRUE,
  ylim = c(1,1.5),
  xlab="Time",
  ylab="Hazard Ratio",
  main="Model 7: Variable Hazard Ratio - First 4 Months"
)
abline(h = 1, col = "red", lty = 2)
```

\newpage
## Model 8: IPW Case Base (Variable HR)

### Generate Model Summary
```{r, warning=FALSE, message=FALSE}
# set seed for output reproducibility
set.seed(s)

# fit model
model.8 <- FSH(
  Delta ~ pspline(time, df = 2) * A,
  data = data.set,
  ratio = fsh.ratio,
  w = ipw.stabilized
)

# generate summary
summary(model.8)
```

### Plot variable HR
```{r, message=FALSE}
# visualize the hazard ratio over time
par(mfrow = c(2, 1))

plot(
  model.8,
  type = "hr",
  newdata = expand.grid(
    time = tau.grid,
    A = trt.types
  ),
  var = "A",
  xvar = "time",
  ci = TRUE,
  ylim = c(0,2),
  xlab="Time",
  ylab="Hazard Ratio",
  main="Model 8: Variable Hazard Ratio"
)
abline(h = 1, col = "red", lty = 2)

plot(
  model.8,
  type = "hr",
  newdata = expand.grid(
    time = t.grid,
    A = trt.types
  ),
  var = "A",
  xvar = "time",
  ci = TRUE,
  ylim = c(1,1.5),
  xlab="Time",
  ylab="Hazard Ratio",
  main="Model 8: Variable Hazard Ratio - First 4 Months"
)
abline(h = 1, col = "red", lty = 2)
```

\newpage
# Survival Curves using MS-Cox and IPW Case Base Models

## SetUp
```{r}
# colors: blue for MS-Cox, red for case base (with matching transparent fills)
col.cox <- "blue"
col.cb  <- "red"

fill.cox <- rgb(0, 0, 1, 0.20)
fill.cb  <- rgb(1, 0, 0, 0.20)
```


## Generate MS-Cox Survival Curve
```{r}
sf.cox.ind <- survfit(
  model.5,
  newdata = individual.profile.data
)

S.cox <- approx(
  sf.cox.ind$time,
  sf.cox.ind$surv,
  xout = tau.grid,
  method = "constant",
  rule = 2
)$y

Lo.cox <- approx(
  sf.cox.ind$time,
  sf.cox.ind$lower,
  xout = tau.grid,
  method = "constant",
  rule = 2
)$y

Up.cox <- approx(
  sf.cox.ind$time,
  sf.cox.ind$upper,
  xout = tau.grid,
  method = "constant",
  rule = 2
)$y
```

## Generate Case Base Survival Curve
```{r, message=FALSE}
set.seed(s)
n.boot <- 200

boot.surv <- replicate(n.boot, {
  idx <- sample(nrow(data.set), replace = TRUE)
  boot.data <- data.set[idx, ]

  boot.fit <- FSH(
    Delta ~ pspline(time, df = 2) * A,
    data = boot.data,
    ratio = 10,
    weights = boot.data$ipw
  )

  ar <- absoluteRisk(
    boot.fit,
    time = tau.grid,
    newdata = individual.profile.data,
    type = "survival"
  )
  ar[, 2]
})

S.cb  <- absoluteRisk(
  model.8,
  time = tau.grid,
  newdata = individual.profile.data,
  type = "survival"
)[, 2]

Lo.cb <- apply(
  boot.surv,
  1,
  quantile,
  probs = 0.025,
  na.rm = TRUE
)

Up.cb <- apply(
  boot.surv,
  1,
  quantile,
  probs = 0.975,
  na.rm = TRUE
)
```

## Plot Survival Curves
```{r}
par(mfrow = c(2, 1))

plot(
  tau.grid,
  S.cox,
  type = "n",
  ylim = c(0, 1),
  xlab = "Days",
  ylab = "Survival Probability",
  main = "Estimated Survival Probability After RHC"
)

# shaded bands first, lines on top
polygon(
  c(tau.grid, rev(tau.grid)),
  c(Lo.cox, rev(Up.cox)),
  col = fill.cox,
  border = NA
)

polygon(
  c(tau.grid, rev(tau.grid)),
  c(Lo.cb, rev(Up.cb)),
  col = fill.cb,
  border = NA
)

lines(tau.grid, S.cox, col = col.cox, lwd = 2, type = "s")
lines(tau.grid, S.cb,  col = col.cb,  lwd = 2)

legend(
  "topright",
  legend = c(
    "semi-parametric (MS-Cox)",
    "parametric (IPW Case Base)"
  ),
  col = c(col.cox, col.cb),
  lty = c(1, 1),
  bg = "gray90"
)

plot(
  tau.grid,
  S.cox,
  type = "n",
  ylim = c(0.6, 1),
  xlim = c(0, 30),
  xlab = "Days",
  ylab = "Survival Probability",
  main = "Estimated Survival Probability After RHC - First Month"
)

# shaded bands first, lines on top
polygon(
  c(tau.grid, rev(tau.grid)),
  c(Lo.cox, rev(Up.cox)),
  col = fill.cox,
  border = NA
)

polygon(
  c(tau.grid, rev(tau.grid)),
  c(Lo.cb, rev(Up.cb)),
  col = fill.cb,
  border = NA
)

lines(tau.grid, S.cox, col = col.cox, lwd = 2, type = "s")
lines(tau.grid, S.cb,  col = col.cb,  lwd = 2)

legend(
  "topright",
  legend = c(
    "semi-parametric (MS-Cox)",
    "parametric (IPW Case Base)"
  ),
  col = c(col.cox, col.cb),
  lty = c(1, 1),
  bg = "gray90"
)
```

\newpage
# RMST Function
```{r}
rmst.fn <- function(time, surv) {
  dt <- diff(time)
  mids <- (head(surv, -1) + tail(surv, -1)) / 2
  c(0, cumsum(mids * dt))
}
```

\newpage
# MS-Cox PH Counterfactual Survival Curves

### Extract MS-Cox Counterfactual Survival Curves
```{r}
sf.cox <- survfit(
  model.5,
  newdata = survival.curves.data
)

S0.cox <- approx(
  sf.cox$time,
  sf.cox$surv[, 1],
  xout = tau.grid,
  method = "constant",
  rule = 2
)$y

S1.cox <- approx(
  sf.cox$time,
  sf.cox$surv[, 2],
  xout = tau.grid,
  method = "constant",
  rule = 2
)$y
```

```{r}
plot(sf.cox, col = c(1,2))
```

### Plot MS-Cox Counterfactual Survival Curves
```{r}
par(mfrow = c(2, 1))

plot(
  tau.grid,
  S0.cox,
  type = "s", 
  lwd = 2,
  col = "red",
  ylim = c(0, 1),
  xlab = "Days",
  ylab = "Survival Probability",
  main = "MS-Cox: Counterfactual Survival Curves"
)

lines(
  tau.grid,
  S1.cox,
  type = "s",
  lwd = 2, 
  col = "blue"
)

legend(
  "topright",
  legend = c("A = 0", "A = 1"),
  col = c("red", "blue"),
  lty = 1,
  lwd = 2
)

plot(
  tau.grid,
  S0.cox,
  type = "s", 
  lwd = 2,
  col = "red",
  ylim = c(0.6, 1),
  xlim = c(0, 30),
  xlab = "Days",
  ylab = "Survival Probability",
  main = "MS-Cox: Counterfactual Survival Curves - First Month"
)

lines(
  tau.grid,
  S1.cox,
  type = "s",
  lwd = 2, 
  col = "blue"
)

legend(
  "topright",
  legend = c("A = 0", "A = 1"),
  col = c("red", "blue"),
  lty = 1,
  lwd = 2
)
```

\newpage
# IPW Case Base Counterfactual Survival Curves

### Extract CB Counterfactual Survival Curves

```{r, warning=FALSE}
Sf.cb <- absoluteRisk(
  object = model.8,
  time = tau.grid,
  newdata = survival.curves.data,
  type = "survival"
)

S0.cb <- Sf.cb[, 2]
S1.cb <- Sf.cb[, 3]
```

### Plot CB Counterfactual Survival Curves
```{r}
par(mfrow = c(2, 1))

plot(
  S0.cb, 
  type = "l", 
  lwd = 2,
  col = "red",
  ylim = c(0, 1),
  xlab = "Days", 
  ylab = "Survival Probability",
  main = "IPW Case Base: Counterfactual Survival Curves"
)

lines(
  S1.cb,
  lwd = 2,
  col = "blue"
)

legend(
  "topright",
  legend = c("A = 0", "A = 1"),
  col = c("red", "blue"),
  lty = 1, 
  lwd = 2
)

plot(
  S0.cb, 
  type = "l", 
  lwd = 2,
  col = "red",
  ylim = c(0.6, 1),
  xlim = c(0, 30),
  xlab = "Days", 
  ylab = "Survival Probability",
  main = "IPW Case Base: Counterfactual Survival Curves - First Month"
)

lines(
  S1.cb,
  lwd = 2,
  col = "blue"
)

legend(
  "topright",
  legend = c("A = 0", "A = 1"),
  col = c("red", "blue"),
  lty = 1, 
  lwd = 2
)
```

\newpage
# RMST Comparison: MS-Cox vs. IPW Case Base

### Calculate RMST for both models
```{r}
RMST.cox <- rmst.fn(tau.grid, S1.cox) - rmst.fn(tau.grid, S0.cox)

RMST.cb <- rmst.fn(tau.grid, S1.cb) - rmst.fn(tau.grid, S0.cb)
```

### Plot Both RMST
```{r}
cols <- c("#8E063B","#023FA5")

plot(
  RMST.cox,
  type = "l", 
  lwd = 2, 
  col = "red",
  xlab = "Tau (Days)", 
  ylab = "RMST Difference",
  main = "Causal RMST up to Tau"
)

lines(
  RMST.cb,
  lwd = 2,
  col = "blue"
)

abline(h = 0, lty = 2)

legend(
  "topright",
   legend = c("MS-Cox", "IPW Case Base"),
   col = c("red", "blue"),
   lwd = 2,
   lty = 1
)
```